### 2017-11-16 paper/manuscript.tex
Create manuscript template

% paper for Zone based traffic assignment algorithm. Short 4 page paper for ICT Express issue
% The aim here is to show an increase in scalibility with a preservation of traffic reduction (also note that the zone system will work with any model)

% for condensed version of paper, all pseudocode is dropped

% TODO:
% launch smaller zone experiments (should not take too long) (zone size of 5 is extremely slow) 
% generate graphs
% final passthrough
% done

\documentclass[conference]{IEEEtran}
\usepackage{mathtools}

\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx

\usepackage{graphicx}
\graphicspath{ {images/} }

\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
\raggedbottom

\title{An Improved Sequential Data Mining Model for Hurricane Trajectory Prediction}
% A novel algorithm for scalable traffic congestion reduction
% An adaptible model for scalability in traffic assignment algorithms
% Z-BAR: A Zone-Based traffic Assignment Algorithm for scalable congestion Reduction

% Is this an Algorithm or a framework: We want to highlight compatability

\author{\IEEEauthorblockN{Jasmin Bissonnette, Caden Marofke, Taylor Cox}
\IEEEauthorblockA{Department of Computer Science\\
University of Manitoba\\
Winnipeg, Manitoba\\
Email: \{bissonnj, marofkec, coxt3\}@myumanitoca.}}

\maketitle

\begin{abstract}

\end{abstract}

\begin{IEEEkeywords}
Hurricane Trajectory, Apriori, Prediction
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

\section{Introduction}

\section{Related Work}

% Methodology and Solution go here

\section{Experimental Setup}

% Need CSV output for metrics

\section{Experimental Results}

\section{Conclusions and Future Work} % what we found:

\begin{thebibliography}{1}
\bibitem{} 
\bibitem{} 
\bibitem{} 
\bibitem{} 
\bibitem{} 
\bibitem{} 
\bibitem{} 
\bibitem{} 
\bibitem{} 
\bibitem{} 
\end{thebibliography}

\end{document}

### 2017-11-18 test_best_fit.py
Introduce Euclidean distance option (Makes no change to result)

@@ -77,7 +77,20 @@ def getBestRule(rules, checkRegions):
   rules.seek(0)
   return bestRuleScore, bestRuleConsequent
 
-def getLatLongDistance(startCoord, endCoord):
+def getLatLongDistance(startCoord, endCoord, haversine=True):
+  if haversine:
+    return getHaversineDistance(startCoord, endCoord)
+  else:
+    return getEuclideanDistance(startCoord, endCoord)
+
+def getEuclideanDistance(startCoord, endCoord):
+  # lat, N/S, y, 0
+  # lon, W/E, x, 1 (negative)
+  # From Wikipedia:
+  # d(p, q) = sqrt( (q1-p1)**2 + (q2-p2)**2 )
+  return math.sqrt( abs( endCoord[1] + startCoord[1])**2 - abs(endCoord[0] - startCoord[0])**2 )
+
+def getHaversineDistance(startCoord, endCoord):
   # lat, N/S, y, 0
   # lon, W/E, x, 1 (negative)
   lon1, lat1, lon2, lat2 = map(math.radians, [startCoord[1], startCoord[0], endCoord[1], endCoord[0]])
@@ -302,7 +315,7 @@ def testRules():
     correctRatio = float(correctCount) / float(totalCount)
     print "Correct to incorrect ratio was %f. This included trajectory extrapolation" % correctRatio
 
-    print "Trajectory is working:", extrapWorks
+    print "Extrapolation is working:", extrapWorks
     print extrapCount
     print totalCount
     print correctCount

### 2017-11-18 basic_trajectory_draw.py
Introduce region overlay switch to draw script

@@ -5,14 +5,18 @@ import numpy as np
 import matplotlib.pyplot as plt
 import sys
 from mpl_toolkits.basemap import Basemap
+from apriori_utils import *
 
 def drawTrajectory():
   fraction = int(sys.argv[2])
+  showRegions = bool(int(sys.argv[3]))
   # Lambert Conformal Conic map.
-  m = Basemap(llcrnrlon=-100.,llcrnrlat=0.,urcrnrlon=-20.,urcrnrlat=57.,
+  m = Basemap(llcrnrlon=-100.,llcrnrlat=0.,urcrnrlon=-20.,urcrnrlat=50.,
               projection='lcc',lat_1=20.,lat_2=40.,lon_0=-60.,
               resolution ='l',area_thresh=1000.)
   # read shapefile.
+  # Need to write a longtitudional line as opposed to a euclidean horizontal line
+
   with open(sys.argv[1], 'r') as inputFile:
     i = 0
     for line in inputFile.readlines():
@@ -21,7 +25,6 @@ def drawTrajectory():
 
         parsedVector = [item.replace("(", "").replace(")", "") for item in lineVector]
         trajectoryVector = [ (float(item.split(":")[0]), float(item.split(":")[1])) for item in parsedVector]
-
         gisXCoords = []
         gisYCoords = []
         for coord in trajectoryVector:
@@ -30,7 +33,6 @@ def drawTrajectory():
           gisYCoords.append(int(newCoord[1]))
 
         m.plot(gisXCoords, gisYCoords, linewidth=0.5, color='r')
-      
       i += 1
 
   m.drawcoastlines()
@@ -39,12 +41,19 @@ def drawTrajectory():
   m.fillcontinents(color='#cc9966',lake_color='#99ffff')
   m.drawparallels(np.arange(10,70,20),labels=[1,1,0,0])
   m.drawmeridians(np.arange(-100,0,20),labels=[0,0,0,1])
+
+  if showRegions:
+    div = 80/float(LONGITUDE_DIVISIONS)
+    m.drawmeridians(np.arange(-100, -20, div),dashes=[1,0],color='b')
+    div2 = 50/float(LATITUDE_DIVISIONS)
+    m.drawparallels(np.arange(0, 50, div2),dashes=[1,0],color='b')
+
   plt.title('Atlantic Hurricane Tracks')
   plt.show()
 
 def main():
-  if len(sys.argv) != 3:
[…]

### 2017-11-23 tuple_regions/apriori_test.py
Start developing tuple-regions model (from Pi et al)

@@ -1,0 +1,117 @@
+
+from apriori_utils import *
+import sys
+import itertools
+import time
+
+# Need to dedupe raw vectors
+
+# For regionalization, need to represent vectors as a scalar and modulo by their bucket
+# How to make vector into a scalar? Norm of vector? Eigenvalue?
+
+# Some of the data entries are impossible (latitude cannot exceed 180. BTW the negative indicates west)
+
+# From some analysis, height is from 7.2->81.0, width from -109.5->63. 73.8 H-H, 172.5 W-W
+
+# As a first pass on regionalization, consider the following:
+""" Suppose we want to group the map into a 10x10 grid
+The question becomes "what sector does my x,y belong to"
+answer:
+73.8 / 10 -> 7.38. So take candidate x / 7.38, truncate to int, that's the x-region. If on the line (modulo = 0), take -1
+ditto for y.
+result: x*10 + y = region #
+
+# tuple regions are started but there is more work to do. First make sure multi-tuple-arrays are working correctly
+"""
+MIN_SUP = 5
+
+# region and safe_incr code moved to util file
+
+def gather_new_frequent_items(resultTable, frequent_k_itemsets):
+  resultTable.append({})
+  for key in frequent_k_itemsets:
+    if frequent_k_itemsets[key] >= MIN_SUP:
+      resultTable[-1][key] = frequent_k_itemsets[key] 
+
+def timeseries_candidate_generation(frequent_k_itemsets, chunk_size):
+  # We have our L_k, use suffix-prefix to get C_k+1
+  candidates = []
+  for curr_index, curr in enumerate(frequent_k_itemsets):
+    for compare_index, compare in enumerate(frequent_k_itemsets):
+
+      # if last chunk size of item equal to first chunksize of item2, add to cands
+      if curr_index != compare_index and curr[(chunk_size*-1):] == compare[:chunk_size]:
+        # we found a possible candidate but we need to do a prune check using frequent_k_itemsets
+        new_candidate = join_entries(curr, compare, chunk_size)
+        required_itemset_checks = itertools.combinations(new_candidate, chunk_size+1)
+        if all(item == curr or item == compare or item in frequent_k_itemsets for item in required_itemset_checks):
+          candidates.append(new_candidate)
+
+  return candidates
+
+def join_entries(left, right, chunk_size):
+  result = left + right[chunk_size:]
+  return result
+
+
+def execute_apriori():
+  with open(sys.argv[1], 'r') as vectors, open(sys.argv[2], 'w') as patterns:
+    level = 1
[…]

### 2017-11-23 tuple_regions/apriori_utils.py

@@ -1,0 +1,64 @@
+# Tuple Region style version. This is so we can emulate the paper's results
+
+LATITUDE_DIVISIONS = 11
+LONGITUDE_DIVISIONS = 18
+
+LAT_MAX_DEGREES = 50.0
+LAT_MIN_DEGREES = 0.0
+LONG_MAX_DEGREES = 100.0
+LONG_MIN_DEGREES = 20.0
+
+DIMENSIONS_LATITUDE = 5
+DIMENSIONS_LONGTITUDE = 5
+
+def gatherCoordinateList(inputStr):
+  parseStr = inputStr.split(' ')[1]
+  if parseStr[-1] == ',':
+    parseStr = parseStr[:-1]
+
+  result = []
+  parseStr = parseStr.strip()
+
+  for vecStr in parseStr.split(','):
+    if len(vecStr):
+      currVec = vecStr.replace('(', '').replace(')', '').split(':')
+      result.append( (float(currVec[0]), float(currVec[1])) )
+
+  return result
+
+def regionalize(coordinate):
+
+  # lat, N/S, y, 0
+  # lon, W/E, x, 1 (negative)
+  # Latitude is counted as degrees N/S of equator => y coords
+  # Longitude is counted as degrees W/E of meridian => x coords
+
+  # may need region pruning here
+
+  return (int(coordinate[0]/DIMENSIONS_LATITUDE), int(coordinate[1]/DIMENSIONS_LONGTITUDE))
+
+def gatherDedupedRegionList(ruleLine):
+  coordinates = gatherCoordinateList(ruleLine)
+  regionList = [regionalize(coord) for coord in coordinates]
+
+  dedupedRegions = []
+  # TODO: change this once we have bounds checking for data
+  i = 0
+  while i < len(regionList) and regionList[i] < 0:
+    i += 1
+  dedupedRegions.append(regionList[i])
+  foundItem = regionList[i]
+
+  while i < len(regionList):
+    if regionList[i] != foundItem and regionList[i] >= 0:
+      foundItem = regionList[i]
+      dedupedRegions.append(regionList[i])
+    i += 1
+
+  return dedupedRegions
+
+def safe_increment(table, key, value=1):
+  try:
+    table[key] += value
+  except KeyError:
+    table[key] = value

### 2017-11-23 tuple_regions/basic_trajectory_draw.py

@@ -1,0 +1,63 @@
+"""
+A Hurricane Trajectory-Rendering script based on examples from matplotlib
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+from mpl_toolkits.basemap import Basemap
+from apriori_utils import *
+
+def drawTrajectory():
+  fraction = int(sys.argv[2])
+  showRegions = bool(int(sys.argv[3]))
+  # Lambert Conformal Conic map.
+  m = Basemap(llcrnrlon=-100.,llcrnrlat=0.,urcrnrlon=-20.,urcrnrlat=50.,
+              projection='lcc',lat_1=20.,lat_2=40.,lon_0=-60.,
+              resolution ='l',area_thresh=1000.)
+  # read shapefile.
+  # Need to write a longtitudional line as opposed to a euclidean horizontal line
+
+  with open(sys.argv[1], 'r') as inputFile:
+    i = 0
+    for line in inputFile.readlines():
+      if i % fraction == 0: # Render 1 in fraction trajectories (nonzero)
+        lineVector = line.split(' ')[1].split(",")[:-1]
+
+        parsedVector = [item.replace("(", "").replace(")", "") for item in lineVector]
+        trajectoryVector = [ (float(item.split(":")[0]), float(item.split(":")[1])) for item in parsedVector]
+        gisXCoords = []
+        gisYCoords = []
+        for coord in trajectoryVector:
+          newCoord = m(coord[1], coord[0]) # Flip the input coordinates only on conversion
+          gisXCoords.append(int(newCoord[0]))
+          gisYCoords.append(int(newCoord[1]))
+
+        m.plot(gisXCoords, gisYCoords, linewidth=0.5, color='r')
+      i += 1
+
+  m.drawcoastlines()
+  m.drawcountries()
+  m.drawmapboundary(fill_color='#99ffff')
+  m.fillcontinents(color='#cc9966',lake_color='#99ffff')
+  m.drawparallels(np.arange(10,70,20),labels=[1,1,0,0])
+  m.drawmeridians(np.arange(-100,0,20),labels=[0,0,0,1])
+
+  if showRegions:
+    div = 80/float(LONGITUDE_DIVISIONS)
+    m.drawmeridians(np.arange(-100, -20, div),dashes=[1,0],color='b')
+    div2 = 50/float(LATITUDE_DIVISIONS)
[…]

### 2017-11-23 tuple_regions/find_data_extrema.py

@@ -1,0 +1,53 @@
+def gatherCoordinateList(inputStr):
+  parseStr = inputStr.split(' ')[1]
+  if parseStr[-1] == ',':
+    parseStr = parseStr[:-1]
+
+  result = []
+  parseStr = parseStr.strip()
+
+  for vecStr in parseStr.split(','):
+    if len(vecStr):
+      currVec = vecStr.replace('(', '').replace(')', '').split(':')
+      result.append( (float(currVec[0]), float(currVec[1])) )
+
+  return result
+
+import sys
+
+def get_extrema():
+  minX = 1000000000
+  minY = 1000000000
+
+  maxX = 0
+  maxY = 0
+
+  with open(sys.argv[1], 'r') as vectors:
+    for line in vectors:
+      coords = gatherCoordinateList(line)
+      for co in coords:
+        if co[0] < minX:
+          minX = co[0]
+
+        if co[0] > maxX:
+          maxX = co[0]
+
+        if co[1] < minY:
+          minY = co[1]
+
+        if co[1] > maxY:
+          maxY = co[1]
+
+  print minX, maxX, minY, maxY
+  print maxX - minX, maxY - minY
+
+def main():
+  if len(sys.argv) != 2:
+    print "Usage: {0} <VectorsInput>".format(sys.argv[0])
+  else:
+    get_extrema()
+
+if __name__ == '__main__':
+  main()
+else:
+  print __name__

### 2017-11-23 tuple_regions/generate_assoc_rules.py

@@ -1,0 +1,113 @@
+# Another quick n dirty python script to generate association rules given a list of frequent itemsets
+# It is assumed that each line in the file is of the form (IS) : Supp
+# Where IS is a ", " delimited series of numbers, which correspond to regions
+# Don't look at singletons, they cannot produce rules
+
+import sys
+import time
+import re
+
+MIN_CONF = 0.25
+
+def getRawRules(itemset):
+  divider = len(itemset) - 1
+  rules = [] # Antecedent, Consequent pairs
+  while divider > 0:
+    antecedent = itemset[0:divider]
+    consequent = itemset[divider:]
+    rules.append( (antecedent, consequent) )
+    divider -= 1
+
+  return rules
+
+def convertToSearchLine(antecedentArray):
+  if len(antecedentArray) == 1:
+    line = antecedentArray[0]
+  else:
+    line = "(" + ", ".join(map(str, antecedentArray)) + ")"
+
+  # Will have to change this to parse strings correctly
+  return str(line).replace("'", "")
+
+def getAntecedentSupport(itemsetsFile, searchLine):
+  with open(itemsetsFile, "r") as itemsets:
+    for line in itemsets:
+      contents = line.split(":")
+      if contents[0].strip() == searchLine:
+        result = int(contents[1].strip())
+
+  return result
+
+def generateRules(itemsetsFile, currItemsetLine): # String line of form (t, t, t) : s
+  generatedRules = []
+
+  currItemsetLine = currItemsetLine.replace(" ", "")
+  currItemsetParsed = currItemsetLine.split(":")
+
+  currItemsetSupport = int(currItemsetParsed[1])
+
+  # Change this to list of tuples
+  currItemset = convertToTupleList(currItemsetParsed[0])
+  # Now we have the full itemset, and its support
+  # These are ordered rules, so we grow the antecedent and shrink the consequent together
+  # ABC, AB>C, A>BC
+  rulePartitions = getRawRules(currItemset)
+
+  # To find confidence, we need support of antecedent
+  for rule in rulePartitions:
+    searchLine = convertToSearchLine(rule[0])
+    # Do a search in the file for the corresponding frequent itemset, and get its support
+    # That gives the confidence of the rule to be inserted
+    antecedentSupport = getAntecedentSupport(itemsetsFile, searchLine)
+
[…]

### 2017-11-23 tuple_regions/test_best_fit.py

@@ -1,0 +1,356 @@
+# Given an incoming trajectory:
+# Cut it into two pieces: check, confirm
+# Check "check" against rule antecedents, get best fit (or show "missing")
+# if not missing, see if "confirm" was right
+
+# To do this, we will need to cut the patterns into two pieces. Training and Testing.
+# First try: quarters. 1335 training 445 testing
+
+# First try: divide test line in half. First half looks for antecedent
+
+# Parse a line into the antecedent, list of strings
+
+import math
+import sys
+import time
+import re
+from apriori_utils import *
+
+EXTRAPOLATION_LIMIT = 10
+EARTH_RADIUS = 6371
+
+def getAttributes(ruleLine):
+  ruleLine = ruleLine.replace(" ", "")
+  confParsed = ruleLine.split(":")
+
+  confidence = float(confParsed[1])
+
+  ruleParsed = confParsed[0].split("=>")
+
+  antecedent = strToTupleArray(ruleParsed[0])
+  consequent = strToTupleArray(ruleParsed[1])
+
+  return antecedent, consequent, confidence
+
+def getMatchingLen(antecedent, checkRegions):
+  # We have antecedent A1 A2 ... An
+  # and regions R1 R1 ... Rm
+  # What is the longest sequence of A's present in R's?
+  # Note that the sequence has to be ordered too
+  maxLen = 0
+  i = 0
+  j = 0
+
+  while i < len(checkRegions):
+    currLen = 0
+    currItem = checkRegions[i]
+    if currItem in antecedent:
+      index = antecedent.index(currItem)
+      j = i
+
+      while j < len(checkRegions) and index < len(antecedent) and checkRegions[j] == antecedent[index]:
+        j += 1
+        index += 1
+        currLen += 1
+
+    if currLen > maxLen:
+      maxLen = currLen
+    i += 1
+
+  return maxLen
+
+def strToTupleArray(arrayString):
+  # May have to fix when we see tuples of size > 1
+  # some code repitition here but that's alright for now
+  tupleArray = []
+  result = arrayString.replace("'", "").replace("[", "").replace("]", "")
+  rawTuples = re.findall("\([0-9-,]+\)" , result)
+  for rawTuple in rawTuples:
+    leftT, rightT = rawTuple[1:len(rawTuple)-1].split(",")
+    tupleArray.append( (leftT, rightT) )
+
+  return tupleArray
+
[…]

### 2017-11-24 test_best_fit.py
Finish tuple regions. Fix Euclidean distance bug. Impl pattern matching as per paper. Tweak both models and training/testing data intervals.

@@ -17,6 +17,7 @@ from apriori_utils import *
 
 EXTRAPOLATION_LIMIT = 10
 EARTH_RADIUS = 6371
+MAX_MINIMUM_MATCH = 5
 
 def getAttributes(ruleLine):
   ruleLine = ruleLine.replace(" ", "")
@@ -91,7 +92,7 @@ def getEuclideanDistance(startCoord, endCoord):
   # lon, W/E, x, 1 (negative)
   # From Wikipedia:
   # d(p, q) = sqrt( (q1-p1)**2 + (q2-p2)**2 )
-  return math.sqrt( abs( endCoord[1] + startCoord[1])**2 - abs(endCoord[0] - startCoord[0])**2 )
+  return math.sqrt( abs( endCoord[1] - startCoord[1])**2 + abs(endCoord[0] - startCoord[0])**2 )
 
 def getHaversineDistance(startCoord, endCoord):
   # lat, N/S, y, 0
@@ -200,8 +201,8 @@ def getCoordCriteriaMatch(currCoord, currCoordList, compareCoordList, currCoordI
 def coordsMatchCriteria(leftCoord, rightCoord, leftCoordList, rightCoordList, leftIndex, rightIndex):
   # See if the two coordinates are within threshold distance and bearing difference
   # These constants are subject to change wrt experiments
-  maxDistanceDiff = 150
-  maxBearingDiff = 10
+  maxDistanceDiff = 100
+  maxBearingDiff = 1
   match = False
   # First, check distance comparison
   distance = getLatLongDistance(leftCoord, rightCoord)
@@ -224,19 +225,47 @@ def coordsMatchCriteria(leftCoord, rightCoord, leftCoordList, rightCoordList, le
 
   return match
 
-def compareResults(bestRuleConsequent, confirmRegions):
+def formerItemsWithinDistance(bestRuleConsequent, confirmRegions):
+  numConseq = len(bestRuleConsequent)
+  numConfrm = len(confirmRegions)
+
+  minLen = min(numConseq, numConfrm) \
+  if min(numConseq, numConfrm) < MAX_MINIMUM_MATCH \
+  else MAX_MINIMUM_MATCH
+
+  bestRuleConsequent = map(characteristicCoordinate, [int(region) for region in bestRuleConsequent[:minLen]])
+  confirmRegions = map(characteristicCoordinate, [int(region) for region in confirmRegions[:minLen]])
+  if minLen == 0:
+    return False
[…]

### 2017-11-24 tuple_regions/apriori_test.py

@@ -23,7 +23,7 @@ result: x*10 + y = region #
 
 # tuple regions are started but there is more work to do. First make sure multi-tuple-arrays are working correctly
 """
-MIN_SUP = 5
+MIN_SUP = 3
 
 # region and safe_incr code moved to util file
 

### 2017-11-24 tuple_regions/test_best_fit.py

@@ -18,6 +18,7 @@ from apriori_utils import *
 
 EXTRAPOLATION_LIMIT = 10
 EARTH_RADIUS = 6371
+MAX_MINIMUM_MATCH = 5 # Max comparison length from paper
 
 def getAttributes(ruleLine):
   ruleLine = ruleLine.replace(" ", "")
@@ -76,7 +77,6 @@ def getBestRule(rules, checkRegions, flex=False):
   bestRuleConsequent = []
   for rule in rules:
     antecedent, consequent, confidence = getAttributes(rule)
-
     if flex:
       matchingLen = getLongestSimilarSequence(antecedent, checkRegions)
     else:
@@ -101,7 +101,7 @@ def getEuclideanDistance(startCoord, endCoord):
   # lon, W/E, x, 1 (negative)
   # From Wikipedia:
   # d(p, q) = sqrt( (q1-p1)**2 + (q2-p2)**2 )
-  return math.sqrt( abs( endCoord[1] + startCoord[1])**2 - abs(endCoord[0] - startCoord[0])**2 )
+  return math.sqrt( abs( endCoord[1] - startCoord[1])**2 + abs(endCoord[0] - startCoord[0])**2 )
 
 def getHaversineDistance(startCoord, endCoord):
   # lat, N/S, y, 0
@@ -232,22 +232,49 @@ def coordsMatchCriteria(leftCoord, rightCoord, leftCoordList, rightCoordList, le
 
   return match
 
-def compareResults(bestRuleConsequent, confirmRegions):
+def formerItemsWithinDistance(bestRuleConsequent, confirmRegions):
+  numConseq = len(bestRuleConsequent)
+  numConfrm = len(confirmRegions)
+
+  minLen = min(numConseq, numConfrm) \
+  if min(numConseq, numConfrm) < MAX_MINIMUM_MATCH \
+  else MAX_MINIMUM_MATCH
+
+  if minLen == 0:
+    return False
+  bestRuleConsequent = [ ( float(x), float(y) ) for x, y in bestRuleConsequent[:minLen]]
+  confirmRegions = [ ( float(x), float(y) ) for x, y in confirmRegions[:minLen]]
+
+  # Make sure all entries have distance <= 1
+  # euclidean distance or haversine distance?
+  distInRange = True
+  i = 0
+  while i < minLen and distInRange:
+    distInRange = getEuclideanDistance(confirmRegions[i], bestRuleConsequent[i]) <= 1
+    i += 1
+
+  return distInRange
[…]

### 2017-11-24 tuple_regions/apriori_test.py
Introduce apriori optimizations

@@ -28,24 +28,62 @@ MIN_SUP = 3
 # region and safe_incr code moved to util file
 
 def gather_new_frequent_items(resultTable, frequent_k_itemsets):
+  print "go"
   resultTable.append({})
   for key in frequent_k_itemsets:
     if frequent_k_itemsets[key] >= MIN_SUP:
       resultTable[-1][key] = frequent_k_itemsets[key] 
+  print "done--"
 
 def timeseries_candidate_generation(frequent_k_itemsets, chunk_size):
-  # We have our L_k, use suffix-prefix to get C_k+1
-  candidates = []
-  for curr_index, curr in enumerate(frequent_k_itemsets):
-    for compare_index, compare in enumerate(frequent_k_itemsets):
+  # this is the candidate generation and pruning
 
-      # if last chunk size of item equal to first chunksize of item2, add to cands
-      if curr_index != compare_index and curr[(chunk_size*-1):] == compare[:chunk_size]:
-        # we found a possible candidate but we need to do a prune check using frequent_k_itemsets
-        new_candidate = join_entries(curr, compare, chunk_size)
-        required_itemset_checks = itertools.combinations(new_candidate, chunk_size+1)
-        if all(item == curr or item == compare or item in frequent_k_itemsets for item in required_itemset_checks):
-          candidates.append(new_candidate)
+  # looks like this is the bottleneck right now (at least it's the worst one)
+  # need to reduce the search space. Should be able to search only a specific subset of the candidate entries (see lecture slides)
+  # We have our L_k, use suffix-prefix to get C_k+1
+
+  # how to speed this way up
+  # take advantage of the fact that the data is sorted now
+
+  candidates = []
+  for curr_index, curr_item in enumerate(frequent_k_itemsets):
+    print curr_index
+
+    compare_index = 0
+    found = False
+    leftChunk = curr_item[(chunk_size*-1):]
+    #print "slower1?"
+    while compare_index < len(frequent_k_itemsets) and not found:
+      found = leftChunk == frequent_k_itemsets[compare_index][:chunk_size]
+      if not found:
+        compare_index += 1
+
+    #print "slower2?"
+    while compare_index < len(frequent_k_itemsets) and leftChunk == frequent_k_itemsets[compare_index][:chunk_size]:
+      new_candidate = join_entries(curr_item, frequent_k_itemsets[compare_index], chunk_size)
+      required_itemset_checks = itertools.combinations(new_candidate, chunk_size+1)
+
+      #print "slower3?"
[…]

### 2017-11-25 tuple_regions/apriori_test.py
Optimize Apriori. Gather new data. Tweak models. Prep for paper

@@ -28,12 +28,10 @@ MIN_SUP = 3
 # region and safe_incr code moved to util file
 
 def gather_new_frequent_items(resultTable, frequent_k_itemsets):
-  print "go"
   resultTable.append({})
   for key in frequent_k_itemsets:
     if frequent_k_itemsets[key] >= MIN_SUP:
       resultTable[-1][key] = frequent_k_itemsets[key] 
-  print "done--"
 
 def timeseries_candidate_generation(frequent_k_itemsets, chunk_size):
   # this is the candidate generation and pruning
@@ -44,10 +42,9 @@ def timeseries_candidate_generation(frequent_k_itemsets, chunk_size):
 
   # how to speed this way up
   # take advantage of the fact that the data is sorted now
-
+  frequent_k_itemsets_dict = { tuple(item):None for item in frequent_k_itemsets }
   candidates = []
   for curr_index, curr_item in enumerate(frequent_k_itemsets):
-    print curr_index
 
     compare_index = 0
     found = False
@@ -64,10 +61,11 @@ def timeseries_candidate_generation(frequent_k_itemsets, chunk_size):
       required_itemset_checks = itertools.combinations(new_candidate, chunk_size+1)
 
       #print "slower3?"
-      if all(item == curr_item or item == frequent_k_itemsets[compare_index] or item in frequent_k_itemsets for item in required_itemset_checks):
+      if all(item == curr_item or item == frequent_k_itemsets[compare_index] or tuple(item) in frequent_k_itemsets_dict for item in required_itemset_checks):
         candidates.append(new_candidate)
       compare_index += 1
-    print "out"
+
+
     # read my lips: no new comparisons!
 
   # candidates = []
@@ -130,6 +128,7 @@ def execute_apriori():
       level += 1
 
     else: # cand gen/prune
+      # scan
       # we are given C_k, , read table, add L_k to final result, generate/prune C_k+1, incr level
       # may want to optimize this too. faster search needed
       # gonna have to eat this one. no faster way it seems

### 2017-11-25 tuple_regions/test_best_fit.py

@@ -250,7 +250,7 @@ def formerItemsWithinDistance(bestRuleConsequent, confirmRegions):
   distInRange = True
   i = 0
   while i < minLen and distInRange:
-    distInRange = getEuclideanDistance(confirmRegions[i], bestRuleConsequent[i]) <= 1
+    distInRange = int(getEuclideanDistance(confirmRegions[i], bestRuleConsequent[i])) <= 1
     i += 1
 
   return distInRange

### 2017-11-25 tuple_regions/unmangle_data.py

@@ -1,0 +1,8 @@
+import sys
+
+with open(sys.argv[1], 'r') as inFile, open(sys.argv[2], 'w') as outFile:
+  for line in inFile:
+    for word in line.split(" "):
+      if len(word) > 0:
+        outFile.write(word+' ')
+

### 2017-11-26 paper/manuscript.tex
Continue tweaking model. Set final datasets. Introduce distance formula writeup

@@
 Winnipeg, Manitoba\\
-Email: \{bissonnj, marofkec, coxt3\}@myumanitoca.}}
+Email: \{bissonnj, marofkec, coxt3\}@myumanitoba.ca}}
 
@@
 
+\subsubsection{Distance Formulas}
+
+% A brief writeup on Haversine vs Eudlicean Distance:
+
+% - What each distance formula is for, and wrt the problem domain
+% - What the paper did vs what we did
+% - Compare different model results
+
+% Distance Formulas:
+
+Accurate distance measurement is a critical task in hurricane trajectory prediction.
+Pattern matching strategies in the proposed hurricane trajectory prediction model depend on coherent distance measurements between real-world spatial points.
+Consider the following prediction-correctness evaulation method proposed by Dong et al (1).
+To determine whether a trajectory prediction is correct, the distance between the first p points of the predicted and actual trajectories is compared against a defined maximum, where p is the minimum length of the two trajectories.
+If all p pairs of points fit within the maximum distance, the trajectory prediction is considered correct.
+
+Two different distance measurement formulas are evaluated in this work.
+This work compares the Euclidean distance formula against the Haversine distance formula.
+The Euclidean distance formula (as used in (1)) is designed to determine the unit distance between two N dimensional points in a straight line.
+In the proposed model, units correspond to degrees of latitude and longtitude.
+In contrast, the Haversine formula is designed to determine the kilometers between two latitude-longtitude points over a large sphere (such as the earth).
+For this reason, the Haversine formula is anticipated to provide a more accurate definition of distance between recorded hurricane points.
+When used in this work, the Euclidean distance formula is evaluated with two dimensions, correspoding respectively to latitude and longtitude.
+The two formulas, denoted Deucl and Dhavr are defined as follows:
+
+% figure 1: Euclidean and Haversine Distance Formulas
+
+% D_eucl(p1, p2) = sqrt( (lat_p2 - lat_p1)^2 + (lon_p2 - lon_p1)^2 )
+
+% D_havr(p1, p2) = < this thing is a monster >
+
+% // Explain any variables. Especially constants like radius of earth
+
+To compare the prediction accuracy between the two proposed formulas, a sample execution of the hurricane trajectory prediction model is executed with identical variables, except for the difference in distance evaluation.
+The distance formulas are interchanged in the prediction correctness checking step of the model, described previously in this section.
+When using the Euclidean formula, the distance threshold is 1.
+This corresponds to the distance threshold proposed in (1).
+When the Haversine formula is employed, the maximal distance is set to 450km.
+This value corresponds to slightly less than approximately one average hurricane-diameter.
+Since the average hurricane diameter is between 333 and 670km (2), two points are considered to be within the distance threshold if they are less than one average hurricane-diameter apart.
+
+Between the Euclidean and Haversine distance formulas, the Haversine formula delivers a increased correctness rate of 39.3\%, corresponding to 72.7\% compared to 33.4\% correctness when the Euclidean formula was employed.
+The realistic nature of the Haversine formula in conjunction with its corresponding maximal distance threshold is able to provide a higher rate of accurate predictions compared to the Euclidean distance formula, which is not designed for non-euclidean coordinate systems such as coordinate systems over spheres.
+
+Table of parameters:
+
+- Weighted Training data:   No         No
+- Training Data:            1950-2000  1950-2000
+- Testing Data:             2001-2015  2001-2015
+- Minsup:                   30         30
+- Minconf:                  0.25       0.25
+- Region Method:            DC         DC
+- Fitness Function:         [1]        [1]
+- Rule Fit Flexible:        No         No
+- Correctness Method:       [1]        [1]
+- Distance Formula:         Euclidean  Haversine
+- Distance Threshold:       1 (unit)   450 (km)
+------------------------------------------------
+- Correctness ratio         33.4\%     72.7\%
+
+(DC: Disretized Coordinates)
+
+% [1] Dong, X & C. Pi, D. (2013). Novel method for hurricane trajectory prediction based on data mining. Natural Hazards and Earth System Sciences. 13. 3211-3220.
+
+% [2] http://www.usno.navy.mil/JTWC/frequently-asked-questions-1/frequently-asked-questions#tcsize (Working on finding better source...)
+
 % Need CSV output for metrics

### 2017-11-26 apriori_test.py

@@ -23,7 +23,7 @@ result: x*10 + y = region #
 
 
 """
-MIN_SUP = 50
+MIN_SUP = 30
 
 # region and safe_incr code moved to util file
 

### 2017-11-26 test_best_fit.py

@@ -201,8 +201,8 @@ def getCoordCriteriaMatch(currCoord, currCoordList, compareCoordList, currCoordI
 def coordsMatchCriteria(leftCoord, rightCoord, leftCoordList, rightCoordList, leftIndex, rightIndex):
   # See if the two coordinates are within threshold distance and bearing difference
   # These constants are subject to change wrt experiments
-  maxDistanceDiff = 100
-  maxBearingDiff = 1
+  maxDistanceDiff = 300
+  maxBearingDiff = 10
   match = False
   # First, check distance comparison
   distance = getLatLongDistance(leftCoord, rightCoord)
@@ -243,7 +243,7 @@ def formerItemsWithinDistance(bestRuleConsequent, confirmRegions):
   i = 0
   while i < minLen and distInRange:
 
-    distInRange = getLatLongDistance(confirmRegions[i], bestRuleConsequent[i]) <= 300
+    distInRange = getLatLongDistance(confirmRegions[i], bestRuleConsequent[i]) <= 600 # Something is really wrong with the block region model. This gets 10%.
     i += 1
 
   return distInRange
@@ -317,7 +317,7 @@ def testRules():
       searchRounds = 0
       complete = False
       while not complete:
-        bestRuleScore, bestRuleConsequent = getBestRule(rules, checkRegions, flex=False)
+        bestRuleScore, bestRuleConsequent = getBestRule(rules, checkRegions, flex=True)
         if bestRuleScore == 0 and searchRounds < EXTRAPOLATION_LIMIT:
           checkRegions.append(
             getExtrapolatedRegion(

### 2017-11-26 tuple_regions/apriori_test.py

@@ -23,7 +23,7 @@ result: x*10 + y = region #
 
 # tuple regions are started but there is more work to do. First make sure multi-tuple-arrays are working correctly
 """
-MIN_SUP = 3
+MIN_SUP = 30
 
 # region and safe_incr code moved to util file
 

### 2017-11-26 tuple_regions/test_best_fit.py

@@ -250,7 +250,7 @@ def formerItemsWithinDistance(bestRuleConsequent, confirmRegions):
   distInRange = True
   i = 0
   while i < minLen and distInRange:
-    distInRange = int(getEuclideanDistance(confirmRegions[i], bestRuleConsequent[i])) <= 1
+    distInRange = int(getHaversineDistance(confirmRegions[i], bestRuleConsequent[i])) <= 450
     i += 1
 
   return distInRange

### 2017-11-28 tuple_regions/apriori_test.py
Introduce weighted data. Introduce switch to ignore extraps. More assets, graphs, writeups

@@ -23,25 +23,25 @@ result: x*10 + y = region #
 
 # tuple regions are started but there is more work to do. First make sure multi-tuple-arrays are working correctly
 """
-MIN_SUP = 30
+WEIGHT_INTERVAL_NOMINAL = 756
+WEIGHT_INTERVAL_YEAR = 50
+HURRICANE_ID_PLACEHOLDER = "AL015555"
+WEIGHT_SCALE = 7.5
+# For weighted data:
+# Data can be weighted nominally or year-over-year
+# Careful! Right now these values are dependent on the specific training data input. This must be fixed
+# A good minsupport value for weight training data appears to be 10. Minconf can stay the same
 
-# region and safe_incr code moved to util file
-
-def gather_new_frequent_items(resultTable, frequent_k_itemsets):
+def gather_new_frequent_items(resultTable, frequent_k_itemsets, minsup):
   resultTable.append({})
   for key in frequent_k_itemsets:
-    if frequent_k_itemsets[key] >= MIN_SUP:
+    if frequent_k_itemsets[key] >= minsup:
       resultTable[-1][key] = frequent_k_itemsets[key] 
 
 def timeseries_candidate_generation(frequent_k_itemsets, chunk_size):
   # this is the candidate generation and pruning
-
-  # looks like this is the bottleneck right now (at least it's the worst one)
-  # need to reduce the search space. Should be able to search only a specific subset of the candidate entries (see lecture slides)
   # We have our L_k, use suffix-prefix to get C_k+1
 
-  # how to speed this way up
-  # take advantage of the fact that the data is sorted now
   frequent_k_itemsets_dict = { tuple(item):None for item in frequent_k_itemsets }
   candidates = []
   for curr_index, curr_item in enumerate(frequent_k_itemsets):
@@ -49,39 +49,19 @@ def timeseries_candidate_generation(frequent_k_itemsets, chunk_size):
     compare_index = 0
     found = False
     leftChunk = curr_item[(chunk_size*-1):]
-    #print "slower1?"
+
     while compare_index < len(frequent_k_itemsets) and not found:
       found = leftChunk == frequent_k_itemsets[compare_index][:chunk_size]
       if not found:
         compare_index += 1
 
-    #print "slower2?"
     while compare_index < len(frequent_k_itemsets) and leftChunk == frequent_k_itemsets[compare_index][:chunk_size]:
       new_candidate = join_entries(curr_item, frequent_k_itemsets[compare_index], chunk_size)
[…]

### 2017-11-28 tuple_regions/apriori_utils.py

@@ -57,7 +57,7 @@ def gatherDedupedRegionList(ruleLine):
 
   return dedupedRegions
 
-def safe_increment(table, key, value=1):
+def safe_increment(table, key, value=1.0):
   try:
     table[key] += value
   except KeyError:

### 2017-11-28 tuple_regions/generate_assoc_rules.py

@@ -34,7 +34,7 @@ def getAntecedentSupport(itemsetsFile, searchLine):
     for line in itemsets:
       contents = line.split(":")
       if contents[0].strip() == searchLine:
-        result = int(contents[1].strip())
+        result = float(contents[1].strip())
 
   return result
 
@@ -44,7 +44,7 @@ def generateRules(itemsetsFile, currItemsetLine): # String line of form (t, t, t
   currItemsetLine = currItemsetLine.replace(" ", "")
   currItemsetParsed = currItemsetLine.split(":")
 
-  currItemsetSupport = int(currItemsetParsed[1])
+  currItemsetSupport = float(currItemsetParsed[1])
 
   # Change this to list of tuples
   currItemset = convertToTupleList(currItemsetParsed[0])

### 2017-11-28 tuple_regions/test_best_fit.py

@@ -310,6 +310,8 @@ def getLongestSimilarSequence(ruleRegions, testRegions):
   return resultLen
 
 def testRules():
+  ignoreExtrapolation = bool( int( sys.argv[3] ) )
+
   with open(sys.argv[1], 'r') as test_vectors, open(sys.argv[2], 'r') as rules:
     correctCount = 0
     totalCount = 0
@@ -326,6 +328,10 @@ def testRules():
       complete = False
       while not complete:
         bestRuleScore, bestRuleConsequent = getBestRule(rules, checkRegions, flex=False)
+        if bestRuleScore == 0 and ignoreExtrapolation:
+          complete = True
+          totalCount -= 1
+          break
 
         if bestRuleScore == 0 and searchRounds < EXTRAPOLATION_LIMIT:
           checkRegions.append(
@@ -372,8 +378,8 @@ def testRules():
     print correctCount
 
 def main():
-  if len(sys.argv) != 3:
-    print "Usage: {0} <VectorsInput> <RulesInput>".format(sys.argv[0])
+  if len(sys.argv) != 4:
+    print "Usage: {0} <VectorsInput> <RulesInput> <IgnoreExtrapolation0/1>".format(sys.argv[0])
   else:
     testRules()
 

### 2017-11-28 tuple_regions/apriori_test.py
Make sure default apriori weight is 1

@@ -78,12 +78,12 @@ def gather_sublists(region_list):
   return result
 
 def init_support_weight(weightedData, weightedNominal):
-  w = WEIGHT_SCALE
+  w = 1.0
   if weightedData:
     if weightedNominal:
-      w /= WEIGHT_INTERVAL_NOMINAL
+      w = WEIGHT_SCALE / WEIGHT_INTERVAL_NOMINAL
     else:
-      w /= WEIGHT_INTERVAL_YEAR
+      w = WEIGHT_SCALE / WEIGHT_INTERVAL_YEAR
 
   return w
 

### 2017-12-03 paper/manuscript.tex
Write paper draft up to AprioriAll

--- 2017-11-16 paper/manuscript.tex
@@
-% paper for Zone based traffic assignment algorithm. Short 4 page paper for ICT Express issue
-% The aim here is to show an increase in scalibility with a preservation of traffic reduction (also note that the zone system will work with any model)
-
-% for condensed version of paper, all pseudocode is dropped
-
-% TODO:
-% launch smaller zone experiments (should not take too long) (zone size of 5 is extremely slow) 
-% generate graphs
-% final passthrough
-% done
-
-\documentclass[conference]{IEEEtran}
+\documentclass[12pt,conference]{IEEEtran}
 \usepackage{mathtools}
@@
 \usepackage{graphicx}
+\usepackage{textcomp}
 \graphicspath{ {images/} }
@@
 
-\title{An Improved Sequential Data Mining Model for Hurricane Trajectory Prediction}
-% A novel algorithm for scalable traffic congestion reduction
-% An adaptible model for scalability in traffic assignment algorithms
-% Z-BAR: A Zone-Based traffic Assignment Algorithm for scalable congestion Reduction
-
-% Is this an Algorithm or a framework: We want to highlight compatability
+\title{Improving the Accuracy of a Sequential Mining Model for Hurricane Trajectory Prediction}
 
@@
 
+In the 21st century alone, hurricanes have caused billions of dollars worth of damage due to property destruction, personal danger, and civil unrest.
+The catastrophic impact of hurricanes and other tropical storms has motivated extensive research into hurricane forecasting.
+Hurricane trajectory prediction is a subfield of hurricane forecasting that traditionally uses meteorological analysis to determine the future path of a given hurricane.
+While meteorological methods of hurricane trajectory prediction have been effective in the past, little attention has been paid to hurricane trajectory prediction with respect to historical frequent pattern mining.
+In this work, we propose multiple dimensions for improving the accuracy of datamining-based hurricane trajectory prediction.
+We propose the use of Haversine distance, a real-world distance formula for improved hurricane pattern matching accuracy.
+We also propose the use of weighted data to favor recent hurricane trends when training the trajectory prediction model.
+Finally, we use training data exclusively from the years 1950 to 2000 to reduce the impact of historical inaccuracies in hurricane recording methodologies.
+After implementing all three improvement dimensions on a contemporary data mining model for hurricane trajectory prediction based on AprioriAll, results show a prediction-correctness rate 10.22\% higher than the current state-of-the-art, at 75.22\%.
+
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Hurricane Trajectory, Apriori, Prediction
+Hurricane Trajectory, AprioriAll, Prediction, Pattern Matching
 \end{IEEEkeywords}
@@
 
+Hurricanes are a specific type of tropical cyclone exclusive to the Atlantic basin \cite{def-hurricane}.
+Hurricanes are divided into five categories according the Saffir-Simpson scale, based on wind-speed and expected damage \cite{hurricane-cat}.
+In the United States alone, hurricanes incur damages valued as high as \$157BN \cite{hurricane-cost}.
+The damages incurred by hurricanes do not only include property damage, but also personal danger and in some cases widespread civil unrest \cite{hurricane-cost-non-financial}.
+These damages motivate extensive research in the field of hurricane prediction analysis.
+Hurricane prediction analysis is traditionally divided into the subfields of hurricane intensity prediction and hurricane trajectory prediction.
+This paper focuses exclusively on the subfield of hurricane trajectory prediction.
+In particular, the goal of this paper is to investigate and improve on the existing-state-of-the art in hurricane trajectory prediction systems based on sequential datamining.
+
+Data mining is the nontrivial extraction of implicit, previously unknown, and potentially useful information from data \cite{datamining-def}.
+Data mining applies to an extremely wide range of applications, including clickstream mining, social network community detection, and artificial intelligence.
+The datamining subdiscipline of particular interest in this work is \textit{sequential} data mining.
+Sequential data mining uses a database of chronologically ordered inputs to mine frequent patterns and interesting association rules.
+These association rules are used to make predictions on future data entries such as future customer transactions.
+More specifically, association rules generated by sequential data mining algorithms including AprioriAll, SPIRIT and PrefixSpan can be used in the field of hurricane trajectory prediction.
+
+To the knowledge of the authors, the existing standard in hurricane trajectory prediction via data mining is developed by Dong et al in \cite{major-hurricane-model}.
+The work cited proposes a hurricane trajectory prediction model based on datamining (HTPDM) that uses a modified variant of AprioriAll, a common sequential mining algorithm \cite{AprioriAll-original}.
+HTPDM executes AprioriAll against historical hurricane trajectories from 1900 to 2000 to generate interesting trajectory association rules.
+These association rules are compared against test data composed of hurricane trajectories from 2001 to 2008.
+HTPDM divides a given testing trajectory into two parts: the initial trajectory $T_{i}$ and the terminal trajectory $T_{t}$.
+$T_{i}$ is compared against all association rule antecedents until a sufficiently matching rule R is found based on a predetermined fitness function.
+The consequent of R is then compared against $T_{t}$ using pattern-matching techniques to determine if the prediction was correct.
+The results of HTPM showed that the model was able to achieve a best-case correctness ratio of 65\%.
+In this work, HTPDM will be developed and various strategies will be employed to improve the correctness ratio.
+Reducing the impact of historical factors, favoring the impact of recent trends, and improving the realistic nature of the hurricane trajectory prediction model result in a new model dubbed WARD-HTP (Weighted-Asset, Realistic-Distance Hurricane Trajectory Predictor).
+The strategies employed in WARD-HTP ultimately result in a correctness ratio of 75.22\%, a 10.22\% improvement over HTPDM.
+
+The remainder of this paper is structured as follows: First, Section II gives a detailed survey of related work in the hurricane prediction and sequential mining disciplines.
+Next, section III will cover the general motivations of this work with respect to hurricane trajectory prediction, and the specific motivations of this work with respect to improving the current state-of-the-art.
+Section IV will provide a description of the problem space, which includes specific implementation considerations in hurricane trajectory prediction.
+Section V will give an exposition of the solutions proposed in WARD-HTP for increasing the correctness ratio in hurricane trajetory prediction.
+Following that, Section VI will detail the experimental setup, including the training data and other experimental variables.
+Section VII will evaluate the results of the experiments, which show that the ideal configurations of WARD-HTP result in a maximal correctness ratio of 75.22\%.
+This work will then conclude with section VIII, capturing the conclusions and directions of future work.
+
 \section{Related Work}
 
-% Methodology and Solution go here
+The development of an accurate hurricane trajectory prediction model is a problem which has received little attention in the data mining literature.
+Works including \cite{hurricane-intensity-1} and \cite{hurricane-intensity-2} have been completed to develop data-mining approaches to hurricane intensity prediction.
+These works however do not describe approaches to hurricane trajectory prediction.
+Hurricane trajectory prediction has been covered in previous meteorological and geographic information system (GIS) works.
+These types of hurricane trajectory forecasting models include those seen in \cite{hurricane-forecasting-1} or \cite{hurricane-forecasting-2}.
+The intersection between data mining and hurricane trajectory prediction has been sparsely investigated in the literature.
+Publications such as \cite{typhoon-clustering} proposed a clustering-based model for typhoon track prediction based on specific characteristics of historical data.
+Typhoons are then clustered into groups depending on the general trajectory patterns they exhibit.
+While the described work provided novel results in typhoon track clustering, it is not clear if the proposed solution is able to predict the trajectories of new storms.
+Additionally, the clusters developed apply specific to typhoons, which are waterborne storms native to the east-pacific seaboard.
+Such storm patterns may not be applicable to patterns found in hurricanes, which are waterborne storms of the west-atlantic seaboard.
+For this reason, the proposed clustering patterns were not used in this work.
+
+The primary baseline of this work is \cite{major-hurricane-model}, which uses a specific implementation of AprioriAll (first proposed in \cite{AprioriAll-original}) to determine frequent hurricane trajectory patterns.
+These trajectory patterns correspond to interesting trajectory association rules, which are used to train a prediction model.
+The implementation of AprioriAll used in \cite{major-hurricane-model} is also used as the foundational model for this work.
+AprioriAll transforms a transaction database into a \textit{customer} database, where each entry contains a customer and an ordered sequence of purchased items.
+The ordered sequence database is then mined for frequent ordered patterns, or frequent sequences.
+The modification of Apriori proposed in \cite{major-hurricane-model} mines frequent ordered patterns and thus skips the transformation step of AprioriAll.
+The model proposed in this work uses historical hurricane trajectory readings from HURDAT, the canonical hurricane database maintained by the National Hurricane Center \cite{HURDAT2-original}.
+The data in HURDAT2 is ordered into $\langle Hurricane, Trajectory \rangle$ pairs, omitting the need for the sequence mapping step required at initialization of AprioriAll.
+
+% We might want some more related work
+
+\section{Motivation}
+
+There are two categories of motivation behind this work.
+The first motivation of this work is to increase the total investigation completed in the hurricane trajectory prediction subdiscipline.
+Little research effort has been allocated to this particular problem space of high economic, and social impact.
+The authors expect this work to constitute a growing effort by academic and government researchers to improve existing methods of hurricane trajectory prediction in the interest of public safety and economic stability.
+
+The second motivation behind this work is the specific goal of improving the correctness ratio in the current standard data mining model for hurricane trajectory prediction.
+As previously described, HTPDM had a best case trajectory prediction correctness ratio of 65\%.
+This means that of all testing trajectory for a which a matching rule was successfully produced, 65\% of such matches were correct predictions.
+When HTPDM accounts for testing trajectories for which no rule could be matched, the correctness ratio decreases by 7.5\% to a final value of 57.5\% correctness.
+This decrease in accuracy is expected at some hurricane trajectories are outliers which cannot be predicted by frequent pattern mining, as will be discussed further in this work.
+
+The motivation of improving the correctness ratio in the model encompasses the challenges of improving the best-case correctness ratio (the correctness ratio when ignoring rule-miss incidents) and the worst-case correctness ratio (the correctness ratio when accounting for rule-miss incidents).
+To resolve the motivations of this work, a series of improvement strategies are proposed as applications to the current hurricane trajectory prediction model based on AprioriAll.
+In general, these strategies seek to improve the realism of the prediction model, reduce historical factors which may negatively affect the model's correctness reatio, and favor recent hurricane trajectory trends which are likely to be reflected in current and future hurricane patterns.
+
+\section{Problem Description}
+
+% What are the things to know about the problem?
+% - Algorithm: AprioriAll ( pseudocode )
+% - hit rate/miss rate
+% - regions ( show diagram )
+% - define accuracy, rule fit, fitness function.. ( equations )
+
+In this section, the problem of achieving accurate hurricane trajectory prediction is explained in detail.
+This problem consists of multiple challenges which apply generally to sequential datamining, and specifically to the problem domain of hurricane location analysis.
+Many of the domain-specific challenges of hurricane trajectory prediction arise from the fact that hurricane points are tracked as a series of $\langle Latitude, Longtitude \rangle$ points, which do \textit{not} directly correspond to traditional (x, y) coordinates, as latitude and longtitude are points imposed upon a sphere.
+Challenges including accurate distance evaluation are results of the transition from a Euclidean to a Polar coordinate system.
+
+The exposition of the hurricane prediction problem domain is divided into three parts.
+The first aspect of the problem domain concerns the underlying sequential mining approach used.
+This includes a thorough description of AprioriAll, its intended usage, and the reaction of AprioriAll to its input parameters.
+Next, the subproblem of region discretization will be discussed.
+Since $\langle Latitude, Longtitude \rangle$ points constitute continuous space, hurricane trajectories are converted to sequences of discrete regions to enable the discovery of frequent patterns.
+Finally, problems involving hurricane trajectory matching will be discussed.
+These problems include pattern matching for determining best fit rules and pattern matching for determining trajectory prediction correctness.
+
+\subsection{AprioriAll}
+
+In HTPDM and WARD-HTP, the algorithm that generates trajectory prediction possibilities is based on AprioriAll.
+AprioriAll was first proposed at IBM research as a method for discovering frequent sequential patterns in transactional databases \cite{AprioriAll-original}.
+In its original implementation, AprioriAll is able to generate assocation rules describing sentences of the form ``Customers who buy P may also buy Q''.
+In the hurricane trajectory prediction problem, AprioriAll is used to generate association ruls describing sentences of the form ``Hurricanes located at P may later be located at Q''.
+In the context of AprioriAll, hurricanes are considered $customers$ and their sequences of recorded points are considered $transactions$.
+Consider the high-level implementation steps of AprioriAll as follows:
+
+\begin{enumerate}
+\item Map the transaction database into a customer database.
+\item Generate frequent sequences based on minimum support.
+\item Generate association rules based on minimum confidence.
+\end{enumerate}
+
+\subsubsection{Mapping}
+
+In step (1) of AprioriAll, a given transaction database \textit{D} is grouped by customers, and the frequent items purchased by customers in individual transcations are recorded in a temporary table.
+The contents of this temporary table are mapped to corresponding sequences to enable the mining of frequent transaction sequences from the original database.
+The intention of step (1) is to map \textit{D} in its original $\langle TransactionID, CustomerID, Items\rangle$ database form into a database \textit{D'} of the form $\langle CustomerID, ItemSequence\rangle$.
+
+The hurricane trajectory model in this work does not implement step (1), as the training data from HURDAT2 was instead translated into a $\langle HurricaneID, CoordinateSequence\rangle$ database as a preprocessing step before any aspects of the model are executed.
+By directly translating the data instead of processing it with AprioriAll, all trajectories in the original database are preserved.
+This includes coordinates which may ultimately be infrequent.
+The sequential mapping of AprioriAll is performed as a preprocessing step so that the database used in the prediction model will always consist of hurricanes and their coordinate sequences.
+This allows the prepared (mapped) database to be mined directly without the need to complete the mapping step every time AprioriAll is executed.
+
+\begin{figure}[h]
+\caption{Sample of mapped Hurricane-Coordinates database}
+\centering
+\includegraphics[scale=1.0]{hurricane-table-sample}
+\end{figure}
+
+The above table shows an example of how the trajectory database is structured after the mapping preprocessing step.
+Only two dimensions are needed in the database, being the hurricane ID and the sequence of latitude-longtitude coordinates.
+All hurricanes in HURDAT2 have an ID corresponding to the location (the Atlantic), the index ordered by appearance in the hurricane year, and the year itself.
+For example, hurricane Katrina of 2005 corresponds to AL122005.
+The preprocessing step proposed is executed exactly once, allowing the hurricane database to be mined without the initial grouping and mapping stage.
+
+\subsubsection{Pattern Generation}
+
+\subsubsection{Rule Generation}
+
+\subsection{Region Discretization}
+
+% Put some visualization here and cite matplotlib
+
+\subsection{Trajectory Matching}
+
+\section{Solution}
+
+\subsection{Haversine Formula}
+
+\subsection{Weighted Training Data}
+
+\subsection{Training Data Year Interval}
 
 \section{Experimental Setup}
-
-% Need CSV output for metrics
 
@@
 \begin{thebibliography}{1}
-\bibitem{} 
-\bibitem{} 
-\bibitem{} 
-\bibitem{} 
-\bibitem{} 
-\bibitem{} 
-\bibitem{} 
-\bibitem{} 
+\bibitem{def-hurricane}  
+\bibitem{hurricane-cat} 
+\bibitem{hurricane-cost} 
+\bibitem{hurricane-cost-non-financial} 
+\bibitem{datamining-def} 
+\bibitem{major-hurricane-model} 
+\bibitem{AprioriAll-original} 
+\bibitem{hurricane-intensity-1}
+\bibitem{hurricane-intensity-2}
+\bibitem{hurricane-forecasting-1}
+\bibitem{hurricane-forecasting-2}
+\bibitem{typhoon-clustering}
+%\bibitem{matplotlib}
+
+\bibitem{HURDAT2-original} 
 \bibitem{} 
@@
 \end{document}
+
+% there will be lots of editing %

### 2017-12-03 apriori_test.py

@@ -23,7 +23,7 @@ result: x*10 + y = region #
 
 
 """
-MIN_SUP = 30
+MIN_SUP = 50
 
 # region and safe_incr code moved to util file
 

### 2017-12-03 tuple_regions/apriori_test.py

@@ -4,29 +4,10 @@ import sys
 import itertools
 import time
 
-# Need to dedupe raw vectors
-
-# For regionalization, need to represent vectors as a scalar and modulo by their bucket
-# How to make vector into a scalar? Norm of vector? Eigenvalue?
-
-# Some of the data entries are impossible (latitude cannot exceed 180. BTW the negative indicates west)
-
-# From some analysis, height is from 7.2->81.0, width from -109.5->63. 73.8 H-H, 172.5 W-W
-
-# As a first pass on regionalization, consider the following:
-""" Suppose we want to group the map into a 10x10 grid
-The question becomes "what sector does my x,y belong to"
-answer:
-73.8 / 10 -> 7.38. So take candidate x / 7.38, truncate to int, that's the x-region. If on the line (modulo = 0), take -1
-ditto for y.
-result: x*10 + y = region #
-
-# tuple regions are started but there is more work to do. First make sure multi-tuple-arrays are working correctly
-"""
 WEIGHT_INTERVAL_NOMINAL = 756
 WEIGHT_INTERVAL_YEAR = 50
 HURRICANE_ID_PLACEHOLDER = "AL015555"
-WEIGHT_SCALE = 7.5
+WEIGHT_SCALE = 2.5
 # For weighted data:
 # Data can be weighted nominally or year-over-year
 # Careful! Right now these values are dependent on the specific training data input. This must be fixed
@@ -91,9 +72,9 @@ def check_weight_update(weightedNominal, w, currId, iterId):
   newId = currId
   newW = w
   if weightedNominal:
-    newW = min(w + (WEIGHT_SCALE / WEIGHT_INTERVAL_NOMINAL), 1)
+    newW = min(w + (WEIGHT_SCALE / WEIGHT_INTERVAL_NOMINAL), 2.5)
   elif currId[-4:] != iterId[-4:]:
-    newW = min(w + (WEIGHT_SCALE / WEIGHT_INTERVAL_YEAR), 1)
+    newW = min(w + (WEIGHT_SCALE / WEIGHT_INTERVAL_YEAR), 2.5)
     newId = iterId
 
   return newW, newId
@@ -130,7 +111,8 @@ def execute_apriori():
         for region in regions:
           safe_increment(freqs, region, w)
 
-        w, currId = check_weight_update(weightedNominal, w, currId, _id)
+        if weightedData:
+          w, currId = check_weight_update(weightedNominal, w, currId, _id)
 
[…]

### 2017-12-03 tuple_regions/basic_trajectory_draw.py

@@ -20,6 +20,7 @@ def drawTrajectory():
   with open(sys.argv[1], 'r') as inputFile:
     i = 0
     for line in inputFile.readlines():
+      color = 'r'
       if i % fraction == 0: # Render 1 in fraction trajectories (nonzero)
         lineVector = line.split(' ')[1].split(",")[:-1]
 
@@ -32,8 +33,27 @@ def drawTrajectory():
           gisXCoords.append(int(newCoord[0]))
           gisYCoords.append(int(newCoord[1]))
 
-        m.plot(gisXCoords, gisYCoords, linewidth=0.5, color='r')
+        m.plot(gisXCoords, gisYCoords, linewidth=1.0, color=color)
       i += 1
+
+
+  # with open(sys.argv[4], 'r') as otherInput:
+  #   i = 0
+  #   for line in otherInput.readlines():
+  #     if True or i % fraction == 0: # Render 1 in fraction trajectories (nonzero)
+  #       lineVector = line.split(' ')[1].split(",")[:-1]
+
+  #       parsedVector = [item.replace("(", "").replace(")", "") for item in lineVector]
+  #       trajectoryVector = [ (float(item.split(":")[0]), float(item.split(":")[1])) for item in parsedVector]
+  #       gisXCoords = []
+  #       gisYCoords = []
+  #       for coord in trajectoryVector:
+  #         newCoord = m(coord[1], coord[0]) # Flip the input coordinates only on conversion
+  #         gisXCoords.append(int(newCoord[0]))
+  #         gisYCoords.append(int(newCoord[1]))
+
+  #       m.plot(gisXCoords, gisYCoords, linewidth=1.0, color='b')
+  #     i += 1
 
   m.drawcoastlines()
   m.drawcountries()
@@ -52,7 +72,7 @@ def drawTrajectory():
   plt.show()
 
 def main():
-  if len(sys.argv) != 4:
+  if False and len(sys.argv) != 4:
     print "Usage: {0} <VectorsInput> <FilterFraction> <RegionOverlaySwitch>".format(sys.argv[0])
   else:
     drawTrajectory()

### 2017-12-03 tuple_regions/generate_assoc_rules.py

@@ -7,8 +7,6 @@ import sys
 import time
 import re
 
-MIN_CONF = 0.25
-
 def getRawRules(itemset):
   divider = len(itemset) - 1
   rules = [] # Antecedent, Consequent pairs
@@ -38,7 +36,7 @@ def getAntecedentSupport(itemsetsFile, searchLine):
 
   return result
 
-def generateRules(itemsetsFile, currItemsetLine): # String line of form (t, t, t) : s
+def generateRules(itemsetsFile, currItemsetLine, minconf=0.25): # String line of form (t, t, t) : s
   generatedRules = []
 
   currItemsetLine = currItemsetLine.replace(" ", "")
@@ -61,7 +59,7 @@ def generateRules(itemsetsFile, currItemsetLine): # String line of form (t, t, t
     antecedentSupport = getAntecedentSupport(itemsetsFile, searchLine)
 
     conf = float(currItemsetSupport) / float(antecedentSupport)
-    if conf >= MIN_CONF:
+    if conf >= minconf:
       generatedRules.append( (rule[0], rule[1], conf) )
 
   return generatedRules
@@ -83,11 +81,12 @@ def convertToTupleList(line):
 
 def createRulesFile():
   totalRules = 0
+  minconf = float(sys.argv[3])
   with open(sys.argv[1], 'r') as itemsets, open(sys.argv[2], 'w') as rules:
     for line in itemsets:
 
       if line[1] == "(": # Multi tuple pattern
-        currRules = generateRules(sys.argv[1], line)
+        currRules = generateRules(sys.argv[1], line, minconf)
         totalRules += len(currRules)
         for rule in currRules:
           rules.write(
@@ -102,8 +101,8 @@ def createRulesFile():
   print "Total rules generated:", totalRules
 
 def main():
-  if len(sys.argv) != 3:
-    print "Usage: {0} <PatternsInput> <RulesOutput>".format(sys.argv[0])
+  if len(sys.argv) != 4:
+    print "Usage: {0} <PatternsInput> <RulesOutput> <MinConf>".format(sys.argv[0])
   else:
     createRulesFile()
 

### 2017-12-03 tuple_regions/test_best_fit.py

@@ -75,6 +75,7 @@ def strToTupleArray(arrayString):
 def getBestRule(rules, checkRegions, flex=False):
   bestRuleScore = 0
   bestRuleConsequent = []
+  bestRule = []
   for rule in rules:
     antecedent, consequent, confidence = getAttributes(rule)
     if flex:
@@ -85,10 +86,11 @@ def getBestRule(rules, checkRegions, flex=False):
     if matchingLen > 0:
       ruleScore = confidence * (1 - math.exp( matchingLen*-1 ) )
       if ruleScore > bestRuleScore:
+        bestRule = antecedent
         bestRuleScore = ruleScore
         bestRuleConsequent = consequent
   rules.seek(0)
-  return bestRuleScore, bestRuleConsequent
+  return bestRule, bestRuleScore, bestRuleConsequent
 
 def getLatLongDistance(startCoord, endCoord, haversine=True):
   if haversine:
@@ -122,7 +124,7 @@ def getFinalBearing(startCoord, endCoord):
   # lon, W/E, x, 1 (negative)
 
   # Determine the direction of the trajectory at termination (for extrapolation and for similar-region comparison)
-  lon1, lat1, lon2, lat2 = map(math.radians, [startCoord[1], startCoord[0], endCoord[1], endCoord[0]])
+  lon1, lat1, lon2, lat2 = map(math.radians, [float(startCoord[1]), float(startCoord[0]), float(endCoord[1]), float(endCoord[0])])
 
   distLon = lon2 - lon1
 
@@ -208,7 +210,7 @@ def getCoordCriteriaMatch(currCoord, currCoordList, compareCoordList, currCoordI
 def coordsMatchCriteria(leftCoord, rightCoord, leftCoordList, rightCoordList, leftIndex, rightIndex):
   # See if the two coordinates are within threshold distance and bearing difference
   # These constants are subject to change wrt experiments
-  maxDistanceDiff = 300
+  maxDistanceDiff = 450
   maxBearingDiff = 10
   match = False
   # First, check distance comparison
@@ -327,7 +329,7 @@ def testRules():
       searchRounds = 0
       complete = False
       while not complete:
-        bestRuleScore, bestRuleConsequent = getBestRule(rules, checkRegions, flex=False)
+        bestRule, bestRuleScore, bestRuleConsequent = getBestRule(rules, checkRegions, flex=False)
         if bestRuleScore == 0 and ignoreExtrapolation:
           complete = True
[…]

### 2017-12-03 paper/manuscript.tex
add regionalization to paper tex

@@
 When HTPDM accounts for testing trajectories for which no rule could be matched, the correctness ratio decreases by 7.5\% to a final value of 57.5\% correctness.
-This decrease in accuracy is expected at some hurricane trajectories are outliers which cannot be predicted by frequent pattern mining, as will be discussed further in this work.
+This decrease in accuracy is expected as some hurricane trajectories are outliers which cannot be predicted by frequent pattern mining, as will be discussed further in this work.
 
@@
 
-% Put some visualization here and cite matplotlib
+Atlantic hurricane trajectories cover a very large span of distance, in some cases travelling more than 10,000 kilometers from first to last advisory \cite{hurricane-distances}.
+As such, the granularity of such trajectories is too small to effectively derive any meaningful patterns \cite{major-hurricane-model}.
+Thus, this work chooses to transform the hurricane trajectories into coarser paths, sacrificing some precision to generate more frequent trajectory patterns over a more general area \cite{ransform-hurricanes-for-discretization}.
+
+In HTPDM and WARD-HTP, hurricane trajectory coordinates are scaled by a constant value of 5 and cast to integer values.
+The below figures show the difference of hurricane trajectory values before and after coordinate discretization.
+The python library matplotlib was used to generate images overlaying hurricane trajectories over a map of the Atlantic seaboard and will be used extensively throughout this paper \cite{matplotlib}.
+
+\begin{figure}[h]
+\caption{Hurricane Trajectory Paths}
+\centering
+\includegraphics[scale=1.0]{hurricane-trajectory-raw-diagram}
+\end{figure}
+
+\begin{figure}[h]
+\caption{Hurricane Trajectory Paths With Discretized Coordinates}
+\centering
+\includegraphics[scale=1.0]{hurricane-trajectory-discretized-diagram}
+\end{figure}
+
+As can be observed in Fig.
+3, coordinate discretization clamps hurricane trajectory points into larger discrete regions.
+As previously described, hurricane coordinates are tracked by latitude and longitude pairs, represented in degrees from the equator and the prime meridian, respectively.
+One degree of latitude is roughly equivalent to 111 km, however the distance of one degree of longitude is variable \cite{lat-long-distance}.
+This results in hurricane coordinates being roughly bounded in discrete regions of 5 degrees, or 555 km.
 
@@
 \bibitem{typhoon-clustering}
-%\bibitem{matplotlib}
+\bibitem{hurricane-distances}
+\bibitem{transform-hurricanes-for-discretization}
+\bibitem{matplotlib}
+\bibitem{lat-long-distance}
 

### 2017-12-05 paper/manuscript.tex
Write paper up to and including problem domain exposition

--- 2017-12-03 paper/manuscript.tex
@@
 While meteorological methods of hurricane trajectory prediction have been effective in the past, little attention has been paid to hurricane trajectory prediction with respect to historical frequent pattern mining.
-In this work, we propose multiple dimensions for improving the accuracy of datamining-based hurricane trajectory prediction.
+In this work, we propose multiple dimensions for improving the accuracy of data mining-based hurricane trajectory prediction.
 We propose the use of Haversine distance, a real-world distance formula for improved hurricane pattern matching accuracy.
@@
 Finally, we use training data exclusively from the years 1950 to 2000 to reduce the impact of historical inaccuracies in hurricane recording methodologies.
-After implementing all three improvement dimensions on a contemporary data mining model for hurricane trajectory prediction based on AprioriAll, results show a prediction-correctness rate 10.22\% higher than the current state-of-the-art, at 75.22\%.
+After implementing all three improvement dimensions on a contemporary data mining model for hurricane trajectory prediction based on AprioriAll, results show a prediction-correctness ratio 10.22\% higher than the current state-of-the-art, at 75.22\%.
 
@@
 This paper focuses exclusively on the subfield of hurricane trajectory prediction.
-In particular, the goal of this paper is to investigate and improve on the existing-state-of-the art in hurricane trajectory prediction systems based on sequential datamining.
+In particular, the goal of this paper is to investigate and improve on the existing-state-of-the art in hurricane trajectory prediction systems based on sequential data mining.
 
-Data mining is the nontrivial extraction of implicit, previously unknown, and potentially useful information from data \cite{datamining-def}.
+Data mining is the nontrivial extraction of implicit, previously unknown, and potentially useful information from data \cite{data mining-def}.
 Data mining applies to an extremely wide range of applications, including clickstream mining, social network community detection, and artificial intelligence.
-The datamining subdiscipline of particular interest in this work is \textit{sequential} data mining.
+The data mining subdiscipline of particular interest in this work is \textit{sequential} data mining.
 Sequential data mining uses a database of chronologically ordered inputs to mine frequent patterns and interesting association rules.
@@
 To the knowledge of the authors, the existing standard in hurricane trajectory prediction via data mining is developed by Dong et al in \cite{major-hurricane-model}.
-The work cited proposes a hurricane trajectory prediction model based on datamining (HTPDM) that uses a modified variant of AprioriAll, a common sequential mining algorithm \cite{AprioriAll-original}.
+The work cited proposes a hurricane trajectory prediction model based on data mining (HTPDM) that uses a modified variant of AprioriAll, a common sequential mining algorithm \cite{AprioriAll-original}.
 HTPDM executes AprioriAll against historical hurricane trajectories from 1900 to 2000 to generate interesting trajectory association rules.
@@
 The modification of Apriori proposed in \cite{major-hurricane-model} mines frequent ordered patterns and thus skips the transformation step of AprioriAll.
+
 The model proposed in this work uses historical hurricane trajectory readings from HURDAT, the canonical hurricane database maintained by the National Hurricane Center \cite{HURDAT2-original}.
 The data in HURDAT2 is ordered into $\langle Hurricane, Trajectory \rangle$ pairs, omitting the need for the sequence mapping step required at initialization of AprioriAll.
+The Atlantic hurricane trajectories recorded in HURDAT2 span a very large distance, some of which reaching over 10,000 kilometers in total distance travelled.
+As a result, the coordinates recorded are limited to those within a range of 0 to 50 degrees north, and 20 to 100 degrees west.
+This degree range is approximately the size of the south Atlantic basin.
 
@@
 In this section, the problem of achieving accurate hurricane trajectory prediction is explained in detail.
-This problem consists of multiple challenges which apply generally to sequential datamining, and specifically to the problem domain of hurricane location analysis.
+This problem consists of multiple challenges which apply generally to sequential data mining, and specifically to the problem domain of hurricane location analysis.
 Many of the domain-specific challenges of hurricane trajectory prediction arise from the fact that hurricane points are tracked as a series of $\langle Latitude, Longtitude \rangle$ points, which do \textit{not} directly correspond to traditional (x, y) coordinates, as latitude and longtitude are points imposed upon a sphere.
@@
 
-\subsubsection{Pattern Generation}
+\subsubsection{Sequence Generation}
+
+Section (2) of AprioriAll initiates the generation of frequent sequences.
+In the case of hurricane trajectory prediction, these sequence correspond to frequent recorded hurricane coordinate sequences.
+Frequent hurricane coordinate sequences are those which occur at least a minimum number of times, as defined by the \textit{minimum support}, or minsup.
+In general terms, the sequence generation step of AprioriAll develops frequent ordered patterns of increasing length, beginning with frequent single-item sequences (corresponding to frequent single-point sequences).
+This step begins by scanning the $\langle HurricaneID, CoordinateSequence\rangle$ database for frequent single-point sequences.
+The frequent single-point trajectories are then converted to candidate two-point trajectories as the result of an inner join.
+The candidate two-point trajectories are scanned against the database to determine which candidates satisfy the minimum support.
+Another inner join generates the candidate 3-length trajectories.
+Once the sequence generation level reaches three, each candidate sequence must be pruned if any of their candidate subsequences are infrequent.
+Continuous execution of AprioriAll until completion results in a collection of frequent hurricane trajectory patterns, as defined by the minumum support.
+
+%AprioriAll Pseudocode
+\begin{algorithm}[H]
+  \caption{Modified AprioriAll: No Litemset phase}
+  \label{aprioriall_for_hurricanes}
+  \begin{algorithmic}[1]
+  \State{$L_{1} \gets \{Frequent\ Single\ Points\}$}
+  \For{($k=2; L_{k-1} \neq \emptyset k++)$}
+    \State{$C_{k} \gets Candidates\ from\ Inner\ Join$}
+    \State{$Prune\ C_{k}\ of\ infrequent\ subsequences\ $}
+    \For{$sequence\ S \in HurricaneDB$}
+      \State{$Increase\ c \in C_{k}\ support\ if\ c \in S$}
+    \EndFor
+    \State{$L_{k} \gets \{c \in C_{k}|\ sup(c) \geq minsup\}$}
+  \EndFor
+  \State{$Frequent Trajectories \gets All(L_{k})$}
+  \end{algorithmic}
+\end{algorithm}
+
+Along with HTPDM, WARD-HTP executes AprioriAll with a slight modification to the existing algorithm proposed in \cite{AprioriAll-original}.
+Whereas L1, the frequent single-item sequences is generated by the L-Itemset (mapping) phase of AprioriAll, the variant of AprioriAll used in hurricane trajectory prediction skips the L-Itemset phase, and thus scans the $\langle HurricaneID, CoordinateSequence\rangle$ directly for frequent single-point trajectories.
+Otherwise, the remainder of AprioriAll is used in this work as originally designed.
+AprioriAll is used to ensure that the order of recorded hurricane coordinate sequences is preserved.
+This is a product of the inner-join step.
+Consider two hurricane trajectories $T_{1}$ and $T_{2}$.
+$T_{1}$ and $T_{2}$ form a new trajectory $T_{3}$ if and only if the \textit{L}-1 terminal coordinates of $T_{1}$ correspond to the \textit{L}-1 initial coordinates of $T_{2}$, where \textit{L} is the current level of AprioriAll being executed.
+For example, if \textit{L}=2, then AprioriAll scans for frequent 2-coordinate sequences, and uses those frequent 2-coordinate sequences to generate 3-coordinate sequences based on trajectory matching.
+If the matching condition is satisfied, a new trajectory is generated.
+The inner-join step of AprioriAll emulates the required hurricane trajectory-matching to develop frequent hurricane trajectory patterns.
+Once all possible frequent hurricane trajectory patterns are discovered, the second step of AprioriAll completes.
 
@@
 
+The final step of AprioriAll is the rule-generation step.
+Though association rule generation is not a specific stage of the AprioriAll algorithm, it is traditionally the final stage of any association rule mining approach.
+The rule generation step is responsible for developing interesting association rules.
+As mentioned, rules generated by the hurricane trajectory prediction model proposed will satisfy sentences of the form ``Hurricanes located at P may later be located at Q''.
+
+Association rules are derived directly from frequent patterns generated by AprioriAll or any modified versions.
+These rules consist of four dimensions: the antecedent, consequent, support and confidence.
+An association rule \textit{R} is generated by partitioning a frequent sequence \textit{S} into two congruent parts.
+The left partition of \textit{S} constitutes the antecedent of \textit{R}, and the latter partition is the consequent.
+The support of \textit{R} is the original support of \textit{S}.
+The confidence of \textit{R} is the support of \textit{S} divided by the support of the subsequence given by the former partition.
+
+Rules with confidence of at least the \textit{minimum confidence} (minconf) are stored in a table of interesting association rules.
+The rules generated by AprioriAll are the candidates for hurricane trajectory pattern matching.
+Hurricane trajectories are matched against all association rule antecedents.
+The consequent of the closest match rule is then used to predict the hurricane terminal trajectory.
+If the association rule consequent matches the actual hurricane terminal trajectory, the prediction is considered correct.
+Without the association rules formed by AprioriAll and the rule generation step, data mining-based hurricane trajectory prediction cannot be executed.
+HTPDM and WARD-HTP both use historically frequent hurricane trajectory to predict the result of future hurricane paths.
+
 \subsection{Region Discretization}
 
-% Put some visualization here and cite matplotlib
+This subsection covers problems related to the discretization of $\langle Latitude, Longtitude \rangle$ points.
+As shown in figure (1), the raw recorded hurricane coordinates are expressed in degrees and minutes.
+The excessively fine-grained nature of the raw data prevents the data from being able to produce meaningful frequent coordinate sequences.
+Therefore, effort must be taken to discretize the coordinate system, allowing more coarse-grained frequent patterns to be produced.
+
+In HTPDM and WARD-HTP, recorded hurricane coordinates are discretized into approximately square blocks of a specific size, expressed in degrees.
+Consider the first row of coordinates shown in figure (1): (12.5:-37.5), (13.0:-39.0), (13.5:-40.5).
+Mapping this coordinate sequence to a discritization block of size 5, the new trajectory becomes (2:-7), (2:-7), (2:-8).
+Duplicate coordinates are not counted, so the trajectory ultimately reduces to (2:-7), (2:-8).
+Note that the degree-size of the discretization block is expressed as an integer, which means that all coordinates are mapped to their lowest divisor relative to the discretization size.
+This enables the coarse-grained mapping of coordinates to their corresponding region block.
+
+% Figures 2 and 3: Trajectories without and with region overlay
+\begin{figure}[htp]
+\caption{Sample Hurricane Trajectory Paths}
+\centering
+\includegraphics[scale=0.55]{Merged-Trajectory-Comparison}
+\end{figure}
+
+Consider figures (2) and (3) shown above, generated using a sample of the recrded hurricane trajectories in HURDAT2 with the visualization tool Matplotlib \cite{matplotlib}.
+As shown in the figures, region discretization organizes continuous hurricane coordinate sequences into discrete region sequences of approximately uniform size and square shape.
+A discretization block size of 5 is selected in HTPDM and is replicated in WARD-HTP.
+A block size of 5 corresponds approximately to a real-world block dimension of 500km by 500km.
+The block dimension size is approximate because the mapping between coordinate degrees and kilometers is imprecise, and is dependent on location in the case of longtitudal points \cite{lat-long-distance}.
+The approximate size of 500km is similar to the average hurricane total diameter \cite{hurricane-distances}, permitting the coarse-grain expression of hurricane trajectories without undue loss in precision.
+
+All aspects of hurricane trajectory prediction with data mining depend on region discretization.
+Frequent trajectory patterns and their resulting trajectory predictions are expressed in terms of the regions where a given hurricane is expected to appear.
+The process of region discretization allows for hurricane coordinate trajectories to be scaled down, allowing the regionalized coordinates to be treated as a simplified coordinate system in subsequent aspects of the prediction model.
 
 \subsection{Trajectory Matching}
+
+The final key aspect of the hurricane trajectory prediction problem concerns accurate hurricane trajectory pattern matching.
+Pattern matching is a concern for hurricane trajectory prediction in two primary areas.
+First, hurricane trajectory matching must be effectively employed when comparing a new hurricane trajectory against the table of hurricane association rules generated by the previously described variant of AprioriAll.
+Hurricane trajectory matching is also used when checking the correctness of a trajectory prediction.
+The consequent of the matched hurricane trajectory rule is compared with the actual hurricane terminal trajectory to determine prediction correctness.
+
+\subsubsection{Rule-Fit Matching}
+
+Pattern matching is used when evaluating the best-fit rule for hurricane prediction purposes.
+Similar to HTPDM, a new hurricane trajectory is passed into the hurricane prediction model proposed in this work, it is separated into two subtrajectories: the initial trajectory $T_{i}$ and the terminal trajectory $T_{t}$.
+$T_{i}$ is compared against the antecedent of every interesting assocation rule satisfying minconf for a match.
+There are two dimensions that determine the best-fit rule for an incoming trajectory requiring prediction.
+These dimensions are matching length and rule confidence.
+The matching length of the initial trajectory $T_{i}$ to a rule antecedent is the longest congruent subsequence of regionalized points within the rule antecedent within $T_{i}$.
+Note that both $T_{i}$ and the rule antecedent sequences have been regionalized per the regionalization methodology outlined.
+Using the fitness function described in HTPDM \cite{major-hurricane-model}, the best-fit rule match for the incoming initial trajectory $T_{i}$ is expressed with the following formula:
+\begin{eqnarray*}
+Match = max( conf_{r} * (1 - e^{-matchLen_{t,r} } )\ ), r \in R
+\end{eqnarray*}
+Where r is a member of the set of all interesting assocation rules \textit{R}, t is the initial trajectory $T_{i}$, and $matchLen_{t,r}$ expresses the matching length between $T_{i}$ and the antecedent sequence of r.
+This formula is meant to be maximized to determine the rule which most generously fits the incoming hurricane trajectory.
+Once a best fit rule is determined, the rule consequent is compared with the \textit{actual} terminal hurricane trajectory to determine if the prediction is correct.
+
+\subsubsection{Prediction Correctness Matching}
+
+Prediction Correctness checking is also powered by hurricane trajectory pattern matching.
+The consequent of the best-fit rule generated by the rule-fit matching stage constitutes the predicted terminal sequence for the incoming hurricane trajectory.
+The best-fit consequent is then compared against the actual terminal trajectory $T_{t}$ to determine the prediction correctness.
+In HTPDM and WARD-HTP, prediction correctness is considered a binary attribute: a prediction is either correct or incorrect.
+If the matched-rule consequent also matches the incoming hurricane terminal trajectory, the prediction is correct and the model's correctness ratio is increased accordingly.
+
+The pattern matching for rule consequents has some differences compared to antecedent pattern matching.
+First, prediction correctness uses matching length with a different definition.
+Whereas matching length for rule antecedents is the longest identical congruent subsequence of the initial trajectory, matching length for the rule consequent is the longest congurent subsequence of points within a specific distance, beginning at the start of the terminal sequence.
+This methodology is also based on HTPDM and is based on the following steps.
+In the algorithm given, $conseq_{r}$ corresponds to the matched rule's consequent sequence, and D corresponds to the minimum distance threshold necessary for prediction correctness.
+In HTPDM, the distance threshold corresponds to a Euclidean distance of 1.
+
+\begin{algorithm}[H]
+  \caption{Consequent Matching for Prediction Correctness}
+  \label{consequent_matching}
+  \begin{algorithmic}[1]
+  \State{$l \gets min(|conseq_{r}|, |T_{t}|)$}
+  \State{$Select\ first\ l\ points\ of\ conseq_{r}$}
+  \State{$Select\ first\ l\ points\ of\ T_{t}$}
+  \If{$All\ corresponding\ points\ within\ D$}
+    \State{$Prediction:\ CORRECT$}
+  \EndIf
+  \end{algorithmic}
+\end{algorithm}
+
+% this may need some editing
 
@@
 \bibitem{hurricane-cost-non-financial} 
-\bibitem{datamining-def} 
+\bibitem{data mining-def} 
 \bibitem{major-hurricane-model} 
+\bibitem{HURDAT2-original} 
 \bibitem{AprioriAll-original} 
@@
 \bibitem{typhoon-clustering}
-%\bibitem{matplotlib}
+\bibitem{matplotlib}
+\bibitem{lat-long-distance}
+\bibitem{hurricane-distances}
 
-\bibitem{HURDAT2-original} 
-\bibitem{} 
-\bibitem{} 
 \end{thebibliography}
@@
 % there will be lots of editing %
+
+% Early callouts:
+% Motivation could be more clear
+% We need to make sure we are explaining rule-miss incidents and our approach to them
+% We also need to make sure we explain what exactly the correctness ratio is and how it changes based on accounting for rule-miss
+% future work: non-binary prediction correctness?
+% More non-text assets?
+% We have done a good job describing the problem space but we need to explain what we problems __are__
+% ie: why is there a solution

### 2017-12-05 tuple_regions/basic_trajectory_draw.py

@@ -63,10 +63,10 @@ def drawTrajectory():
   m.drawmeridians(np.arange(-100,0,20),labels=[0,0,0,1])
 
   if showRegions:
-    div = 80/float(LONGITUDE_DIVISIONS)
-    m.drawmeridians(np.arange(-100, -20, div),dashes=[1,0],color='b')
-    div2 = 50/float(LATITUDE_DIVISIONS)
-    m.drawparallels(np.arange(0, 50, div2),dashes=[1,0],color='b')
+    div = 5
+    m.drawmeridians(np.arange(-140, -20, div),dashes=[1,0],color='b')
+    div2 = 5
+    m.drawparallels(np.arange(0, 70, div2),dashes=[1,0],color='b')
 
   plt.title('Atlantic Hurricane Tracks')
   plt.show()

### 2017-12-06 paper/manuscript.tex
little tex fix
Complete Problem Description and start Soluiton section

@@
 \usepackage{graphicx}
-\usepackage{textcomp}
+%\usepackage{textcomp}
 \graphicspath{ {images/} }
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \caption{Sample of mapped Hurricane-Coordinates database}
@@
 
-The above table shows an example of how the trajectory database is structured after the mapping preprocessing step.
+Figure (1) shows an example of how the trajectory database is structured after the mapping preprocessing step.
 Only two dimensions are needed in the database, being the hurricane ID and the sequence of latitude-longtitude coordinates.
@@
 Section (2) of AprioriAll initiates the generation of frequent sequences.
-In the case of hurricane trajectory prediction, these sequence correspond to frequent recorded hurricane coordinate sequences.
+In the case of hurricane trajectory prediction, these sequences correspond to frequent recorded hurricane coordinate sequences.
 Frequent hurricane coordinate sequences are those which occur at least a minimum number of times, as defined by the \textit{minimum support}, or minsup.
@@
 
-Consider figures (2) and (3) shown above, generated using a sample of the recrded hurricane trajectories in HURDAT2 with the visualization tool Matplotlib \cite{matplotlib}.
-As shown in the figures, region discretization organizes continuous hurricane coordinate sequences into discrete region sequences of approximately uniform size and square shape.
+Consider figure (2), generated using a sample of the recrded hurricane trajectories in HURDAT2 with the visualization tool Matplotlib \cite{matplotlib}.
+As shown in the pair of images, region discretization organizes continuous hurricane coordinate sequences into discrete region sequences of approximately uniform size and square shape.
 A discretization block size of 5 is selected in HTPDM and is replicated in WARD-HTP.
@@
 
-% this may need some editing
+The hurricane trajecory prediction model based on data mining as originally proposed in \cite{major-hurricane-model} implements all aspects of the problem domain described.
+Modified AprioriAll, downscaled region discretization and pattern matching with respect to association rule antecedents and consequents are respectively employed by HTPDM in order to develop an effective system for predicting future hurricane trajectories.
+Based on Atlantic hurricane trajectory data from 1900 to 2008, the model generates a best-case correctness ratio of 65\% and a worst-case ratio of 57.5\%.
+The training data used in the original model consists of trajectories from 1900 to 2000, with testing data supplied by trajectories from 2001 to 2008.
+At a high level, the model is developed and tested by following the steps below, using all implementation strategies discussed.
+
+% Model formal explanation could be better. Maybe do pseudocode
+
+\begin{enumerate}
+\item Generate interesting trajectory rules from AprioriAll
+\item Determine matching rule for incoming trajectory 
+\item Check if the matched rule's consequent points are sufficiently close
+\end{enumerate}
+
+The differentiation between the model's best and worst-case hit rates is dependent on the ability of an incoming test trajectory to be matched with a trajectory association rule.
+As described in the pattern matching stage, the initial subtrajectory of an incoming trajectory is compared against all interesting rules to determine a best-fit match.
+If no such match is found, HTPDM attempts to extrapolate a future point of the initial trajectory, using the extrapolated point as a new option for determing a rule antecedent match.
+When accounting for all tested trajectories, including those which were matched to a rule as a result of trajectory extrapolation, the correctness ratio was 57.5\%.
+Therefore, the best case correctness ratio of HTPDM is the ratio which ignores all test trajectories which did not match with any rules, resulting in a final value of 65\%.
+
+Despite establishing a convincing initial start point, the best-case correctness ratio of 65\% may be too low for HTPDM to be considered realiable in the real-world.
+This is problematic for those interested in hurricane trajectory prediction as a model with only 65\% accuracy may not be able to deliver crucial insight into hurricane trajectory prediction, particularly with respect to landfall location prediction.
+Both correctness ratios given must be improved in order for data mining to be considered a reliable approach to predicting the trajectories of future hurricanes.
 
@@
 
-\subsection{Haversine Formula}
+In order to improve the correctness ratio of datamining-based hurricane trajectory prediction, this work identifies and explores three opportunities for worst-case and best-case ratio improvement.
+The opportunities for improvement are the improvement of realism, reduction of historical impact, and favorability towards recent hurricane trends.
+Therefore, the opportunities for improvement are realized by three unique dimensions.
+These dimensions are the use of a distance formula designed for spherical coordinate systems, the Haversine Distance Formula, the use of weighted training data when executing the modified variant of AprioriAll, and the modification of the training data to only use historical trajectories from 1950 to 2000.
+The implementation of these three dimensions results in a new hurricane trajectory prediction model dubbed WARD-HTP (Weighted-Asset, Realistic-Distance Hurricane Trajectory Predictor).
+The combined contributions of WARD-HTP are expected to result in improvements to the best-case and worst-case correctness ratios.
+
+\subsection{Haversine Distance Formula}
 
@@
 % We also need to make sure we explain what exactly the correctness ratio is and how it changes based on accounting for rule-miss
+% Also need to explicitly call out what our minconf was and why
 % future work: non-binary prediction correctness?
@@
 % ie: why is there a solution
+% we will need to re-read the paper and keep an eye out for sections that suffer from scope creep. Problem description first comes to mind.

### 2017-12-06 paper/manuscript.tex
add haversine text from the doc, some cleanup

@@
 Only two dimensions are needed in the database, being the hurricane ID and the sequence of latitude-longtitude coordinates.
-All hurricanes in HURDAT2 have an ID corresponding to the location (the Atlantic), the index ordered by appearance in the hurricane year, and the year itself.
-For example, hurricane Katrina of 2005 corresponds to AL122005.
+All hurricanes in HURDAT2 have an ID corresponding to the location (i.e.
+the Atlantic), the index ordered by appearance in the hurricane year, and the year itself.
+For example, hurricane Katrina of year 2005 corresponds to AL122005.
 The preprocessing step proposed is executed exactly once, allowing the hurricane database to be mined without the initial grouping and mapping stage.
@@
 
-In order to improve the correctness ratio of datamining-based hurricane trajectory prediction, this work identifies and explores three opportunities for worst-case and best-case ratio improvement.
-The opportunities for improvement are the improvement of realism, reduction of historical impact, and favorability towards recent hurricane trends.
+In order to improve the correctness ratio of datamining-based hurricane trajectory prediction, this work identifies and explores three opportunities for worst-case and best-case ratio refinement.
+The opportunities identified are the introduction of more realism in calculation formulas, reduction of historical impact on data used, and favorability towards recent hurricane trends.
 Therefore, the opportunities for improvement are realized by three unique dimensions.
@@
 \subsection{Haversine Distance Formula}
+
+Accurate distance measurement is a critical task in hurricane trajectory prediction.
+Pattern matching strategies in the proposed hurricane trajectory prediction model depend on coherent distance measurements between real-world spatial points.
+Consider the following prediction-correctness evaluation method proposed by \cite{major-hurricane-model}.
+To determine whether a trajectory prediction is correct, the distance between the first p points of the predicted and actual trajectories is compared against a defined maximum, where p is the minimum length of the two trajectories.
+If all p pairs of points fit within the maximum distance, the trajectory prediction is considered correct.
+
+Two different distance measurement formulas are examined in this work for use in the evaluation method described.
+This work compares the Euclidean distance formula against the Haversine distance formula.
+The Euclidean distance formula (as used in \cite{major-hurricane-model}) is designed to determine the unit distance between two N dimensional points in a straight line.
+In the proposed model, units correspond to degrees of latitude and longitude.
+In contrast, the Haversine formula is designed to determine the kilometers between two latitude-longitude points over a large sphere (such as the earth).
+For this reason, the Haversine formula is anticipated to provide a more accurate definition of distance between recorded hurricane points.
+When used in this work, the Euclidean distance formula is evaluated with two dimensions, corresponding respectively to latitude and longitude.
+The two formulas, denoted D_eucl and D_havr are defined as follows:
+
+\begin{eqnarray*}
+D_eucl(p1, p2) = sqrt( (lat_p2 - lat_p1)^2 + (lon_p2 - lon_p1)^2 )
+\end{eqnarray*}
+
+% TODO: paste the full haversine formula here
+\begin{eqnarray*}
+D_havr(p1, p2) = 
+\end{eqnarray*}
+
+To compare the prediction accuracy between the two proposed distance formulas, a sample execution of the hurricane trajectory prediction model is executed with identical variables, except for the difference in distance evaluation.
+The distance formulas are interchanged in the prediction correctness checking step of the model, described previously in this section.
+When using the Euclidean formula, the distance threshold is 1.
+This corresponds to the distance threshold proposed in \cite{major-hurricane-model}.
+When the Haversine formula is employed, the maximal distance is varied from 300 to 450 km.
+This value corresponds to slightly less than approximately one average hurricane-diameter.
+Since the average hurricane diameter is between 333 and 670 km \cite{hurricane-distances}, two points are considered to be within the distance threshold if they are less than one average hurricane-diameter apart.
+This size was decided upon because of the coordinate discretization method employed in the work of \cite{major-hurricane-model}, which as discussed in Region Discretization section, scales all hurricane coordinates by a constant value of 5.
+Therefore, the euclidean distance threshold of 1 in the discretized coordinate system corresponds to a square region approximately equal to 500 km by 500 km.
+As such, the maximal distance ranges used for the Haversine formula are much less than used in the Euclidean formula.
+The expected outcome of this experiment is to receive more granular results using the Haversine formula compared to the Euclidean formula.
 
@@
 \section{Experimental Results}
+
+\subsection{Haversine Distance Formula}
+
+As can be observed below in Figure (), between the Euclidean and Haversine distance formulas, the Haversine formula with a distance threshold of 450 km delivers a increased correctness rate of 38.49\%, corresponding to 74.77\% compared to 36.28\% correctness when the Euclidean formula was employed.
+It is noted however that a distance threshold of 300 km (representing the lower bound of average hurricane sizes) results in an increased correctness rate of 19.91\% over the Euclidean formula, but remains a lower correctness rate compared to a Haversine distance of 450 km.
+The realistic nature of the Haversine formula in conjunction with its corresponding maximal distance threshold is able to provide a higher rate of accurate predictions compared to the Euclidean distance formula, which is not designed for non-euclidean coordinate systems such as coordinate systems over spheres.
+
+% TODO: make Excel charts based on data from the Doc, paste them here
+\begin{figure}[ht]
+\caption{Comparative Execution of Euclidean and Haversine Distance Formulas}
+\centering
+%\includegraphics[scale=1.0]{ <TABLE NAME HERE> }
+\end{figure}
+
+\subsection{Weighted Training Data}
+
+\subsection{Training Data Year Interval}
 

### 2017-12-07 project_final/run_experiment.sh
Reorganized project. Added simple bash script to run everything.

@@ -1,0 +1,6 @@
+#!/bin/sh
+python scripts/generate_frequent_patterns.py data/vectors_training_1950_2000.txt output/patterns.txt 37.75 1 1
+
+python scripts/generate_assoc_rules.py output/patterns.txt output/rules.txt 0.25
+
+python scripts/test_best_fit.py data/vectors_testing_2001_2015.txt output/rules.txt 1

### 2017-12-07 project_final/scripts/apriori_utils.py

@@ -1,0 +1,63 @@
+
+LATITUDE_DIVISIONS = 11
+LONGITUDE_DIVISIONS = 18
+
+LAT_MAX_DEGREES = 50.0
+LAT_MIN_DEGREES = 0.0
+LONG_MAX_DEGREES = 100.0
+LONG_MIN_DEGREES = 20.0
+
+DIMENSIONS_LATITUDE = 5
+DIMENSIONS_LONGTITUDE = 5
+
+def gatherCoordinateList(inputStr):
+  parseStr = inputStr.split(' ')[1]
+  if parseStr[-1] == ',':
+    parseStr = parseStr[:-1]
+
+  result = []
+  parseStr = parseStr.strip()
+
+  for vecStr in parseStr.split(','):
+    if len(vecStr):
+      currVec = vecStr.replace('(', '').replace(')', '').split(':')
+      result.append( (float(currVec[0]), float(currVec[1])) )
+
+  return result
+
+def regionalize(coordinate):
+
+  # lat, N/S, y, 0
+  # lon, W/E, x, 1 (negative)
+  # Latitude is counted as degrees N/S of equator => y coords
+  # Longitude is counted as degrees W/E of meridian => x coords
+
+  # may need region pruning here
+
+  return (int(coordinate[0]/DIMENSIONS_LATITUDE), int(coordinate[1]/DIMENSIONS_LONGTITUDE))
+
+def gatherDedupedRegionList(ruleLine):
+  coordinates = gatherCoordinateList(ruleLine)
+  regionList = [regionalize(coord) for coord in coordinates]
+
+  dedupedRegions = []
+  # TODO: change this once we have bounds checking for data
+  i = 0
+  while i < len(regionList) and regionList[i] < 0:
+    i += 1
+  dedupedRegions.append(regionList[i])
+  foundItem = regionList[i]
+
+  while i < len(regionList):
+    if regionList[i] != foundItem and regionList[i] >= 0:
+      foundItem = regionList[i]
+      dedupedRegions.append(regionList[i])
+    i += 1
+
+  return dedupedRegions
+
+def safe_increment(table, key, value=1.0):
+  try:
+    table[key] += value
+  except KeyError:
+    table[key] = value

### 2017-12-07 project_final/scripts/generate_frequent_patterns.py

@@ -1,0 +1,172 @@
+
+from apriori_utils import *
+import sys
+import itertools
+import time
+
+WEIGHT_INTERVAL_NOMINAL = 756
+WEIGHT_INTERVAL_YEAR = 50
+HURRICANE_ID_PLACEHOLDER = "AL015555"
+WEIGHT_SCALE = 2.5
+# For weighted data:
+# Data can be weighted nominally or year-over-year
+# Careful! Right now these values are dependent on the specific training data input. This must be fixed
+# A good minsupport value for weight training data appears to be 10. Minconf can stay the same
+
+def gather_new_frequent_items(resultTable, frequent_k_itemsets, minsup):
+  resultTable.append({})
+  for key in frequent_k_itemsets:
+    if frequent_k_itemsets[key] >= minsup:
+      resultTable[-1][key] = frequent_k_itemsets[key] 
+
+def timeseries_candidate_generation(frequent_k_itemsets, chunk_size):
+  # this is the candidate generation and pruning
+  # We have our L_k, use suffix-prefix to get C_k+1
+
+  frequent_k_itemsets_dict = { tuple(item):None for item in frequent_k_itemsets }
+  candidates = []
+  for curr_index, curr_item in enumerate(frequent_k_itemsets):
+
+    compare_index = 0
+    found = False
+    leftChunk = curr_item[(chunk_size*-1):]
+
+    while compare_index < len(frequent_k_itemsets) and not found:
+      found = leftChunk == frequent_k_itemsets[compare_index][:chunk_size]
+      if not found:
+        compare_index += 1
+
+    while compare_index < len(frequent_k_itemsets) and leftChunk == frequent_k_itemsets[compare_index][:chunk_size]:
+      new_candidate = join_entries(curr_item, frequent_k_itemsets[compare_index], chunk_size)
+      required_itemset_checks = itertools.combinations(new_candidate, chunk_size+1)
+
+      if all(item == curr_item or item == frequent_k_itemsets[compare_index] or tuple(item) in frequent_k_itemsets_dict for item in required_itemset_checks):
+        candidates.append(new_candidate)
+      compare_index += 1
+
+  return candidates
+
+def join_entries(left, right, chunk_size):
+  result = left + right[chunk_size:]
+  return result
+
+def gather_sublists(region_list):
+  result = []
+  for size in range(1, len(region_list)):
+    for index in range(len(region_list)):
+      result.append(region_list[index : index + size])
[…]

### 2017-12-07 project_final/scripts/test_best_fit.py

@@ -1,0 +1,399 @@
+# Given an incoming trajectory:
+# Cut it into two pieces: check, confirm
+# Check "check" against rule antecedents, get best fit (or show "missing")
+# if not missing, see if "confirm" was right
+
+# To do this, we will need to cut the patterns into two pieces. Training and Testing.
+# First try: quarters. 1335 training 445 testing
+
+# First try: divide test line in half. First half looks for antecedent
+
+# Parse a line into the antecedent, list of strings
+
+import math
+import sys
+import time
+import re
+from apriori_utils import *
+
+EXTRAPOLATION_LIMIT = 10
+EARTH_RADIUS = 6371
+MAX_MINIMUM_MATCH = 5 # Max comparison length from paper
+
+def getAttributes(ruleLine):
+  ruleLine = ruleLine.replace(" ", "")
+  confParsed = ruleLine.split(":")
+
+  confidence = float(confParsed[1])
+
+  ruleParsed = confParsed[0].split("=>")
+
+  antecedent = strToTupleArray(ruleParsed[0])
+  consequent = strToTupleArray(ruleParsed[1])
+
+  return antecedent, consequent, confidence
+
+def getMatchingLen(antecedent, checkRegions):
+  # We have antecedent A1 A2 ... An
+  # and regions R1 R1 ... Rm
+  # What is the longest sequence of A's present in R's?
+  # Note that the sequence has to be ordered too
+  maxLen = 0
+  i = 0
+  j = 0
+
+  while i < len(checkRegions):
+    currLen = 0
+    currItem = checkRegions[i]
+    if currItem in antecedent:
+      index = antecedent.index(currItem)
+      j = i
+
+      while j < len(checkRegions) and index < len(antecedent) and checkRegions[j] == antecedent[index]:
+        j += 1
+        index += 1
+        currLen += 1
+
+    if currLen > maxLen:
+      maxLen = currLen
+    i += 1
+
+  return maxLen
+
+def strToTupleArray(arrayString):
+  # May have to fix when we see tuples of size > 1
+  # some code repitition here but that's alright for now
+  tupleArray = []
+  result = arrayString.replace("'", "").replace("[", "").replace("]", "")
+  rawTuples = re.findall("\([0-9-,]+\)" , result)
+  for rawTuple in rawTuples:
+    leftT, rightT = rawTuple[1:len(rawTuple)-1].split(",")
[…]

### 2017-12-07 project_final/scripts/tools/basic_trajectory_draw.py

@@ -1,0 +1,83 @@
+"""
+A Hurricane Trajectory-Rendering script based on examples from matplotlib
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+from mpl_toolkits.basemap import Basemap
+from apriori_utils import *
+
+def drawTrajectory():
+  fraction = int(sys.argv[2])
+  showRegions = bool(int(sys.argv[3]))
+  # Lambert Conformal Conic map.
+  m = Basemap(llcrnrlon=-100.,llcrnrlat=0.,urcrnrlon=-20.,urcrnrlat=50.,
+              projection='lcc',lat_1=20.,lat_2=40.,lon_0=-60.,
+              resolution ='l',area_thresh=1000.)
+  # read shapefile.
+  # Need to write a longtitudional line as opposed to a euclidean horizontal line
+
+  with open(sys.argv[1], 'r') as inputFile:
+    i = 0
+    for line in inputFile.readlines():
+      color = 'r'
+      if i % fraction == 0: # Render 1 in fraction trajectories (nonzero)
+        lineVector = line.split(' ')[1].split(",")[:-1]
+
+        parsedVector = [item.replace("(", "").replace(")", "") for item in lineVector]
+        trajectoryVector = [ (float(item.split(":")[0]), float(item.split(":")[1])) for item in parsedVector]
+        gisXCoords = []
+        gisYCoords = []
+        for coord in trajectoryVector:
+          newCoord = m(coord[1], coord[0]) # Flip the input coordinates only on conversion
+          gisXCoords.append(int(newCoord[0]))
+          gisYCoords.append(int(newCoord[1]))
+
+        m.plot(gisXCoords, gisYCoords, linewidth=1.0, color=color)
+      i += 1
+
+
+  # with open(sys.argv[4], 'r') as otherInput:
+  #   i = 0
+  #   for line in otherInput.readlines():
+  #     if True or i % fraction == 0: # Render 1 in fraction trajectories (nonzero)
+  #       lineVector = line.split(' ')[1].split(",")[:-1]
+
+  #       parsedVector = [item.replace("(", "").replace(")", "") for item in lineVector]
+  #       trajectoryVector = [ (float(item.split(":")[0]), float(item.split(":")[1])) for item in parsedVector]
+  #       gisXCoords = []
+  #       gisYCoords = []
[…]

### 2017-12-07 project_final/scripts/tools/generate_coord_vectors.py

@@ -1,0 +1,36 @@
+# A simple script to parse atlantic_storms.csv into a plaintext list of HurricaneID -> Trajectory
+# Hurricanes with fewer than 3 trajectories are not included in the list, but are recorded as pruned
+import sys
+
+def parseHurricanes():
+  with open(sys.argv[1], 'r') as inputFile, open(sys.argv[2], 'w') as outputFile, open(sys.argv[3], 'w') as prunedFile:
+    currentId = ""
+    currentVector = []
+
+    for line in inputFile.readlines()[1:]:
+      lineId = line.split(",")[1]
+      if lineId != currentId: # New ID!
+        if len(currentVector) >= 3: # Did we build enough data for this hurricane to be interesting?
+          outputFile.write(currentId+" "),
+          for item in currentVector:
+            outputFile.write(item),
+          outputFile.write("\n")
+        elif len(currentId) > 0:
+          prunedFile.write(currentId+"\n")
+
+        currentVector = [] # Print or not, move onto next ID/coordinates
+        currentId = lineId
+
+      outStr = "(" + line.split(",")[6] + ":" + line.split(",")[7] + "),"
+      currentVector.append(outStr)
+
+def main():
+  if len(sys.argv) != 4:
+    print "Usage: {0} <StormsCsvInput> <VectorsOutput> <PrunedStormsOutput>".format(sys.argv[0])
+  else:
+    parseHurricanes()
+
+if __name__ == '__main__':
+  main()
+else:
+  print __name__

### 2017-12-07 project_final/scripts/tools/parse_paper_vectors.py

@@ -1,0 +1,39 @@
+import sys
+
+YEAR_INDEX = 0
+NUM_INDEX = 4
+
+LAT_INDEX = 6
+LON_INDEX = 7
+
+def parseHurricanes():
+  with open(sys.argv[1], 'r') as inputFile, open(sys.argv[2], 'w') as outputFile:
+    currentId = ""
+    currentVector = []
+
+    for line in inputFile.readlines()[1:]:
+      lineSplit = line.split(" ")
+      lineId = lineSplit[YEAR_INDEX] + lineSplit[NUM_INDEX]
+
+      if lineId != currentId: # New ID!
+        outputFile.write(currentId+" ")
+        for item in currentVector:
+          outputFile.write(item)
+        outputFile.write("\n")
+
+        currentVector = [] # Print or not, move onto next ID/coordinates
+        currentId = lineId
+
+      outStr = "(" + lineSplit[LAT_INDEX] + ":" + lineSplit[LON_INDEX] + "),"
+      currentVector.append(outStr)
+
+def main():
+  if len(sys.argv) != 3:
+    print "Usage: {0} <StormsCsvInput> <VectorsOutput>".format(sys.argv[0])
+  else:
+    parseHurricanes()
+
+if __name__ == '__main__':
+  main()
+else:
+  print __name__

### 2017-12-07 paper/manuscript.tex
Complete paper first draft

--- 2017-12-06 paper/manuscript.tex
@@
 \usepackage{graphicx}
-%\usepackage{textcomp}
+\usepackage{amsmath}
+
+\usepackage{stfloats}
+
+\DeclareMathOperator{\atantwo}{atan2}
 \graphicspath{ {images/} }
@@
 In this work, we propose multiple dimensions for improving the accuracy of data mining-based hurricane trajectory prediction.
-We propose the use of Haversine distance, a real-world distance formula for improved hurricane pattern matching accuracy.
+We propose the use of Haversine distance, a real-world distance formula for reliable hurricane pattern matching accuracy.
 We also propose the use of weighted data to favor recent hurricane trends when training the trajectory prediction model.
 Finally, we use training data exclusively from the years 1950 to 2000 to reduce the impact of historical inaccuracies in hurricane recording methodologies.
-After implementing all three improvement dimensions on a contemporary data mining model for hurricane trajectory prediction based on AprioriAll, results show a prediction-correctness ratio 10.22\% higher than the current state-of-the-art, at 75.22\%.
+After implementing all three improvement dimensions on a contemporary data mining model for hurricane trajectory prediction based on AprioriAll, results show a prediction-correctness ratio up to 10.0\% higher than the current state-of-the-art, at 75.0\%.
 
@@
 
-Hurricanes are a specific type of tropical cyclone exclusive to the Atlantic basin \cite{def-hurricane}.
-Hurricanes are divided into five categories according the Saffir-Simpson scale, based on wind-speed and expected damage \cite{hurricane-cat}.
-In the United States alone, hurricanes incur damages valued as high as \$157BN \cite{hurricane-cost}.
+Hurricanes are a specific type of tropical cyclone exclusive to the Atlantic basin.
+Hurricanes are divided into five categories according the Saffir-Simpson scale, based on wind-speed and expected damage.
+In the United States alone, hurricanes incur damages which can reach \$157BN \cite{hurricane-cost}.
 The damages incurred by hurricanes do not only include property damage, but also personal danger and in some cases widespread civil unrest \cite{hurricane-cost-non-financial}.
@@
 This paper focuses exclusively on the subfield of hurricane trajectory prediction.
-In particular, the goal of this paper is to investigate and improve on the existing-state-of-the art in hurricane trajectory prediction systems based on sequential data mining.
+In particular, the goal of this paper is to investigate and refine the existing-state-of-the art in hurricane trajectory prediction systems based on sequential data mining.
 
@@
 The consequent of R is then compared against $T_{t}$ using pattern-matching techniques to determine if the prediction was correct.
-The results of HTPM showed that the model was able to achieve a best-case correctness ratio of 65\%.
-In this work, HTPDM will be developed and various strategies will be employed to improve the correctness ratio.
+The results of HTPM showed that the model was able to achieve a best-case correctness ratio of 65.0\%.
+In this work, HTPDM will be developed and various strategies will be employed to increase the correctness ratio.
 Reducing the impact of historical factors, favoring the impact of recent trends, and improving the realistic nature of the hurricane trajectory prediction model result in a new model dubbed WARD-HTP (Weighted-Asset, Realistic-Distance Hurricane Trajectory Predictor).
-The strategies employed in WARD-HTP ultimately result in a correctness ratio of 75.22\%, a 10.22\% improvement over HTPDM.
+The strategies employed in WARD-HTP ultimately result in a correctness ratio of 75.0\%, a 10\% improvement over HTPDM.
 
@@
 Following that, Section VI will detail the experimental setup, including the training data and other experimental variables.
-Section VII will evaluate the results of the experiments, which show that the ideal configurations of WARD-HTP result in a maximal correctness ratio of 75.22\%.
+Section VII will evaluate the results of the experiments, which show that the ideal configurations of WARD-HTP result in a maximal correctness ratio of 75.0\%.
 This work will then conclude with section VIII, capturing the conclusions and directions of future work.
@@
 Additionally, the clusters developed apply specific to typhoons, which are waterborne storms native to the east-pacific seaboard.
-Such storm patterns may not be applicable to patterns found in hurricanes, which are waterborne storms of the west-atlantic seaboard.
+Such storm patterns may not be applicable to patterns found in hurricanes, which are waterborne storms of the west-atlantic basin.
 For this reason, the proposed clustering patterns were not used in this work.
@@
 Little research effort has been allocated to this particular problem space of high economic, and social impact.
-The authors expect this work to constitute a growing effort by academic and government researchers to improve existing methods of hurricane trajectory prediction in the interest of public safety and economic stability.
+The authors expect this work to contribute to the growing effort by academic and government researchers to develop reliable methods of hurricane trajectory prediction in the interest of public safety and economic stability.
 
 The second motivation behind this work is the specific goal of improving the correctness ratio in the current standard data mining model for hurricane trajectory prediction.
-As previously described, HTPDM had a best case trajectory prediction correctness ratio of 65\%.
-This means that of all testing trajectory for a which a matching rule was successfully produced, 65\% of such matches were correct predictions.
+As previously described, HTPDM had a best case trajectory prediction correctness ratio of 65.0\%.
+This means that of all testing trajectory for a which a matching rule was successfully produced, 65.0\% of such matches were correct predictions.
 When HTPDM accounts for testing trajectories for which no rule could be matched, the correctness ratio decreases by 7.5\% to a final value of 57.5\% correctness.
@@
   \State{$L_{1} \gets \{Frequent\ Single\ Points\}$}
-  \For{($k=2; L_{k-1} \neq \emptyset k++)$}
+  \For{($k=2; L_{k-1} \neq \emptyset; k++)$}
     \State{$C_{k} \gets Candidates\ from\ Inner\ Join$}
@@
     \For{$sequence\ S \in HurricaneDB$}
-      \State{$Increase\ c \in C_{k}\ support\ if\ c \in S$}
+      \State{$Increment\ c \in C_{k}\ support\ if\ c \in S$}
     \EndFor
@@
 $T_{1}$ and $T_{2}$ form a new trajectory $T_{3}$ if and only if the \textit{L}-1 terminal coordinates of $T_{1}$ correspond to the \textit{L}-1 initial coordinates of $T_{2}$, where \textit{L} is the current level of AprioriAll being executed.
-For example, if \textit{L}=2, then AprioriAll scans for frequent 2-coordinate sequences, and uses those frequent 2-coordinate sequences to generate 3-coordinate sequences based on trajectory matching.
+For example, if \textit{L}$=$2, then AprioriAll scans for frequent 2-coordinate sequences, and uses those frequent 2-coordinate sequences to generate 3-coordinate sequences based on trajectory matching.
 If the matching condition is satisfied, a new trajectory is generated.
@@
 Frequent trajectory patterns and their resulting trajectory predictions are expressed in terms of the regions where a given hurricane is expected to appear.
-The process of region discretization allows for hurricane coordinate trajectories to be scaled down, allowing the regionalized coordinates to be treated as a simplified coordinate system in subsequent aspects of the prediction model.
+The process of region discretization scales all hurricane coordinate trajectories downard, allowing the regionalized coordinates to be treated as a simplified coordinate system in subsequent aspects of the prediction model.
 
@@
 Modified AprioriAll, downscaled region discretization and pattern matching with respect to association rule antecedents and consequents are respectively employed by HTPDM in order to develop an effective system for predicting future hurricane trajectories.
-Based on Atlantic hurricane trajectory data from 1900 to 2008, the model generates a best-case correctness ratio of 65\% and a worst-case ratio of 57.5\%.
+Based on Atlantic hurricane trajectory data from 1900 to 2008, the model generates a best-case correctness ratio of 65.0\% and a worst-case ratio of 57.5\%.
 The training data used in the original model consists of trajectories from 1900 to 2000, with testing data supplied by trajectories from 2001 to 2008.
@@
 When accounting for all tested trajectories, including those which were matched to a rule as a result of trajectory extrapolation, the correctness ratio was 57.5\%.
-Therefore, the best case correctness ratio of HTPDM is the ratio which ignores all test trajectories which did not match with any rules, resulting in a final value of 65\%.
+Therefore, the best case correctness ratio of HTPDM is the ratio which ignores all test trajectories which did not match with any rules, resulting in a final value of 65.0\%.
 
-Despite establishing a convincing initial start point, the best-case correctness ratio of 65\% may be too low for HTPDM to be considered realiable in the real-world.
-This is problematic for those interested in hurricane trajectory prediction as a model with only 65\% accuracy may not be able to deliver crucial insight into hurricane trajectory prediction, particularly with respect to landfall location prediction.
-Both correctness ratios given must be improved in order for data mining to be considered a reliable approach to predicting the trajectories of future hurricanes.
+Despite establishing a convincing initial start point, the best-case correctness ratio of 65.0\% may be too low for HTPDM to be considered realiable in the real-world.
+This is problematic for those interested in hurricane trajectory prediction as a model with only 65.0\% accuracy may not be able to deliver crucial insight into hurricane trajectory prediction, particularly with respect to landfall location prediction.
+Both correctness ratios given must be increased in order for data mining to be considered a practical approach to predicting the trajectories of future hurricanes.
 
@@
 
-In order to improve the correctness ratio of datamining-based hurricane trajectory prediction, this work identifies and explores three opportunities for worst-case and best-case ratio improvement.
-The opportunities for improvement are the improvement of realism, reduction of historical impact, and favorability towards recent hurricane trends.
-Therefore, the opportunities for improvement are realized by three unique dimensions.
-These dimensions are the use of a distance formula designed for spherical coordinate systems, the Haversine Distance Formula, the use of weighted training data when executing the modified variant of AprioriAll, and the modification of the training data to only use historical trajectories from 1950 to 2000.
+In order to increase the correctness ratio of datamining-based hurricane trajectory prediction, this work identifies and explores three opportunities for worst-case and best-case ratio refinement.
+The opportunities identified are the introduction of more realism in calculation formulas, reduction of historical impact on data used, and favorability towards recent hurricane trends.
+The opportunities for improvement are realized by three unique dimensions.
+First, the Haversine Distance Formula is employed for calculating realistic distances in spherical coordinate systems.
+Second, the use of weighted training data will be employed in AprioriAll to favor more recently recored hurricane trajectories.
+Finally, the training data used will only constitute hurricanes from 1950 onwards to remove the risk of relying on misrecorded historical data.
 The implementation of these three dimensions results in a new hurricane trajectory prediction model dubbed WARD-HTP (Weighted-Asset, Realistic-Distance Hurricane Trajectory Predictor).
-The combined contributions of WARD-HTP are expected to result in improvements to the best-case and worst-case correctness ratios.
+The combined contributions of WARD-HTP are expected to result in the desired increase to the best-case and worst-case correctness ratios.
 
@@
 
+The first dimension of improvement introduced by WARD-HTP is the introduction of a realistic distance formula for hurricane trajectory comparison.
+Two different distance measurement formulas are explored in this work for use in the correctness checking stage of the hurricane trajectory prediction model.
+This work compares the Euclidean distance formula against the Haversine distance formula.
+The Euclidean distance formula (as used in \cite{major-hurricane-model}) is designed to determine the unit distance between two N dimensional points on a plane.
+In HTPDM, units correspond to degrees of latitude and longitude.
+The Euclidean distance formula is evaluated with two dimensions, corresponding respectively to latitude and longitude values.
+
+In contrast to the Euclidean formula, the Haversine Distance Formula is designed to determine the kilometers between two latitude-longitude points over a large sphere, such as the earth's surface.
+Despite the fact that the earth's surface is not a perfect sphere \cite{lat-long-distance}, the Haversine Distance Formula is sufficiently accurate when used within a specific region such as the Atlantic Basin.
+The two formulas, denoted $D_{eucl}$ and $D_{havr}$ are defined as follows:
+\begin{align*}
+& (1)\ D_{eucl}(p, q) = \sqrt{(\Delta Lat)^2 + (\Delta Lon)^2}\\
+& (2)\ D_{havr}(p, q) = 2R(\atantwo(\sqrt{\alpha}, \sqrt{1-\alpha}))\\
+& \alpha = \sin^2(\frac{\Delta Lat}2) + \cos(p_{lat}) + \cos(q_{lat}) + \sin^2(\frac{\Delta Lon}2)
+% & \Delta Lat = (q_{lat} - p_{lat})/2\\
+% & \Delta Lon = (q_{lon} - p_{lon})/2
+\end{align*}
+% H is in RADS, Eucl in degrees
+Where $\Delta Lat$ and $\Delta Lon$ respectively denote the differences in latitude and longtitude between points q and p.
+Note that the Euclidean distance formula expresses Latitude and Longtitude in degrees and minutes, while the Haversine formula expresses Latitude and Longtitude exclusively in radians.
+In addition, whereas the Euclidean formula expresses distance in terms of units on a plane, the Haversine formula expresses distance in terms of portions of a 1-unit sphere.
+Therefore, to express distance in kilometers, the result of Haversine formula must be multiplied by R, which denotes the radius of the earth in kilometers (approximately 6371km \cite{earth-radius-kilometers}).
+
+The euclidean distance in the discretized coordinate system corresponds to a difference in square regions approximately equal to the specified region dimension.
+Similar to the example outlined in section 4.2, suppose the region dimension of the discretized coordinate system is once again 5.
+After discretizing the coordinate system into regions of square size, a Euclidean distance of 1 between two 5-degree square regions $R_{1}$ and $R_{2}$ means that $R_{1}$ and $R_{2}$ are approximately 1 region, or 5 degrees, or 500km apart.
+
+When using the Euclidean formula, the distance threshold is 1.
+This corresponds to the distance threshold proposed in \cite{major-hurricane-model}.
+When the Haversine formula is employed, the maximal distance is varied from 300 to 450 km.
+This value corresponds to slightly less than approximately one average hurricane-diameter.
+Since the average hurricane diameter is between 333 and 670 km \cite{hurricane-distances}, two points are considered to be within the distance threshold if they are less than one average hurricane-diameter apart.
+
+The maximal distance ranges used for the Haversine formula are much less than used in the Euclidean formula.
+Therefore, the expected outcome of employing the Haversine formula is the production of finer-grained hurricane trajectory prediction results when compared to results generated by use of the Euclidean distance formula.
+
 \subsection{Weighted Training Data}
+
+The second approach to increasing trajectory prediction correctness deployed in WARD-HTP is the usage of weighted training data in AprioriAll.
+Applying weights to recorded trajectories in chronological order allows AprioriAll to favor recently produced trajectory patterns by increasing the support and confidence of such recently produced trajectories.
+While the weighting of training data results in some trajectories being counted with a support greater than 1, this effect is balanced by counting earlier recorded trajectories with a support less than 1.
+In WARD-HTP, we weigh hurricane trajectories on a scale increasing linearly, resulting in a tipping point \textit{T} where records before \textit{T} are underweighted and records after \textit{T} are overweighted.
+Depending on the weighing approach used, the weights of all records does not necessarily need to be balanced.
+
+The following algorithm applies further modifications to AprioriAll in order to weigh the stored hurricane records.
+In this algorithm, W identifies the total weighing scale to be used in the modification.
+For example, if W$=$3, then the final recored hurricane trajectory in the hurricane database HDB will be counted with a support of 3.
+In addition, the first trajectory will be counted with a support of 3/$\lvert$HDB$\rvert$.
+The support modifier increases by 3/$\lvert$HDB$\rvert$ for each entry in HDB.
+
+% give weighted algorithm and explain what + why
+%Weighted AprioriAll Pseudocode
+% Modify this for weighted stuff
+\begin{algorithm}[H]
+  \caption{Modified AprioriAll With Weighted Support}
+  \label{aprioriall_for_hurricanes}
+  \begin{algorithmic}[1]
+  \For{($k=1; L_{k-1} \neq \emptyset; k++)$}
+    \State{$ w \gets \frac{W}{\lvert HDB \rvert}$}
+    \State{$C_{k} \gets Candidates\ (Inner\ Join\ or\ Singletons)$}
+    \State{$Prune\ C_{k}\ of\ infrequent\ subsequences\ $}
+    \For{$sequence\ S \in HDB$}
+      \State{$Increase\ c \in C_{k}\ support\ by\ w\ if\ c \in S$}
+      \State{$w \gets w + \frac{W}{\lvert HDB \rvert}$}
+    \EndFor
+    \State{$L_{k} \gets \{c \in C_{k}|\ sup(c) \geq minsup\}$}
+  \EndFor
+  \State{$Frequent Trajectories \gets All(L_{k})$}
+  \end{algorithmic}
+\end{algorithm}
+
+The algorithm described assumes that the HDB is ordered chronologically.
+This ensures that recently produced hurricane trajectory records correspond to a higher weight.
+Iteration through the HDB increases the current weight w by the weighting scale W divided by the size of the HDB.
+Therefore the final record in the HDB is guaranteed to have a weighted support exactly equal to W.
+Alternatively, the support weight modifier could also change every year.
+This means that W would be divided by the number of years which the database records span, and the value of w would increase every time the records reach a new year.
+Though index-based weighting is shown above in algorithm (3), both weighting strategies will be evaluated further in this work.
+
+Under the assumption of chronologically ordered data, introducing weighted support to AprioriAll will result in higher support for patterns which occur in recent years.
+Fewer repeated records will be needed to generate frequent sequences from recent data, and more repeated records will be needed to generate frequent sequence from older data.
+Therefore, the introduction of weighted support to AprioriAll is expected to further increase the hurricane trajectory prediction correctness ratio for current and future trajectories being tested upon.
 
@@
 
+The final improvement dimension proposed in WARD-HTP is the reconsideration of 1900 to 2000 as the trajectory prediction training data interval.
+In its original form, HTPDM is trained on hurricane trajectory data from 1900 to 2000.
+However, incorporating training data from years as early as 1900 introduces a risk of skewing frequent trajectory patterns in favor of historical errors and anomalies.
+While weighted support in AprioriAll helps reduce the impact of historical anomalies, the primitivity of early hurricane recording technology makes hurricane trajectories recorded before 1950 much less reliable than those recorded after \cite{hurricane-recording-history}
+
+The hurricane trajectory prediction model proposed in this work relies on test data exclusively from trajectories recorded between the years 1950 and 2000.
+The motivation being selecting this specific year interval is the attempted reduction of historical factors in the generation of frequent trajectory patterns (and subsequent interesting trajectory association rules).
+Paired with the Haversine Distance Formula and weighted AprioriAll approaches, tweaking the training data year interval is expected to result in an additional increase of the model correctness ratio.
+Additionally, association rules generated by training data from 1950 and onwards are expected to be more reliable as the influence of historical factors will be removed.
+
 \section{Experimental Setup}
+
+In its fully implemented form, WARD-HTP was run against various input parameters to determine an optimal configuration.
+WARD-HTP was evaluated against varying distance formula styles, weighted training data approaches (including the absence of weighted data), and training data year intervals.
+Evaluation of WARD-HTP was conducted on an eight-core Intel(R) i7-4720 processor at 3.35GHz, with 8 GB of memory using a Python implementation.
+The training and testing data was supplied from Hurricane DATabase 2 (HURDAT2), which is maintained by the National Oceanic and Atmospheric Administration \cite{HURDAT2-original}.
+
+Apart from the three dimensions of improvement outlined, one of the key differences between HTPDM and WARD-HTP is that WARD-HTP does not attempt to extrapolate points for hurricane trajecories with which no rules were successfully matched.
+Therefore, results shown from WARD-HTP capture the correctness ratio when considering all test trajectories, and the correctness ratio when exclusively considering trajectory with which a rule was immediately successfully matched.
+Rule-match extrapolation was omitted from WARD-HTP due to the nature of the trajectories recored in HURDAT2.
+
+\begin{figure}[htp]
+\caption{Non-Matchable Hurricane Trajectory Paths}
+\centering
+\includegraphics[scale=0.6]{Rule-Miss-Diagram}
+\end{figure}
+
+% careful about color: what if we publish in black and white
+Figure (3) illustrates the characteristics of trajectories which were not able to match with any interesting rules.
+The particular diagram was generated from training data ranging from 1950 to 2000, and testing data from 2001 to 2015.
+Red trajectories represent a subset of the training data, while blue trajectories are those which did not match with any association rules.
+It is visually clear to see that non-matching trajectories were highly dissimilar from training data trajectories.
+As a result, no amount of point extrapolation would be able to result in a reliable terminal trajectory prediction.
+Some of the non-matching trajectories appear to line up with training trajectories, but are moving in an opposite direction.
+Therefore, any points they match with would also not provide a reliable result.
 
@@
 
+This section of the paper documents the experimental results of all improvement dimensions proposed in WARD-HTP.
+Each dimension of Haversine Distance, Weighted Training Data and Training Data Year Interval of 1950-2000 contributes iteratively to a total growth in correctness ratio of 10.0\% when compared to HTPDM, for a total of 75.0\%.
+Whereas the worst-case correctness ratio (ratio accounting for rule-miss incidents) of HTPDM was 57.5\%, the worst-case correctness ratio of WARD-HTP was 72.58\% when configured in the same manner that generated the best-case result of 75.0\%.
+
+The worst-case to best-case difference in HTPDM is 7.5\%, whereas the difference between the two cases in WARD-HTP is 2.44\%.
+This results in a difference decrease of 5.08\%.
+Not only is the difference in ratios reduced when using WARD-HTP, the best and worse case ratios are also increased, by 10.0\% and 15.08\% respectively.
+
+\subsection{Haversine Distance Formula}
+\begin{figure*}[!bp]
+\centering
+\includegraphics[scale=1.0]{Distance-Formula-Results}
+\caption{Results of Different Distance Formulas}
+\end{figure*}
+
+The Distance Formula section of the experiments in this work compare the Euclidean Distance approach as seen in HTPDM against two separate Haversine Distance approaches.
+The first approach imposes a distance threshold \textit{D} of 300km while the second sets \textit{D} to 450km.
+300km was selected as a sample value for testing the Haversine formula, while 450km was selected as a value slightly less than the approximate average hurricane diameter discussed in section 4-B.
+
+Aside from differences in the distance approaches, all other variables in this experiment are held constant.
+The model was trained on trajectories recorded between 1950 and 2000 and tested on tracjectories recorded between 2001 and 2015.
+The parameters of AprioriAll were set to minsup$=$3.75\% (28.3125) and minconf$=$0.25.
+The training data in AprioriAll was unweighted (weighing scale$=$1), and rule-miss extrapolation was disabled as previous discussed.
+
+As shown in figure (4), the Euclidean distance option resulted in much lower correctness ratios than both the 300km and 450km Haversine approaches.
+However, the difference between best and worst case correctness ratios is lowest when using Euclidean distance.
+Since increasing the ratios is prioritized over reducing the ratio difference, the Haversine approach of 450km produces the best result.
+The distance threshold of 450km introduces more flexibility into which predictions count as correct.
+The increase in flexibility is permissible as 450km still corresponds to a size small than the approximate average hurricane diameter.
+The results of this section show that using the Haversine Distance formula with a threshold of 450km is the ideal approach to evaluating distance when predicting hurricane trajectories with data mining.
+
+\subsection{Weighted Training Data}
+
+\begin{figure*}[!bp]
+\centering
+\includegraphics[scale=0.8]{Patterns-Rules-Weight-Metrics}
+\caption{Minsup and Weighing Strategies Against Patterns and Rules}
+\end{figure*}
+
+In this section of the evaluation against WARD-HTP, multiple full executions of the model from AprioriAll to the correctness ratio analysis are completed.
+Each execution of WARD-HTP is performed on a unique pairing of minimum support and the training data weighing strategy with the weighing scale W set to 1 or 2.5.
+Apart from the minimum support and the weighing strategy, all other parameters of the model are held constant.
+
+Three different weighing strategies are evaluated in this experiment.
+First, the absense of data weighing is explored (W$=$1).
+Next, weighing on an annual scale is considered.
+In this strategy, the value of w increases by W divided by the total number of years in the data Y (in the case where training data is supplied from 1950 to 2000, Y$=$50).
+The third strategy employed is the index-weight strategy, where the weight is updated at each HDB entry as seen in section 5-B.
+
+Various values of minimum support are also explored in this series of evaluations.
+Minimum support values correspondong to 2.5\%, 3.75\% and 5\% of the number of training data trajectories are paired against the three weighing strategies discussed above.
+While a higher minimum support will generate fewer interesting trajectory association rules, the rules generated will be more reliable for future predictions.
+
+In figures (5) and (6), the number of frequent trajectories and trajectory association rules generated is compared alongride the best and worse care hit rates produced for given pairings of minsup and the weighing strategy.
+Note that the minconf value is held at 0.25 for all experiments, along with the training data of 1950-2000 from HURDAT2, testing data from 2001-2015, and distance formula set to Haversine Distance with a threshold of 450.
+
+The maximal best-case hit rate for WARD-HTP was generated with an index-weighted strategy with minimum support of 2.5\%, corresponding to 18.875 when used with data from 1950 to 2000.
+The resulting hit rate was 75.0\%, with a worst-case of 72.58\%.
+Note that minimum support of 5\% with index-based weighting generated a very slightly higher best-case hit rate at 75.22\%, albeit with a lower worst-case ratio of 67.33\%.
+We select 75\% as the optimal generated data point because a .22\% decrease in best-case hit rate is certainly a ``fair trade'' for a 5.25\% increase in worst-case hit rate.
+
+The year-based and index-based weighting strategies generated identical model results at minsup$=$2.5\%.
+We select the year-based weighting strategy as the optimal approach due to the fact that the year-based strategy evenly counts trajectory records with respect to time.
+In the year-based strategy, all records from the most recent year are counted with at the maximum weight scale, whereas the index-based strategy only counts one record at the maximum.
+While the two weighting strategies resulted in identical results in this scenario, the year-based strategy is preferred as it encourages an increase in support weight closely coupled to real-world historical trends.
+
+% talk a little more then done
+\begin{figure*}[!tp]
+\centering
+\includegraphics[scale=0.8]{Hit-Rate-Weight-Metrics}
+\caption{Minsup and Weighing Strategies Against Correctness Ratios}
+\end{figure*}
+
+\subsection{Training Data Year Interval}
+
+% some training data notes
+The final component of the evaluation undertaken in this work compares the correctness ratios generated by various training and testing data intervals.
+Training data from 1900 to 2000 and 1950 to 2000 is used against testing data from 2001 to 2008 and 2001 to 2015.
+All other variables are held constant with minsup$=$2.5\%, weighting strategy set to yearly and the distance formula set to Haversine with \textit{D}$=$450.
+Note that the data ranges used by HTPDM were 1900-2000 for training data and 2001-2008 for test data.
+
+% figure
+\begin{figure}[htp]
+\centering
+\includegraphics[scale=1.0]{Year-Interval-Metrics}
+\caption{Year Interval Options Against Correctness Ratios}
+\end{figure}
+% explain. done.
+
+Figure (7) shows the correctness ratios generated by all possibly training data and testing data pairings.
+Note that ratios in brackets denote the worse-case ratio while ratios outside of brackets refer to the best-case.
+Between all possible options, we find that the training data interval of 1950-2000 results in the best correctness ratios for testing data from 2001-2015 when applying all other parameters we deemed optimal throughout the previous results of this work.
+Regardless of correctness ratio, 2001-2015 should always be selected as the test data interval to encourage analysis against more recent hurricane patterns.
+The results show that no training data-testing data pairing was able to generate correctness ratios higher than the ones generated by training the model with data from 1950 to 2000, and testing it on data as recently recorded as 2015.
+
 \section{Conclusions and Future Work} % what we found:
 
+This work proposed a new model named the Weighted-Asset, Realistic-Distance Hurricane Trajectory Predictor (WARD-HTP) for improving the accuracy of existing data mining approaches to hurricane trajectory prediction.
+An AprioriAll-based hurricane trajectory prediction model similar to \cite{major-hurricane-model} was developed to demonstrate the unique contributions of WARD-HTP.
+There are three such contributions brought forth in the development and implementation of the model proposed.
+These contributions are the use of training data from 1950 to 2000, the use of weighted training data, or weighted assets (WA), and the use of the Haversine Distance Formula for realistic distance (RD).
+The three contributions combined form WARD-HTP, and produced a hurricane trajectory prediction correctness ratio of exactly 75.0\%, which is 10\% higher than the current state-of-the-art.
+When account for trajectories which were not able to match with any association rules, the correctness ratio of WARD-HTP is 72.58\%, which is still 5.08\% higher than the worst case state-of-the-art ratio of 57.5\% as seen in \cite{major-hurricane-model}
+
+There are mutliple areas to explore in the hurricane trajectory prediction problem space with respect to future work.
+Experiments in this work show that all datasets contain a subset of trajectories for which no association rule could be matched.
+While \cite{major-hurricane-model} proposed an extrapolation method to recover from rule-miss-incidents, more work must be completed to reduce the number of non-matchable trajectories.
+Many of the mismatched trajectories are outliers with very little conformance to existing hurricane trajectories as seen in figure (3).
+Additionally, more experiments will be run on different inputs including training data from other regions such as the Pacific region.
+Further approaches to increasing the correctness ratio of WARD-HTP will also be investiaged.
+This includes the investigation of alternative methods for defining training data weight.
+
 \begin{thebibliography}{1}
-\bibitem{def-hurricane}  
-\bibitem{hurricane-cat} 
-\bibitem{hurricane-cost} 
-\bibitem{hurricane-cost-non-financial} 
-\bibitem{data mining-def} 
-\bibitem{major-hurricane-model} 
-\bibitem{HURDAT2-original} 
-\bibitem{AprioriAll-original} 
-\bibitem{hurricane-intensity-1}
-\bibitem{hurricane-intensity-2}
-\bibitem{hurricane-forecasting-1}
-\bibitem{hurricane-forecasting-2}
-\bibitem{typhoon-clustering}
-\bibitem{matplotlib}
-\bibitem{lat-long-distance}
-\bibitem{hurricane-distances}
+\bibitem{hurricane-cost} Pielke Jr, Roger A., et al. Normalized hurricane damage in the United States: 1900–2005. Natural Hazards Review 9.1 (2008): 29-42
+\bibitem{hurricane-cost-non-financial} F. F. Townsend, The Federal Response to Hurricane Katrina: Lessons Learned, United States Federal Government, Washington, DC, USA, 2006
+\bibitem{data mining-def} S. Sumathi and S. N. Sivanandam, Introduction to Dataming and its Applications: Springer Publishing 2006, ISBN: 978-3-540-34350-9
+\bibitem{major-hurricane-model} Dong, X and C.
+Pi, D.
+(2013).
+Novel method for hurricane trajectory prediction based on data mining.
+Natural Hazards and Earth System Sciences.
+13.
+3211-3220.
+10.5194/nhess-13-3211-2013.
+\bibitem{HURDAT2-original} C. Landsea, J. Franklin and J. Beven (2015) The revised Atlantic hurricane database (HURDAT2). National Oceanic and Atmospheric Administration
+\bibitem{AprioriAll-original} R. Agrawal and R. Srikant, Mining sequential patterns, Proceedings of the Eleventh International Conference on Data Engineering, Taipei, 1995, pp. 3-14.
+doi: 10.1109/ICDE.1995.380415
+\bibitem{hurricane-intensity-1} Y.
+Su, S.
+Chelluboina, M.
+Hahsler and M.
+H.
+Dunham, A New Data Mining Model for Hurricane Intensity Prediction, 2010 IEEE International Conference on Data Mining Workshops, Sydney, NSW, 2010, pp.
+98-105
+doi: 10.1109/ICDMW.2010.158
+\bibitem{hurricane-intensity-2} R.
+Yang, J.
+Tang and D.
+Sun (2011), Association Rule Mining Application for Atlantic Tropical Cyclone Intensity Changes, American Meteorological Society.
+doi: 10.1175/WAF-D-10-05029.1
+\bibitem{hurricane-forecasting-1} Rozanova, Olga S., Jui-Ling Yu, and Chin-Kun Hu.
+Typhoon eye trajectory based on a mathematical model: Comparing with observational data.
+Nonlinear Analysis: Real World Applications 11.3 (2010): 1847-1861.
+\bibitem{hurricane-forecasting-2} Chandan Roy and Rita Kovordanyi, Tropical cyclone track forecasting techniques: A review, 2012, Atmospheric research, (104-105), 40-69.
+doi: 10.1016/j.atmosres.2011.09.012.
+\bibitem{typhoon-clustering} Kim, Hyeong-Seog, et al. Pattern classification of typhoon tracks using the fuzzy c-means clustering method. Journal of Climate 24.2 (2011): 488-508.
+\bibitem{matplotlib} J. D. Hunter, Matplotlib: A 2D Graphics Environment, in Computing in Science and Engineering, vol. 9, no. 3, pp. 90-95, May-June 2007.
+doi: 10.1109/MCSE.2007.55
+\bibitem{lat-long-distance} Menno-Jan, K.
+(2013).
+Map use: reading, analysis, interpretation: by Kimberling, AJ, Buckley, AR, Muehrcke, PC and Muehrcke, JO, Redlands, California, ESRI Press Academic, 2012.
+ISBN 978-1-589-48279-1 Seventh edition.
+\bibitem{hurricane-distances} D. Brown. (2013) Forecast Verification: Quantifying Forecast Uncertainty. National Hurricane Conference: Hurricane Readiness for Coastal Communities
+\bibitem{earth-radius-kilometers} Chambat, F., and Valette, B. (2001). Mean radius, mass, and inertia for reference Earth models. Physics of the Earth and Planetary Interiors, 124(3), 237-253.
+\bibitem{hurricane-recording-history} Landsea, C.
+W., Vecchi, G.
+A., Bengtsson, L., and Knutson, T.
+R.
+(2010).
+Impact of duration thresholds on Atlantic tropical cyclone counts.
+Journal of Climate, 23(10), 2508-2519.
 
@@
 % we will need to re-read the paper and keep an eye out for sections that suffer from scope creep. Problem description first comes to mind.
+
+% we need to make sure we are not jumping to far off the deep end with hurricanes and forgetting to discuss d a t a  m i n i n g

### 2017-12-08 paper/manuscript.tex
minor grammar/spelling fixes

@@
 We propose the use of Haversine distance, a real-world distance formula for reliable hurricane pattern matching accuracy.
-We also propose the use of weighted data to favor recent hurricane trends when training the trajectory prediction model.
+We also propose the use of weighted data to favour recent hurricane trends when training the trajectory prediction model.
 Finally, we use training data exclusively from the years 1950 to 2000 to reduce the impact of historical inaccuracies in hurricane recording methodologies.
@@
 Hurricanes are divided into five categories according the Saffir-Simpson scale, based on wind-speed and expected damage.
-In the United States alone, hurricanes incur damages which can reach \$157BN \cite{hurricane-cost}.
+In the United States alone, hurricanes which can incur \$157BN in damages \cite{hurricane-cost}.
 The damages incurred by hurricanes do not only include property damage, but also personal danger and in some cases widespread civil unrest \cite{hurricane-cost-non-financial}.
@@
 Data mining applies to an extremely wide range of applications, including clickstream mining, social network community detection, and artificial intelligence.
-The data mining subdiscipline of particular interest in this work is \textit{sequential} data mining.
+The data mining sub-discipline of particular interest in this work is \textit{sequential} data mining.
 Sequential data mining uses a database of chronologically ordered inputs to mine frequent patterns and interesting association rules.
@@
 In this work, HTPDM will be developed and various strategies will be employed to increase the correctness ratio.
-Reducing the impact of historical factors, favoring the impact of recent trends, and improving the realistic nature of the hurricane trajectory prediction model result in a new model dubbed WARD-HTP (Weighted-Asset, Realistic-Distance Hurricane Trajectory Predictor).
+Reducing the impact of historical factors, favouring the impact of recent trends, and improving the realistic nature of the hurricane trajectory prediction model result in a new model dubbed WARD-HTP (Weighted-Asset, Realistic-Distance Hurricane Trajectory Predictor).
 The strategies employed in WARD-HTP ultimately result in a correctness ratio of 75.0\%, a 10\% improvement over HTPDM.
@@
 The remainder of this paper is structured as follows: First, Section II gives a detailed survey of related work in the hurricane prediction and sequential mining disciplines.
-Next, section III will cover the general motivations of this work with respect to hurricane trajectory prediction, and the specific motivations of this work with respect to improving the current state-of-the-art.
+Next, section III will cover the general motivations of this work with respect to hurricane trajectory prediction and the specific motivations of this work with respect to improving the current state-of-the-art.
 Section IV will provide a description of the problem space, which includes specific implementation considerations in hurricane trajectory prediction.
-Section V will give an exposition of the solutions proposed in WARD-HTP for increasing the correctness ratio in hurricane trajetory prediction.
+Section V will give an exposition of the solutions proposed in WARD-HTP for increasing the correctness ratio in hurricane trajectory prediction.
 Following that, Section VI will detail the experimental setup, including the training data and other experimental variables.
@@
 While the described work provided novel results in typhoon track clustering, it is not clear if the proposed solution is able to predict the trajectories of new storms.
-Additionally, the clusters developed apply specific to typhoons, which are waterborne storms native to the east-pacific seaboard.
-Such storm patterns may not be applicable to patterns found in hurricanes, which are waterborne storms of the west-atlantic basin.
+Additionally, the clusters developed apply specifically to typhoons, which are waterborne storms native to the east-pacific seaboard.
+Such storm patterns may not be applicable to patterns found in hurricanes, which are waterborne storms of the west Atlantic basin.
 For this reason, the proposed clustering patterns were not used in this work.
@@
 The data in HURDAT2 is ordered into $\langle Hurricane, Trajectory \rangle$ pairs, omitting the need for the sequence mapping step required at initialization of AprioriAll.
-The Atlantic hurricane trajectories recorded in HURDAT2 span a very large distance, some of which reaching over 10,000 kilometers in total distance travelled.
+The Atlantic hurricane trajectories recorded in HURDAT2 span a very large distance, some of which reaching over 10,000 kilometres in total distance travelled.
 As a result, the coordinates recorded are limited to those within a range of 0 to 50 degrees north, and 20 to 100 degrees west.
@@
 There are two categories of motivation behind this work.
-The first motivation of this work is to increase the total investigation completed in the hurricane trajectory prediction subdiscipline.
-Little research effort has been allocated to this particular problem space of high economic, and social impact.
+The first motivation of this work is to increase the total investigation completed in the hurricane trajectory prediction sub-discipline.
+Little research effort has been allocated to this particular problem space of high economic and social impact.
 The authors expect this work to contribute to the growing effort by academic and government researchers to develop reliable methods of hurricane trajectory prediction in the interest of public safety and economic stability.
@@
 The second motivation behind this work is the specific goal of improving the correctness ratio in the current standard data mining model for hurricane trajectory prediction.
-As previously described, HTPDM had a best case trajectory prediction correctness ratio of 65.0\%.
-This means that of all testing trajectory for a which a matching rule was successfully produced, 65.0\% of such matches were correct predictions.
-When HTPDM accounts for testing trajectories for which no rule could be matched, the correctness ratio decreases by 7.5\% to a final value of 57.5\% correctness.
+As previously described, HTPDM achieved a best case trajectory prediction correctness ratio of 65.0\%.
+This means that of all testing trajectory for which a matching rule was successfully produced, 65.0\% of such matches were correct predictions.
+When HTPDM accounted for testing trajectories for which no rule could be matched, the correctness ratio decreased by 7.5\% to a final value of 57.5\% correctness.
 This decrease in accuracy is expected as some hurricane trajectories are outliers which cannot be predicted by frequent pattern mining, as will be discussed further in this work.
@@
 To resolve the motivations of this work, a series of improvement strategies are proposed as applications to the current hurricane trajectory prediction model based on AprioriAll.
-In general, these strategies seek to improve the realism of the prediction model, reduce historical factors which may negatively affect the model's correctness reatio, and favor recent hurricane trajectory trends which are likely to be reflected in current and future hurricane patterns.
+In general, these strategies seek to improve the realism of the prediction model, reduce historical factors which may negatively affect the model's correctness ratio, and favour recent hurricane trajectory trends which are likely to be reflected in current and future hurricane patterns.
 
@@
 In this section, the problem of achieving accurate hurricane trajectory prediction is explained in detail.
-This problem consists of multiple challenges which apply generally to sequential data mining, and specifically to the problem domain of hurricane location analysis.
-Many of the domain-specific challenges of hurricane trajectory prediction arise from the fact that hurricane points are tracked as a series of $\langle Latitude, Longtitude \rangle$ points, which do \textit{not} directly correspond to traditional (x, y) coordinates, as latitude and longtitude are points imposed upon a sphere.
+This problem consists of multiple challenges which apply generally to sequential data mining and specifically to the problem domain of hurricane location analysis.
+Many of the domain-specific challenges of hurricane trajectory prediction arise from the fact that hurricane points are tracked as a series of $\langle Latitude, Longitude \rangle$ points, which do \textit{not} directly correspond to traditional (x, y) coordinates, as latitude and longitude are points imposed upon a sphere.
 Challenges including accurate distance evaluation are results of the transition from a Euclidean to a Polar coordinate system.
@@
 Next, the subproblem of region discretization will be discussed.
-Since $\langle Latitude, Longtitude \rangle$ points constitute continuous space, hurricane trajectories are converted to sequences of discrete regions to enable the discovery of frequent patterns.
+Since $\langle Latitude, Longitude \rangle$ points constitute continuous space, hurricane trajectories are converted to sequences of discrete regions to enable the discovery of frequent patterns.
 Finally, problems involving hurricane trajectory matching will be discussed.
@@
 AprioriAll was first proposed at IBM research as a method for discovering frequent sequential patterns in transactional databases \cite{AprioriAll-original}.
-In its original implementation, AprioriAll is able to generate assocation rules describing sentences of the form ``Customers who buy P may also buy Q''.
-In the hurricane trajectory prediction problem, AprioriAll is used to generate association ruls describing sentences of the form ``Hurricanes located at P may later be located at Q''.
+In its original implementation, AprioriAll is able to generate association rules describing sentences of the form ``Customers who buy P may also buy Q''.
+In the hurricane trajectory prediction problem, AprioriAll is used to generate association rules describing sentences of the form ``Hurricanes located at P may later be located at Q''.
 In the context of AprioriAll, hurricanes are considered $customers$ and their sequences of recorded points are considered $transactions$.
@@
 
-In step (1) of AprioriAll, a given transaction database \textit{D} is grouped by customers, and the frequent items purchased by customers in individual transcations are recorded in a temporary table.
+In step (1) of AprioriAll, a given transaction database \textit{D} is grouped by customers, and the frequent items purchased by customers in individual transactions are recorded in a temporary table.
 The contents of this temporary table are mapped to corresponding sequences to enable the mining of frequent transaction sequences from the original database.
@@
 Figure (1) shows an example of how the trajectory database is structured after the mapping preprocessing step.
-Only two dimensions are needed in the database, being the hurricane ID and the sequence of latitude-longtitude coordinates.
+Only two dimensions are needed in the database, being the hurricane ID and the sequence of latitude-longitude coordinates.
 All hurricanes in HURDAT2 have an ID corresponding to the location (i.e.
@@
 Once the sequence generation level reaches three, each candidate sequence must be pruned if any of their candidate subsequences are infrequent.
-Continuous execution of AprioriAll until completion results in a collection of frequent hurricane trajectory patterns, as defined by the minumum support.
+Continuous execution of AprioriAll until completion results in a collection of frequent hurricane trajectory patterns, as defined by the minimum support.
 
@@
 \begin{algorithm}[H]
-  \caption{Modified AprioriAll: No Litemset phase}
+  \caption{Modified AprioriAll: No L-Itemset phase}
   \label{aprioriall_for_hurricanes}
@@
 
-This subsection covers problems related to the discretization of $\langle Latitude, Longtitude \rangle$ points.
+This subsection covers problems related to the discretization of $\langle Latitude, Longitude \rangle$ points.
 As shown in figure (1), the raw recorded hurricane coordinates are expressed in degrees and minutes.
-The excessively fine-grained nature of the raw data prevents the data from being able to produce meaningful frequent coordinate sequences.
+The excessively fine-grained nature of the raw data prevents it from being able to produce meaningful frequent coordinate sequences.
 Therefore, effort must be taken to discretize the coordinate system, allowing more coarse-grained frequent patterns to be produced.
@@
 In HTPDM and WARD-HTP, recorded hurricane coordinates are discretized into approximately square blocks of a specific size, expressed in degrees.
-Consider the first row of coordinates shown in figure (1): (12.5:-37.5), (13.0:-39.0), (13.5:-40.5).
-Mapping this coordinate sequence to a discritization block of size 5, the new trajectory becomes (2:-7), (2:-7), (2:-8).
+Consider the third row of coordinates shown in figure (1): (12.5:-37.5), (13.0:-39.0), (13.5:-40.5).
+Mapping this coordinate sequence to a discretization block of size 5, the new trajectory becomes (2:-7), (2:-7), (2:-8).
 Duplicate coordinates are not counted, so the trajectory ultimately reduces to (2:-7), (2:-8).
@@
 
-Consider figure (2), generated using a sample of the recrded hurricane trajectories in HURDAT2 with the visualization tool Matplotlib \cite{matplotlib}.
+Consider figure (2), generated using a sample of the recorded hurricane trajectories in HURDAT2 with the visualization tool Matplotlib \cite{matplotlib}.
 As shown in the pair of images, region discretization organizes continuous hurricane coordinate sequences into discrete region sequences of approximately uniform size and square shape.
@@
 A block size of 5 corresponds approximately to a real-world block dimension of 500km by 500km.
-The block dimension size is approximate because the mapping between coordinate degrees and kilometers is imprecise, and is dependent on location in the case of longtitudal points \cite{lat-long-distance}.
+The block dimension size is approximate because the mapping between coordinate degrees and kilometres is imprecise, and is dependent on location in the case of longitudal points \cite{lat-long-distance}.
 The approximate size of 500km is similar to the average hurricane total diameter \cite{hurricane-distances}, permitting the coarse-grain expression of hurricane trajectories without undue loss in precision.
@@
 Frequent trajectory patterns and their resulting trajectory predictions are expressed in terms of the regions where a given hurricane is expected to appear.
-The process of region discretization scales all hurricane coordinate trajectories downard, allowing the regionalized coordinates to be treated as a simplified coordinate system in subsequent aspects of the prediction model.
+The process of region discretization scales all hurricane coordinate trajectories downward, allowing the regionalized coordinates to be treated as a simplified coordinate system in subsequent aspects of the prediction model.
 
@@
 Pattern matching is used when evaluating the best-fit rule for hurricane prediction purposes.
-Similar to HTPDM, a new hurricane trajectory is passed into the hurricane prediction model proposed in this work, it is separated into two subtrajectories: the initial trajectory $T_{i}$ and the terminal trajectory $T_{t}$.
-$T_{i}$ is compared against the antecedent of every interesting assocation rule satisfying minconf for a match.
+Similar to HTPDM, a new hurricane trajectory is passed into the hurricane prediction model proposed in this work, it is separated into two sub-trajectories: the initial trajectory $T_{i}$ and the terminal trajectory $T_{t}$.
+$T_{i}$ is compared against the antecedent of every interesting association rule satisfying minconf for a match.
 There are two dimensions that determine the best-fit rule for an incoming trajectory requiring prediction.
@@
 \end{eqnarray*}
-Where r is a member of the set of all interesting assocation rules \textit{R}, t is the initial trajectory $T_{i}$, and $matchLen_{t,r}$ expresses the matching length between $T_{i}$ and the antecedent sequence of r.
+Where r is a member of the set of all interesting association rules \textit{R}, t is the initial trajectory $T_{i}$, and $matchLen_{t,r}$ expresses the matching length between $T_{i}$ and the antecedent sequence of r.
 This formula is meant to be maximized to determine the rule which most generously fits the incoming hurricane trajectory.
@@
 First, prediction correctness uses matching length with a different definition.
-Whereas matching length for rule antecedents is the longest identical congruent subsequence of the initial trajectory, matching length for the rule consequent is the longest congurent subsequence of points within a specific distance, beginning at the start of the terminal sequence.
+Whereas matching length for rule antecedents is the longest identical congruent subsequence of the initial trajectory, matching length for the rule consequent is the longest congruent subsequence of points within a specific distance, beginning at the start of the terminal sequence.
 This methodology is also based on HTPDM and is based on the following steps.
@@
 
-The hurricane trajecory prediction model based on data mining as originally proposed in \cite{major-hurricane-model} implements all aspects of the problem domain described.
+The hurricane trajectory prediction model based on data mining as originally proposed in \cite{major-hurricane-model} implements all aspects of the problem domain described.
 Modified AprioriAll, downscaled region discretization and pattern matching with respect to association rule antecedents and consequents are respectively employed by HTPDM in order to develop an effective system for predicting future hurricane trajectories.
@@
 The training data used in the original model consists of trajectories from 1900 to 2000, with testing data supplied by trajectories from 2001 to 2008.
-At a high level, the model is developed and tested by following the steps below, using all implementation strategies discussed.
+At a high level, the model is developed and tested by following the steps below, using all implementation strategies discussed:
 
@@
 \begin{enumerate}
-\item Generate interesting trajectory rules from AprioriAll
-\item Determine matching rule for incoming trajectory 
-\item Check if the matched rule's consequent points are sufficiently close
+\item Generate interesting trajectory rules from AprioriAll.
+\item Determine matching rule for incoming trajectory.
+\item Check if the matched rule's consequent points are sufficiently close.
 \end{enumerate}
@@
 The differentiation between the model's best and worst-case hit rates is dependent on the ability of an incoming test trajectory to be matched with a trajectory association rule.
-As described in the pattern matching stage, the initial subtrajectory of an incoming trajectory is compared against all interesting rules to determine a best-fit match.
-If no such match is found, HTPDM attempts to extrapolate a future point of the initial trajectory, using the extrapolated point as a new option for determing a rule antecedent match.
+As described in the pattern matching stage, the initial sub-trajectory of an incoming trajectory is compared against all interesting rules to determine a best-fit match.
+If no such match is found, HTPDM attempts to extrapolate a future point of the initial trajectory, using the extrapolated point as a new option for determining a rule antecedent match.
 When accounting for all tested trajectories, including those which were matched to a rule as a result of trajectory extrapolation, the correctness ratio was 57.5\%.
@@
 
-Despite establishing a convincing initial start point, the best-case correctness ratio of 65.0\% may be too low for HTPDM to be considered realiable in the real-world.
+Despite establishing a convincing initial start point, the best-case correctness ratio of 65.0\% may be too low for HTPDM to be considered reliable in the real-world.
 This is problematic for those interested in hurricane trajectory prediction as a model with only 65.0\% accuracy may not be able to deliver crucial insight into hurricane trajectory prediction, particularly with respect to landfall location prediction.
@@
 
-In order to increase the correctness ratio of datamining-based hurricane trajectory prediction, this work identifies and explores three opportunities for worst-case and best-case ratio refinement.
-The opportunities identified are the introduction of more realism in calculation formulas, reduction of historical impact on data used, and favorability towards recent hurricane trends.
+In order to increase the correctness ratio of data mining-based hurricane trajectory prediction, this work identifies and explores three opportunities for worst-case and best-case ratio refinement.
+The opportunities identified are the introduction of more realism in calculation formulas, reduction of historical impact on data used, and favourability towards recent hurricane trends.
 The opportunities for improvement are realized by three unique dimensions.
 First, the Haversine Distance Formula is employed for calculating realistic distances in spherical coordinate systems.
-Second, the use of weighted training data will be employed in AprioriAll to favor more recently recored hurricane trajectories.
-Finally, the training data used will only constitute hurricanes from 1950 onwards to remove the risk of relying on misrecorded historical data.
+Second, the use of weighted training data will be employed in AprioriAll to favour more recently recorded hurricane trajectories.
+Finally, the training data used will only constitute hurricanes from 1950 onwards to remove the risk of relying on mis-recorded historical data.
 The implementation of these three dimensions results in a new hurricane trajectory prediction model dubbed WARD-HTP (Weighted-Asset, Realistic-Distance Hurricane Trajectory Predictor).
@@
 
-In contrast to the Euclidean formula, the Haversine Distance Formula is designed to determine the kilometers between two latitude-longitude points over a large sphere, such as the earth's surface.
+In contrast to the Euclidean formula, the Haversine Distance Formula is designed to determine the kilometres between two latitude-longitude points over a large sphere, such as the earth's surface.
 Despite the fact that the earth's surface is not a perfect sphere \cite{lat-long-distance}, the Haversine Distance Formula is sufficiently accurate when used within a specific region such as the Atlantic Basin.
@@
 % H is in RADS, Eucl in degrees
-Where $\Delta Lat$ and $\Delta Lon$ respectively denote the differences in latitude and longtitude between points q and p.
-Note that the Euclidean distance formula expresses Latitude and Longtitude in degrees and minutes, while the Haversine formula expresses Latitude and Longtitude exclusively in radians.
+Where $\Delta Lat$ and $\Delta Lon$ respectively denote the differences in latitude and longitude between points q and p.
+Note that the Euclidean distance formula expresses Latitude and Longitude in degrees and minutes, while the Haversine formula expresses Latitude and Longitude exclusively in radians.
 In addition, whereas the Euclidean formula expresses distance in terms of units on a plane, the Haversine formula expresses distance in terms of portions of a 1-unit sphere.
-Therefore, to express distance in kilometers, the result of Haversine formula must be multiplied by R, which denotes the radius of the earth in kilometers (approximately 6371km \cite{earth-radius-kilometers}).
+Therefore, to express distance in kilometres, the result of Haversine formula must be multiplied by R, which denotes the radius of the earth in kilometres (approximately 6371km \cite{earth-radius-kilometres}).
 
@@
 The second approach to increasing trajectory prediction correctness deployed in WARD-HTP is the usage of weighted training data in AprioriAll.
-Applying weights to recorded trajectories in chronological order allows AprioriAll to favor recently produced trajectory patterns by increasing the support and confidence of such recently produced trajectories.
+Applying weights to recorded trajectories in chronological order allows AprioriAll to favour recently produced trajectory patterns by increasing the support and confidence of such recently produced trajectories.
 While the weighting of training data results in some trajectories being counted with a support greater than 1, this effect is balanced by counting earlier recorded trajectories with a support less than 1.
@@
 In this algorithm, W identifies the total weighing scale to be used in the modification.
-For example, if W$=$3, then the final recored hurricane trajectory in the hurricane database HDB will be counted with a support of 3.
+For example, if W$=$3, then the final recorded hurricane trajectory in the hurricane database HDB will be counted with a support of 3.
 In addition, the first trajectory will be counted with a support of 3/$\lvert$HDB$\rvert$.
@@
 In its original form, HTPDM is trained on hurricane trajectory data from 1900 to 2000.
-However, incorporating training data from years as early as 1900 introduces a risk of skewing frequent trajectory patterns in favor of historical errors and anomalies.
+However, incorporating training data from years as early as 1900 introduces a risk of skewing frequent trajectory patterns in favour of historical errors and anomalies.
 While weighted support in AprioriAll helps reduce the impact of historical anomalies, the primitivity of early hurricane recording technology makes hurricane trajectories recorded before 1950 much less reliable than those recorded after \cite{hurricane-recording-history}
@@
 
-Apart from the three dimensions of improvement outlined, one of the key differences between HTPDM and WARD-HTP is that WARD-HTP does not attempt to extrapolate points for hurricane trajecories with which no rules were successfully matched.
+Apart from the three dimensions of improvement outlined, one of the key differences between HTPDM and WARD-HTP is that WARD-HTP does not attempt to extrapolate points for hurricane trajectories with which no rules were successfully matched.
 Therefore, results shown from WARD-HTP capture the correctness ratio when considering all test trajectories, and the correctness ratio when exclusively considering trajectory with which a rule was immediately successfully matched.
-Rule-match extrapolation was omitted from WARD-HTP due to the nature of the trajectories recored in HURDAT2.
+Rule-match extrapolation was omitted from WARD-HTP due to the nature of the trajectories recorded in HURDAT2.
 
@@
 
-% careful about color: what if we publish in black and white
+% careful about colour: what if we publish in black and white
 Figure (3) illustrates the characteristics of trajectories which were not able to match with any interesting rules.
@@
 As a result, no amount of point extrapolation would be able to result in a reliable terminal trajectory prediction.
-Some of the non-matching trajectories appear to line up with training trajectories, but are moving in an opposite direction.
-Therefore, any points they match with would also not provide a reliable result.
+Some of the non-matching trajectories appear to align with training trajectories, but are moving in an opposite direction.
+Therefore, any points that match would also not provide a reliable result.
 
@@
 
-This section of the paper documents the experimental results of all improvement dimensions proposed in WARD-HTP.
+This section documents the experimental results of all improvement dimensions proposed in WARD-HTP.
 Each dimension of Haversine Distance, Weighted Training Data and Training Data Year Interval of 1950-2000 contributes iteratively to a total growth in correctness ratio of 10.0\% when compared to HTPDM, for a total of 75.0\%.
@@
 Aside from differences in the distance approaches, all other variables in this experiment are held constant.
-The model was trained on trajectories recorded between 1950 and 2000 and tested on tracjectories recorded between 2001 and 2015.
+The model was trained on trajectories recorded between 1950 and 2000 and tested on trajectories recorded between 2001 and 2015.
 The parameters of AprioriAll were set to minsup$=$3.75\% (28.3125) and minconf$=$0.25.
@@
 The distance threshold of 450km introduces more flexibility into which predictions count as correct.
-The increase in flexibility is permissible as 450km still corresponds to a size small than the approximate average hurricane diameter.
+The increase in flexibility is permissible as 450km still corresponds to a size smaller than the approximate average hurricane diameter.
 The results of this section show that using the Haversine Distance formula with a threshold of 450km is the ideal approach to evaluating distance when predicting hurricane trajectories with data mining.
@@
 Three different weighing strategies are evaluated in this experiment.
-First, the absense of data weighing is explored (W$=$1).
+First, the absence of data weighing is explored (W$=$1).
 Next, weighing on an annual scale is considered.
@@
 Various values of minimum support are also explored in this series of evaluations.
-Minimum support values correspondong to 2.5\%, 3.75\% and 5\% of the number of training data trajectories are paired against the three weighing strategies discussed above.
+Minimum support values corresponding to 2.5\%, 3.75\% and 5\% of the number of training data trajectories are paired against the three weighing strategies discussed above.
 While a higher minimum support will generate fewer interesting trajectory association rules, the rules generated will be more reliable for future predictions.
 
-In figures (5) and (6), the number of frequent trajectories and trajectory association rules generated is compared alongride the best and worse care hit rates produced for given pairings of minsup and the weighing strategy.
+In figures (5) and (6), the number of frequent trajectories and trajectory association rules generated is compared alongside the best and worse case hit rates produced for given pairings of minsup and the weighing strategy.
 Note that the minconf value is held at 0.25 for all experiments, along with the training data of 1950-2000 from HURDAT2, testing data from 2001-2015, and distance formula set to Haversine Distance with a threshold of 450.
@@
 The resulting hit rate was 75.0\%, with a worst-case of 72.58\%.
-Note that minimum support of 5\% with index-based weighting generated a very slightly higher best-case hit rate at 75.22\%, albeit with a lower worst-case ratio of 67.33\%.
+Note that a minimum support of 5\% with index-based weighting generated a very slightly higher best-case hit rate at 75.22\%, albeit with a lower worst-case ratio of 67.33\%.
 We select 75\% as the optimal generated data point because a .22\% decrease in best-case hit rate is certainly a ``fair trade'' for a 5.25\% increase in worst-case hit rate.
@@
 We select the year-based weighting strategy as the optimal approach due to the fact that the year-based strategy evenly counts trajectory records with respect to time.
-In the year-based strategy, all records from the most recent year are counted with at the maximum weight scale, whereas the index-based strategy only counts one record at the maximum.
+In the year-based strategy, all records from the most recent year are counted at the maximum weight scale, whereas the index-based strategy only counts one record at the maximum.
 While the two weighting strategies resulted in identical results in this scenario, the year-based strategy is preferred as it encourages an increase in support weight closely coupled to real-world historical trends.
@@
 
-Figure (7) shows the correctness ratios generated by all possibly training data and testing data pairings.
+Figure (7) shows the correctness ratios generated by all possible training data and testing data pairings.
 Note that ratios in brackets denote the worse-case ratio while ratios outside of brackets refer to the best-case.
@@
 Regardless of correctness ratio, 2001-2015 should always be selected as the test data interval to encourage analysis against more recent hurricane patterns.
-The results show that no training data-testing data pairing was able to generate correctness ratios higher than the ones generated by training the model with data from 1950 to 2000, and testing it on data as recently recorded as 2015.
+The results show that no training-data, testing-data pairing was able to generate correctness ratios higher than the ones generated by training the model with data from 1950 to 2000, and testing it on data as recently recorded as 2015.
 
@@
 The three contributions combined form WARD-HTP, and produced a hurricane trajectory prediction correctness ratio of exactly 75.0\%, which is 10\% higher than the current state-of-the-art.
-When account for trajectories which were not able to match with any association rules, the correctness ratio of WARD-HTP is 72.58\%, which is still 5.08\% higher than the worst case state-of-the-art ratio of 57.5\% as seen in \cite{major-hurricane-model}
+When accounting for trajectories which were not able to match with any association rules, the correctness ratio of WARD-HTP is 72.58\%, which is still 5.08\% higher than the worst case state-of-the-art ratio of 57.5\% as seen in \cite{major-hurricane-model}.
 
-There are mutliple areas to explore in the hurricane trajectory prediction problem space with respect to future work.
+There are multiple areas to explore in the hurricane trajectory prediction problem space with respect to future work.
 Experiments in this work show that all datasets contain a subset of trajectories for which no association rule could be matched.
@@
 Additionally, more experiments will be run on different inputs including training data from other regions such as the Pacific region.
-Further approaches to increasing the correctness ratio of WARD-HTP will also be investiaged.
+Further approaches to increasing the correctness ratio of WARD-HTP will also be investigated.
 This includes the investigation of alternative methods for defining training data weight.
@@
 \bibitem{hurricane-distances} D. Brown. (2013) Forecast Verification: Quantifying Forecast Uncertainty. National Hurricane Conference: Hurricane Readiness for Coastal Communities
-\bibitem{earth-radius-kilometers} Chambat, F., and Valette, B. (2001). Mean radius, mass, and inertia for reference Earth models. Physics of the Earth and Planetary Interiors, 124(3), 237-253.
+\bibitem{earth-radius-kilometres} Chambat, F., and Valette, B. (2001). Mean radius, mass, and inertia for reference Earth models. Physics of the Earth and Planetary Interiors, 124(3), 237-253.
 \bibitem{hurricane-recording-history} Landsea, C.

### 2017-12-09 COMP4710-Bissonnette-Marofke-Cox-Project-Fall2017/paper/comp4710-bissonnette-marofke-cox-manuscript.tex
Polish repo for final project handin

--- 2017-12-08 paper/manuscript.tex
@@
 In this work, we propose multiple dimensions for improving the accuracy of data mining-based hurricane trajectory prediction.
-We propose the use of Haversine distance, a real-world distance formula for reliable hurricane pattern matching accuracy.
-We also propose the use of weighted data to favour recent hurricane trends when training the trajectory prediction model.
-Finally, we use training data exclusively from the years 1950 to 2000 to reduce the impact of historical inaccuracies in hurricane recording methodologies.
+We propose the use of the Haversine distance formula for realistic hurricane trajectory distance comparison.
+We also propose the use of weighted training data to favour recent hurricane trends when determining a trajectory prediction.
+Finally, we use training data exclusively from the years 1950 to 2000 to reduce the impact of historical climate trends and inaccuracies in hurricane recording methodologies.
 After implementing all three improvement dimensions on a contemporary data mining model for hurricane trajectory prediction based on AprioriAll, results show a prediction-correctness ratio up to 10.0\% higher than the current state-of-the-art, at 75.0\%.
@@
 Hurricanes are divided into five categories according the Saffir-Simpson scale, based on wind-speed and expected damage.
-In the United States alone, hurricanes which can incur \$157BN in damages \cite{hurricane-cost}.
+In the United States alone, the damages incurred by a single hurricane have reached as high \$157BN \cite{hurricane-cost}.
 The damages incurred by hurricanes do not only include property damage, but also personal danger and in some cases widespread civil unrest \cite{hurricane-cost-non-financial}.
@@
 These association rules are used to make predictions on future data entries such as future customer transactions.
-More specifically, association rules generated by sequential data mining algorithms including AprioriAll, SPIRIT and PrefixSpan can be used in the field of hurricane trajectory prediction.
+More specifically, association rules generated by sequential data mining algorithms including AprioriAll, SPIRIT and PrefixSpan can be used as training data for a hurricane trajectory prediction model.
 
-To the knowledge of the authors, the existing standard in hurricane trajectory prediction via data mining is developed by Dong et al in \cite{major-hurricane-model}.
+To the knowledge of the authors, the existing standard in hurricane trajectory prediction via data mining is established by Dong et al in \cite{major-hurricane-model}.
 The work cited proposes a hurricane trajectory prediction model based on data mining (HTPDM) that uses a modified variant of AprioriAll, a common sequential mining algorithm \cite{AprioriAll-original}.
@@
 The results of HTPM showed that the model was able to achieve a best-case correctness ratio of 65.0\%.
-In this work, HTPDM will be developed and various strategies will be employed to increase the correctness ratio.
-Reducing the impact of historical factors, favouring the impact of recent trends, and improving the realistic nature of the hurricane trajectory prediction model result in a new model dubbed WARD-HTP (Weighted-Asset, Realistic-Distance Hurricane Trajectory Predictor).
+In this work, a model similar to HTPDM will be implemented and various strategies will be employed to increase the correctness ratio.
+Reducing the impact of historical factors, favouring the impact of recent trends, and improving the realism of the hurricane trajectory prediction model result in a new model dubbed WARD-HTP (Weighted-Asset, Realistic-Distance Hurricane Trajectory Predictor).
 The strategies employed in WARD-HTP ultimately result in a correctness ratio of 75.0\%, a 10\% improvement over HTPDM.
@@
 Section VII will evaluate the results of the experiments, which show that the ideal configurations of WARD-HTP result in a maximal correctness ratio of 75.0\%.
-This work will then conclude with section VIII, capturing the conclusions and directions of future work.
+Finally, section VIII will capture all conclusions and proposed directions for future work.
 
@@
 While the described work provided novel results in typhoon track clustering, it is not clear if the proposed solution is able to predict the trajectories of new storms.
-Additionally, the clusters developed apply specifically to typhoons, which are waterborne storms native to the east-pacific seaboard.
-Such storm patterns may not be applicable to patterns found in hurricanes, which are waterborne storms of the west Atlantic basin.
-For this reason, the proposed clustering patterns were not used in this work.
+Additionally, the clusters developed apply specifically to typhoons, which are tropical storms native to the east-pacific seaboard.
+Such storm patterns may not be applicable to patterns found in hurricanes, which are exclusive to the Atlantic basin.
+For these reasons, the proposed clustering patterns were not used in this work.
 
@@
 These trajectory patterns correspond to interesting trajectory association rules, which are used to train a prediction model.
-The implementation of AprioriAll used in \cite{major-hurricane-model} is also used as the foundational model for this work.
+The implementation of AprioriAll used in \cite{major-hurricane-model} also serves as the foundation for the model proposed in this work.
 AprioriAll transforms a transaction database into a \textit{customer} database, where each entry contains a customer and an ordered sequence of purchased items.
 The ordered sequence database is then mined for frequent ordered patterns, or frequent sequences.
-The modification of Apriori proposed in \cite{major-hurricane-model} mines frequent ordered patterns and thus skips the transformation step of AprioriAll.
+The modification proposed in \cite{major-hurricane-model} directly mines frequent ordered patterns and thus skips the transformation step of AprioriAll.
 
-The model proposed in this work uses historical hurricane trajectory readings from HURDAT, the canonical hurricane database maintained by the National Hurricane Center \cite{HURDAT2-original}.
-The data in HURDAT2 is ordered into $\langle Hurricane, Trajectory \rangle$ pairs, omitting the need for the sequence mapping step required at initialization of AprioriAll.
-The Atlantic hurricane trajectories recorded in HURDAT2 span a very large distance, some of which reaching over 10,000 kilometres in total distance travelled.
+The model proposed in this work uses historical hurricane trajectory readings from HURDAT2, the canonical hurricane database maintained by the National Hurricane Center \cite{HURDAT2-original}.
+The data in HURDAT2 was sorted into $\langle Hurricane, Trajectory \rangle$ pairs, removing the need for the sequence mapping step required at initialization of AprioriAll.
+The Atlantic hurricane trajectories recorded in HURDAT2 span a very large area, some of which reaching over 10,000 kilometres between the first and last recorded points.
 As a result, the coordinates recorded are limited to those within a range of 0 to 50 degrees north, and 20 to 100 degrees west.
 This degree range is approximately the size of the south Atlantic basin.
-
-% We might want some more related work
 
@@
 There are two categories of motivation behind this work.
-The first motivation of this work is to increase the total investigation completed in the hurricane trajectory prediction sub-discipline.
+The first motivation of this work is to add to the total investigation completed in the data mining-based hurricane trajectory prediction subdiscipline.
 Little research effort has been allocated to this particular problem space of high economic and social impact.
-The authors expect this work to contribute to the growing effort by academic and government researchers to develop reliable methods of hurricane trajectory prediction in the interest of public safety and economic stability.
+This work is expected to contribute to the growing effort by academic and government researchers to develop reliable methods of hurricane trajectory prediction in the interest of public safety and economic stability.
 
@@
 This means that of all testing trajectory for which a matching rule was successfully produced, 65.0\% of such matches were correct predictions.
-When HTPDM accounted for testing trajectories for which no rule could be matched, the correctness ratio decreased by 7.5\% to a final value of 57.5\% correctness.
+When HTPDM accounted for testing trajectories with which no rule could be matched, the correctness ratio decreased by 7.5\% to a worst-case value of 57.5\% correctness.
 This decrease in accuracy is expected as some hurricane trajectories are outliers which cannot be predicted by frequent pattern mining, as will be discussed further in this work.
@@
 The motivation of improving the correctness ratio in the model encompasses the challenges of improving the best-case correctness ratio (the correctness ratio when ignoring rule-miss incidents) and the worst-case correctness ratio (the correctness ratio when accounting for rule-miss incidents).
-To resolve the motivations of this work, a series of improvement strategies are proposed as applications to the current hurricane trajectory prediction model based on AprioriAll.
+To fulfill the motivations of this work, a series of improvement strategies are proposed for the current standard hurricane trajectory prediction model based on AprioriAll.
 In general, these strategies seek to improve the realism of the prediction model, reduce historical factors which may negatively affect the model's correctness ratio, and favour recent hurricane trajectory trends which are likely to be reflected in current and future hurricane patterns.
@@
 
-% What are the things to know about the problem?
-% - Algorithm: AprioriAll ( pseudocode )
-% - hit rate/miss rate
-% - regions ( show diagram )
-% - define accuracy, rule fit, fitness function.. ( equations )
-
-In this section, the problem of achieving accurate hurricane trajectory prediction is explained in detail.
-This problem consists of multiple challenges which apply generally to sequential data mining and specifically to the problem domain of hurricane location analysis.
-Many of the domain-specific challenges of hurricane trajectory prediction arise from the fact that hurricane points are tracked as a series of $\langle Latitude, Longitude \rangle$ points, which do \textit{not} directly correspond to traditional (x, y) coordinates, as latitude and longitude are points imposed upon a sphere.
-Challenges including accurate distance evaluation are results of the transition from a Euclidean to a Polar coordinate system.
+In this section, the problems surrounding the achievement of accurate hurricane trajectory prediction are explained in detail.
+These problems consist of multiple challenges which apply generally to sequential data mining and specifically to hurricane distance, trajectory and location analysis.
+Many of the domain-specific challenges of hurricane trajectory prediction arise from the fact that hurricane points are tracked as a series of $\langle Latitude, Longitude \rangle$ points, which do \textit{not} directly correspond to traditional (x, y) coordinates, as latitude and longitude coordinates are points imposed upon a sphere.
+Challenges such as accurate distance evaluation are the result of the need to transition from a Euclidean to a Polar coordinate system.
 
@@
 Next, the subproblem of region discretization will be discussed.
-Since $\langle Latitude, Longitude \rangle$ points constitute continuous space, hurricane trajectories are converted to sequences of discrete regions to enable the discovery of frequent patterns.
+Since $\langle Latitude, Longitude \rangle$ points describe continuous space, hurricane trajectories are converted to sequences of discrete regions to enable the discovery of frequent patterns.
 Finally, problems involving hurricane trajectory matching will be discussed.
@@
 AprioriAll was first proposed at IBM research as a method for discovering frequent sequential patterns in transactional databases \cite{AprioriAll-original}.
-In its original implementation, AprioriAll is able to generate association rules describing sentences of the form ``Customers who buy P may also buy Q''.
+In its original implementation, AprioriAll is able to generate association rules describing sentences of the form ``Customers who buy P may later buy Q''.
 In the hurricane trajectory prediction problem, AprioriAll is used to generate association rules describing sentences of the form ``Hurricanes located at P may later be located at Q''.
@@
 
-In step (1) of AprioriAll, a given transaction database \textit{D} is grouped by customers, and the frequent items purchased by customers in individual transactions are recorded in a temporary table.
+In step (1) of AprioriAll, a given transaction database \textit{D} is grouped by its customers, and the frequent items purchased by customers in individual transactions are recorded in a temporary table.
 The contents of this temporary table are mapped to corresponding sequences to enable the mining of frequent transaction sequences from the original database.
@@
 
-Section (2) of AprioriAll initiates the generation of frequent sequences.
-In the case of hurricane trajectory prediction, these sequences correspond to frequent recorded hurricane coordinate sequences.
+Section (2) of AprioriAll is responsible for the generation of frequent sequences.
+In the case of hurricane trajectory prediction, these sequences correspond to frequently recorded hurricane coordinate sequences.
 Frequent hurricane coordinate sequences are those which occur at least a minimum number of times, as defined by the \textit{minimum support}, or minsup.
@@
 Once the sequence generation level reaches three, each candidate sequence must be pruned if any of their candidate subsequences are infrequent.
-Continuous execution of AprioriAll until completion results in a collection of frequent hurricane trajectory patterns, as defined by the minimum support.
+Continuous execution of AprioriAll until termination results in the collection of all frequent hurricane trajectory patterns defined by the minimum support.
 
@@
 \begin{algorithm}[H]
-  \caption{Modified AprioriAll: No L-Itemset phase}
+  \caption{Modified AprioriAll: No Mapping Phase}
   \label{aprioriall_for_hurricanes}
@@
 
-Along with HTPDM, WARD-HTP executes AprioriAll with a slight modification to the existing algorithm proposed in \cite{AprioriAll-original}.
-Whereas L1, the frequent single-item sequences is generated by the L-Itemset (mapping) phase of AprioriAll, the variant of AprioriAll used in hurricane trajectory prediction skips the L-Itemset phase, and thus scans the $\langle HurricaneID, CoordinateSequence\rangle$ directly for frequent single-point trajectories.
+As also seen in HTPDM, WARD-HTP executes AprioriAll with a slight modification to the existing algorithm proposed in \cite{AprioriAll-original}.
+Whereas L1 (the set of frequent single-item sequences) is generated by the L-Itemset (mapping) phase of AprioriAll, the variant of AprioriAll used in hurricane trajectory prediction skips the L-Itemset phase and thus scans the $\langle HurricaneID, CoordinateSequence\rangle$ database directly for frequent single-point trajectories.
 Otherwise, the remainder of AprioriAll is used in this work as originally designed.
@@
 If the matching condition is satisfied, a new trajectory is generated.
-The inner-join step of AprioriAll emulates the required hurricane trajectory-matching to develop frequent hurricane trajectory patterns.
+The inner-join step of AprioriAll emulates the required hurricane trajectory-matching to discover frequent hurricane trajectory patterns.
 Once all possible frequent hurricane trajectory patterns are discovered, the second step of AprioriAll completes.
@@
 
-The final step of AprioriAll is the rule-generation step.
+The third and final step of AprioriAll is the rule-generation step.
 Though association rule generation is not a specific stage of the AprioriAll algorithm, it is traditionally the final stage of any association rule mining approach.
-The rule generation step is responsible for developing interesting association rules.
+The rule generation step is responsible for identifying and recording interesting association rules.
 As mentioned, rules generated by the hurricane trajectory prediction model proposed will satisfy sentences of the form ``Hurricanes located at P may later be located at Q''.
@@
 The left partition of \textit{S} constitutes the antecedent of \textit{R}, and the latter partition is the consequent.
-The support of \textit{R} is the original support of \textit{S}.
+The support of \textit{R} is the support of the unpartitioned pattern \textit{S}.
 The confidence of \textit{R} is the support of \textit{S} divided by the support of the subsequence given by the former partition.
@@
 Without the association rules formed by AprioriAll and the rule generation step, data mining-based hurricane trajectory prediction cannot be executed.
-HTPDM and WARD-HTP both use historically frequent hurricane trajectory to predict the result of future hurricane paths.
+HTPDM and WARD-HTP both use historically frequent hurricane trajectories to predict the result of future hurricane paths.
 
@@
 \begin{figure}[htp]
-\caption{Sample Hurricane Trajectory Paths}
+\caption{Hurricane Trajectory Paths Without and With Regionalization}
 \centering
@@
 The final key aspect of the hurricane trajectory prediction problem concerns accurate hurricane trajectory pattern matching.
-Pattern matching is a concern for hurricane trajectory prediction in two primary areas.
+Pattern matching is a concern for hurricane trajectory prediction in two key areas.
 First, hurricane trajectory matching must be effectively employed when comparing a new hurricane trajectory against the table of hurricane association rules generated by the previously described variant of AprioriAll.
@@
 Where r is a member of the set of all interesting association rules \textit{R}, t is the initial trajectory $T_{i}$, and $matchLen_{t,r}$ expresses the matching length between $T_{i}$ and the antecedent sequence of r.
-This formula is meant to be maximized to determine the rule which most generously fits the incoming hurricane trajectory.
-Once a best fit rule is determined, the rule consequent is compared with the \textit{actual} terminal hurricane trajectory to determine if the prediction is correct.
+The match formula is meant to be maximized in order to determine the rule which most generously fits the incoming hurricane trajectory.
+Once the rule of best fit is determined, the rule's consequent is compared with the \textit{actual} hurricane terminal trajectory to determine if the prediction is correct.
 
@@
 The pattern matching for rule consequents has some differences compared to antecedent pattern matching.
-First, prediction correctness uses matching length with a different definition.
+First, the definition of matching length is altered when considering prediction correctness.
 Whereas matching length for rule antecedents is the longest identical congruent subsequence of the initial trajectory, matching length for the rule consequent is the longest congruent subsequence of points within a specific distance, beginning at the start of the terminal sequence.
-This methodology is also based on HTPDM and is based on the following steps.
-In the algorithm given, $conseq_{r}$ corresponds to the matched rule's consequent sequence, and D corresponds to the minimum distance threshold necessary for prediction correctness.
+This methodology is also based on HTPDM and is based on the following steps shown in algorithm (2).
+In the algorithm given, $conseq_{r}$ corresponds to the matched rule's consequent sequence and D corresponds to the minimum distance threshold necessary for prediction correctness.
 In HTPDM, the distance threshold corresponds to a Euclidean distance of 1.
@@
 
-% Model formal explanation could be better. Maybe do pseudocode
-
 \begin{enumerate}
@@
 If no such match is found, HTPDM attempts to extrapolate a future point of the initial trajectory, using the extrapolated point as a new option for determining a rule antecedent match.
-When accounting for all tested trajectories, including those which were matched to a rule as a result of trajectory extrapolation, the correctness ratio was 57.5\%.
+When accounting for all tested trajectories, including those which were matched to a rule as a result of trajectory extrapolation or never matched with any rules, the correctness ratio was 57.5\%.
 Therefore, the best case correctness ratio of HTPDM is the ratio which ignores all test trajectories which did not match with any rules, resulting in a final value of 65.0\%.
@@
 Despite establishing a convincing initial start point, the best-case correctness ratio of 65.0\% may be too low for HTPDM to be considered reliable in the real-world.
-This is problematic for those interested in hurricane trajectory prediction as a model with only 65.0\% accuracy may not be able to deliver crucial insight into hurricane trajectory prediction, particularly with respect to landfall location prediction.
+This is problematic for those who advocate the use of data mining in hurricane trajectory prediction as a model with only 65.0\% accuracy may not be able to deliver crucial hurricane trajectory prediction insights, particularly with respect to landfall location prediction.
 Both correctness ratios given must be increased in order for data mining to be considered a practical approach to predicting the trajectories of future hurricanes.
@@
 In order to increase the correctness ratio of data mining-based hurricane trajectory prediction, this work identifies and explores three opportunities for worst-case and best-case ratio refinement.
-The opportunities identified are the introduction of more realism in calculation formulas, reduction of historical impact on data used, and favourability towards recent hurricane trends.
+The opportunities identified are the introduction of increased realism in distance calculation, the reduction of historical impact on data used, and the favourability towards recent hurricane trajectory trends.
 The opportunities for improvement are realized by three unique dimensions.
@@
 The implementation of these three dimensions results in a new hurricane trajectory prediction model dubbed WARD-HTP (Weighted-Asset, Realistic-Distance Hurricane Trajectory Predictor).
-The combined contributions of WARD-HTP are expected to result in the desired increase to the best-case and worst-case correctness ratios.
+As shown further in this work, the combined contributions of WARD-HTP result in the desired increase to the best-case and worst-case model correctness ratios.
 
@@
 
-In contrast to the Euclidean formula, the Haversine Distance Formula is designed to determine the kilometres between two latitude-longitude points over a large sphere, such as the earth's surface.
+In contrast to the Euclidean formula, the Haversine Distance Formula is designed to determine the distance in kilometres between two latitude-longitude points over a large sphere, such as the earth's surface.
 Despite the fact that the earth's surface is not a perfect sphere \cite{lat-long-distance}, the Haversine Distance Formula is sufficiently accurate when used within a specific region such as the Atlantic Basin.
@@
 & \alpha = \sin^2(\frac{\Delta Lat}2) + \cos(p_{lat}) + \cos(q_{lat}) + \sin^2(\frac{\Delta Lon}2)
-% & \Delta Lat = (q_{lat} - p_{lat})/2\\
-% & \Delta Lon = (q_{lon} - p_{lon})/2
 \end{align*}
+
 % H is in RADS, Eucl in degrees
@@
 Note that the Euclidean distance formula expresses Latitude and Longitude in degrees and minutes, while the Haversine formula expresses Latitude and Longitude exclusively in radians.
-In addition, whereas the Euclidean formula expresses distance in terms of units on a plane, the Haversine formula expresses distance in terms of portions of a 1-unit sphere.
+In addition, whereas the Euclidean formula expresses distance in terms of units on a plane, the Haversine formula expresses distance in terms of portions of a single unit sphere.
 Therefore, to express distance in kilometres, the result of Haversine formula must be multiplied by R, which denotes the radius of the earth in kilometres (approximately 6371km \cite{earth-radius-kilometres}).
@@
 While the weighting of training data results in some trajectories being counted with a support greater than 1, this effect is balanced by counting earlier recorded trajectories with a support less than 1.
-In WARD-HTP, we weigh hurricane trajectories on a scale increasing linearly, resulting in a tipping point \textit{T} where records before \textit{T} are underweighted and records after \textit{T} are overweighted.
-Depending on the weighing approach used, the weights of all records does not necessarily need to be balanced.
+In WARD-HTP, we weigh hurricane trajectories on a scale increasing linearly.
+This results in a tipping point \textit{T} where records before \textit{T} are underweighted and records after \textit{T} are overweighted.
+Depending on the weighing approach used, the combined weight of all records does not necessarily need to be balanced.
 
@@
 
-% give weighted algorithm and explain what + why
-%Weighted AprioriAll Pseudocode
-% Modify this for weighted stuff
 \begin{algorithm}[H]
@@
 In its original form, HTPDM is trained on hurricane trajectory data from 1900 to 2000.
-However, incorporating training data from years as early as 1900 introduces a risk of skewing frequent trajectory patterns in favour of historical errors and anomalies.
-While weighted support in AprioriAll helps reduce the impact of historical anomalies, the primitivity of early hurricane recording technology makes hurricane trajectories recorded before 1950 much less reliable than those recorded after \cite{hurricane-recording-history}
+However, incorporating training data from years as early as 1900 introduces a risk of skewing frequent trajectory patterns in favour of historical recording errors and climatological anomalies.
+While weighted support in AprioriAll helps reduce the impact of historical factors, the primitivity of early hurricane recording technology makes hurricane trajectories recorded before 1950 much less reliable than those recorded after \cite{hurricane-recording-history}.
 
@@
 
-% careful about colour: what if we publish in black and white
 Figure (3) illustrates the characteristics of trajectories which were not able to match with any interesting rules.
@@
 Some of the non-matching trajectories appear to align with training trajectories, but are moving in an opposite direction.
-Therefore, any points that match would also not provide a reliable result.
+Therefore, any points that matched such trajectories would also not provide a reliable result.
 
@@
 This section documents the experimental results of all improvement dimensions proposed in WARD-HTP.
-Each dimension of Haversine Distance, Weighted Training Data and Training Data Year Interval of 1950-2000 contributes iteratively to a total growth in correctness ratio of 10.0\% when compared to HTPDM, for a total of 75.0\%.
+The respective dimensions of Haversine Distance, Weighted Training Data and Training Data Year Interval of 1950-2000 contribute iteratively to a total growth in best-case correctness ratio of 10.0\% when compared to HTPDM, for a total of 75.0\%.
 Whereas the worst-case correctness ratio (ratio accounting for rule-miss incidents) of HTPDM was 57.5\%, the worst-case correctness ratio of WARD-HTP was 72.58\% when configured in the same manner that generated the best-case result of 75.0\%.
@@
 This results in a difference decrease of 5.08\%.
-Not only is the difference in ratios reduced when using WARD-HTP, the best and worse case ratios are also increased, by 10.0\% and 15.08\% respectively.
+Not only is the difference in ratios reduced when using WARD-HTP, the best and worse case ratios are also increased, by 10.0\% and 5.08\% respectively.
 
@@
 The first approach imposes a distance threshold \textit{D} of 300km while the second sets \textit{D} to 450km.
-300km was selected as a sample value for testing the Haversine formula, while 450km was selected as a value slightly less than the approximate average hurricane diameter discussed in section 4-B.
+300km was selected as a sample value for testing the Haversine formula, while 450km was selected as a value slightly less than the approximate average hurricane diameter as discussed in section 4-B.
 
@@
 As shown in figure (4), the Euclidean distance option resulted in much lower correctness ratios than both the 300km and 450km Haversine approaches.
-However, the difference between best and worst case correctness ratios is lowest when using Euclidean distance.
-Since increasing the ratios is prioritized over reducing the ratio difference, the Haversine approach of 450km produces the best result.
+The Haversine Distance approach with \textit{D} set to 450km produces the highest best and worse-case correctness ratios.
+While the difference between ratios is smallest when using the Euclidean distance formula, the overall increase in both ratios makes the Haversine distance at \textit{D}$=$450km the optimal selection.
 The distance threshold of 450km introduces more flexibility into which predictions count as correct.
@@
 
-In this section of the evaluation against WARD-HTP, multiple full executions of the model from AprioriAll to the correctness ratio analysis are completed.
-Each execution of WARD-HTP is performed on a unique pairing of minimum support and the training data weighing strategy with the weighing scale W set to 1 or 2.5.
-Apart from the minimum support and the weighing strategy, all other parameters of the model are held constant.
+In this section of the evaluation against WARD-HTP, multiple complete executions of the model, from AprioriAll to the correctness ratio analysis, are completed.
+Each execution of WARD-HTP is performed on a unique pairing of minimum support and training data weighting strategy with the weighing scale W set to 2.5.
+The weighting strategy options also includes the absence of any weighting strategy.
+Apart from the minimum support and the weighting strategy, all other parameters of the model are held constant.
 
-Three different weighing strategies are evaluated in this experiment.
-First, the absence of data weighing is explored (W$=$1).
-Next, weighing on an annual scale is considered.
+Three different weighting strategies are evaluated in this experiment.
+First, the absence of data weighing is explored.
+Next, weighting on an annual scale is considered.
 In this strategy, the value of w increases by W divided by the total number of years in the data Y (in the case where training data is supplied from 1950 to 2000, Y$=$50).
-The third strategy employed is the index-weight strategy, where the weight is updated at each HDB entry as seen in section 5-B.
+The third strategy employed is the index-weight strategy, where the weight is updated at each HDB entry.
 
@@
 Minimum support values corresponding to 2.5\%, 3.75\% and 5\% of the number of training data trajectories are paired against the three weighing strategies discussed above.
-While a higher minimum support will generate fewer interesting trajectory association rules, the rules generated will be more reliable for future predictions.
+While a higher minimum support will generate fewer interesting trajectory association rules, the rules generated may be more reliable for future predictions.
 
@@
 
-The maximal best-case hit rate for WARD-HTP was generated with an index-weighted strategy with minimum support of 2.5\%, corresponding to 18.875 when used with data from 1950 to 2000.
+The maximal best-case hit rate for WARD-HTP was generated with a year-weighted strategy and minimum support set to 2.5\%, corresponding to 18.875 when used with data from 1950 to 2000.
 The resulting hit rate was 75.0\%, with a worst-case of 72.58\%.
-Note that a minimum support of 5\% with index-based weighting generated a very slightly higher best-case hit rate at 75.22\%, albeit with a lower worst-case ratio of 67.33\%.
+Note that a minimum support of 5\% with index-based weighting generated a very slightly higher best-case hit rate at 75.22\%, though with a lower worst-case ratio of 67.33\%.
 We select 75\% as the optimal generated data point because a .22\% decrease in best-case hit rate is certainly a ``fair trade'' for a 5.25\% increase in worst-case hit rate.
@@
 In the year-based strategy, all records from the most recent year are counted at the maximum weight scale, whereas the index-based strategy only counts one record at the maximum.
-While the two weighting strategies resulted in identical results in this scenario, the year-based strategy is preferred as it encourages an increase in support weight closely coupled to real-world historical trends.
+While the two weighting strategies resulted in identical results in this scenario, the year-based strategy is preferred as it closely couples the increase in support weight to real-world historical hurricane trajectory trends.
 
-% talk a little more then done
 \begin{figure*}[!tp]
@@
 
-% some training data notes
 The final component of the evaluation undertaken in this work compares the correctness ratios generated by various training and testing data intervals.
@@
 
-% figure
 \begin{figure}[htp]
@@
 \end{figure}
-% explain. done.
 
@@
 Regardless of correctness ratio, 2001-2015 should always be selected as the test data interval to encourage analysis against more recent hurricane patterns.
-The results show that no training-data, testing-data pairing was able to generate correctness ratios higher than the ones generated by training the model with data from 1950 to 2000, and testing it on data as recently recorded as 2015.
+The results show that no training-data/testing-data pairing was able to generate correctness ratios higher than the ones generated by training the model with data from 1950 to 2000, and testing it on data as recently recorded as 2015.
 
@@
 These contributions are the use of training data from 1950 to 2000, the use of weighted training data, or weighted assets (WA), and the use of the Haversine Distance Formula for realistic distance (RD).
-The three contributions combined form WARD-HTP, and produced a hurricane trajectory prediction correctness ratio of exactly 75.0\%, which is 10\% higher than the current state-of-the-art.
+The three contributions combined form WARD-HTP and produce a hurricane trajectory prediction correctness ratio of exactly 75.0\%, which is 10\% higher than the current state-of-the-art.
 When accounting for trajectories which were not able to match with any association rules, the correctness ratio of WARD-HTP is 72.58\%, which is still 5.08\% higher than the worst case state-of-the-art ratio of 57.5\% as seen in \cite{major-hurricane-model}.
@@
 \end{document}
-
-% there will be lots of editing %
-
-% Early callouts:
-% Motivation could be more clear
-% We need to make sure we are explaining rule-miss incidents and our approach to them
-% We also need to make sure we explain what exactly the correctness ratio is and how it changes based on accounting for rule-miss
-% Also need to explicitly call out what our minconf was and why
-% future work: non-binary prediction correctness?
-% More non-text assets?
-% We have done a good job describing the problem space but we need to explain what we problems __are__
-% ie: why is there a solution
-% we will need to re-read the paper and keep an eye out for sections that suffer from scope creep. Problem description first comes to mind.
-
-% we need to make sure we are not jumping to far off the deep end with hurricanes and forgetting to discuss d a t a  m i n i n g

### 2017-12-09 paper/manuscript.tex
Polish repo for final project handin

@@
 In the year-based strategy, all records from the most recent year are counted at the maximum weight scale, whereas the index-based strategy only counts one record at the maximum.
-While the two weighting strategies resulted in identical results in this scenario, the year-based strategy is preferred as it encourages an increase in support weight closely coupled to real-world historical trends.
+While the two weighting strategies resulted in identical results in this scenario, the year-based strategy is preferred as it closely couples the increase in support weight to real-world historical hurricane trajectory trends.
 

### 2017-12-09 COMP4710-Bissonnette-Marofke-Cox-Project-Fall2017/run_experiment.sh

@@ -1,0 +1,6 @@
+#!/bin/sh	
+python scripts/generate_frequent_patterns.py data/vectors_training_1950_2000.txt output/patterns.txt 18.875 1 0
+
+python scripts/generate_assoc_rules.py output/patterns.txt output/rules.txt 0.25
+
+python scripts/test_best_fit.py data/vectors_testing_2001_2015.txt output/rules.txt 1 > output_result.txt

### 2017-12-09 COMP4710-Bissonnette-Marofke-Cox-Project-Fall2017/scripts/generate_assoc_rules.py

@@ -1,0 +1,112 @@
+# Generates association rules given a list of frequent itemsets
+# It is assumed that each line in the file is of the form (IS) : Supp
+# Where IS is a ", " delimited series of numbers, which correspond to regions
+# Ignore singletons, they cannot produce rules
+
+import sys
+import time
+import re
+
+def getRawRules(itemset):
+  divider = len(itemset) - 1
+  rules = [] # Antecedent, Consequent pairs
+  while divider > 0:
+    antecedent = itemset[0:divider]
+    consequent = itemset[divider:]
+    rules.append( (antecedent, consequent) )
+    divider -= 1
+
+  return rules
+
+def convertToSearchLine(antecedentArray):
+  if len(antecedentArray) == 1:
+    line = antecedentArray[0]
+  else:
+    line = "(" + ", ".join(map(str, antecedentArray)) + ")"
+
+  # Will have to change this to parse strings correctly
+  return str(line).replace("'", "")
+
+def getAntecedentSupport(itemsetsFile, searchLine):
+  with open(itemsetsFile, "r") as itemsets:
+    for line in itemsets:
+      contents = line.split(":")
+      if contents[0].strip() == searchLine:
+        result = float(contents[1].strip())
+
+  return result
+
+def generateRules(itemsetsFile, currItemsetLine, minconf=0.25): # String line of form (t, t, t) : s
+  generatedRules = []
+
+  currItemsetLine = currItemsetLine.replace(" ", "")
+  currItemsetParsed = currItemsetLine.split(":")
+
+  currItemsetSupport = float(currItemsetParsed[1])
+
+  # Change this to list of tuples
+  currItemset = convertToTupleList(currItemsetParsed[0])
+  # Now we have the full itemset, and its support
+  # These are ordered rules, so we grow the antecedent and shrink the consequent together
+  # ABC, AB>C, A>BC
+  rulePartitions = getRawRules(currItemset)
+
+  # To find confidence, we need support of antecedent
+  for rule in rulePartitions:
+    searchLine = convertToSearchLine(rule[0])
+    # Do a search in the file for the corresponding frequent itemset, and get its support
+    # That gives the confidence of the rule to be inserted
+    antecedentSupport = getAntecedentSupport(itemsetsFile, searchLine)
+
+    conf = float(currItemsetSupport) / float(antecedentSupport)
[…]

### 2017-12-09 COMP4710-Bissonnette-Marofke-Cox-Project-Fall2017/scripts/generate_frequent_patterns.py

@@ -1,0 +1,171 @@
+
+from apriori_utils import *
+import sys
+import itertools
+import time
+
+WEIGHT_INTERVAL_NOMINAL = 756
+WEIGHT_INTERVAL_YEAR = 50
+HURRICANE_ID_PLACEHOLDER = "AL015555"
+WEIGHT_SCALE = 2.5
+# For weighted data:
+# Data can be weighted nominally or year-over-year
+# These values are dependent on the specific training data input. Specifically, change the YEAR interval and NOMINAL interval as appropriate for different inputs
+
+def gather_new_frequent_items(resultTable, frequent_k_itemsets, minsup):
+  resultTable.append({})
+  for key in frequent_k_itemsets:
+    if frequent_k_itemsets[key] >= minsup:
+      resultTable[-1][key] = frequent_k_itemsets[key] 
+
+def timeseries_candidate_generation(frequent_k_itemsets, chunk_size):
+  # this is the candidate generation and pruning
+  # We have our L_k, use suffix-prefix to get C_k+1
+
+  frequent_k_itemsets_dict = { tuple(item):None for item in frequent_k_itemsets }
+  candidates = []
+  for curr_index, curr_item in enumerate(frequent_k_itemsets):
+
+    compare_index = 0
+    found = False
+    leftChunk = curr_item[(chunk_size*-1):]
+
+    while compare_index < len(frequent_k_itemsets) and not found:
+      found = leftChunk == frequent_k_itemsets[compare_index][:chunk_size]
+      if not found:
+        compare_index += 1
+
+    while compare_index < len(frequent_k_itemsets) and leftChunk == frequent_k_itemsets[compare_index][:chunk_size]:
+      new_candidate = join_entries(curr_item, frequent_k_itemsets[compare_index], chunk_size)
+      required_itemset_checks = itertools.combinations(new_candidate, chunk_size+1)
+
+      if all(item == curr_item or item == frequent_k_itemsets[compare_index] or tuple(item) in frequent_k_itemsets_dict for item in required_itemset_checks):
+        candidates.append(new_candidate)
+      compare_index += 1
+
+  return candidates
+
+def join_entries(left, right, chunk_size):
+  result = left + right[chunk_size:]
+  return result
+
+def gather_sublists(region_list):
+  result = []
+  for size in range(1, len(region_list)):
+    for index in range(len(region_list)):
+      result.append(region_list[index : index + size])
+
+  return result
+
[…]

### 2017-12-09 COMP4710-Bissonnette-Marofke-Cox-Project-Fall2017/scripts/test_best_fit.py

@@ -1,0 +1,391 @@
+# Given an incoming trajectory:
+# Cut it into two pieces: check, confirm
+# Check "check" against rule antecedents, get best fit (or show "missing")
+# if not missing, see if "confirm" was right
+
+# To do this, we will need to cut the patterns into two pieces. Training and Testing.
+# First try: quarters. 1335 training 445 testing
+
+# First try: divide test line in half. First half looks for antecedent
+
+# Parse a line into the antecedent, list of strings
+
+import math
+import sys
+import time
+import re
+from apriori_utils import *
+
+EXTRAPOLATION_LIMIT = 10
+EARTH_RADIUS = 6371
+MAX_MINIMUM_MATCH = 5 # Max comparison length from paper
+
+def getAttributes(ruleLine):
+  ruleLine = ruleLine.replace(" ", "")
+  confParsed = ruleLine.split(":")
+
+  confidence = float(confParsed[1])
+
+  ruleParsed = confParsed[0].split("=>")
+
+  antecedent = strToTupleArray(ruleParsed[0])
+  consequent = strToTupleArray(ruleParsed[1])
+
+  return antecedent, consequent, confidence
+
+def getMatchingLen(antecedent, checkRegions):
+  # We have antecedent A1 A2 ... An
+  # and regions R1 R1 ... Rm
+  # What is the longest sequence of A's present in R's?
+  # Note that the sequence has to be ordered too
+  maxLen = 0
+  i = 0
+  j = 0
+
+  while i < len(checkRegions):
+    currLen = 0
+    currItem = checkRegions[i]
+    if currItem in antecedent:
+      index = antecedent.index(currItem)
+      j = i
+
+      while j < len(checkRegions) and index < len(antecedent) and checkRegions[j] == antecedent[index]:
+        j += 1
+        index += 1
+        currLen += 1
+
+    if currLen > maxLen:
+      maxLen = currLen
+    i += 1
+
+  return maxLen
+
+def strToTupleArray(arrayString):
+  # May have to fix when we see tuples of size > 1
+  # some code repitition here but that's alright for now
+  tupleArray = []
+  result = arrayString.replace("'", "").replace("[", "").replace("]", "")
+  rawTuples = re.findall("\([0-9-,]+\)" , result)
+  for rawTuple in rawTuples:
+    leftT, rightT = rawTuple[1:len(rawTuple)-1].split(",")
[…]

### 2017-12-09 project_final/run_experiment.sh

@@ -1,6 +1,0 @@
-#!/bin/sh
-python scripts/generate_frequent_patterns.py data/vectors_training_1950_2000.txt output/patterns.txt 37.75 1 1
-
-python scripts/generate_assoc_rules.py output/patterns.txt output/rules.txt 0.25
-
-python scripts/test_best_fit.py data/vectors_testing_2001_2015.txt output/rules.txt 1

### 2017-12-09 project_final/scripts/apriori_utils.py

@@ -1,63 +1,0 @@
-
-LATITUDE_DIVISIONS = 11
-LONGITUDE_DIVISIONS = 18
-
-LAT_MAX_DEGREES = 50.0
-LAT_MIN_DEGREES = 0.0
-LONG_MAX_DEGREES = 100.0
-LONG_MIN_DEGREES = 20.0
-
-DIMENSIONS_LATITUDE = 5
-DIMENSIONS_LONGTITUDE = 5
-
-def gatherCoordinateList(inputStr):
-  parseStr = inputStr.split(' ')[1]
-  if parseStr[-1] == ',':
-    parseStr = parseStr[:-1]
-
-  result = []
-  parseStr = parseStr.strip()
-
-  for vecStr in parseStr.split(','):
-    if len(vecStr):
-      currVec = vecStr.replace('(', '').replace(')', '').split(':')
-      result.append( (float(currVec[0]), float(currVec[1])) )
-
-  return result
-
-def regionalize(coordinate):
-
-  # lat, N/S, y, 0
-  # lon, W/E, x, 1 (negative)
-  # Latitude is counted as degrees N/S of equator => y coords
-  # Longitude is counted as degrees W/E of meridian => x coords
-
-  # may need region pruning here
-
-  return (int(coordinate[0]/DIMENSIONS_LATITUDE), int(coordinate[1]/DIMENSIONS_LONGTITUDE))
-
-def gatherDedupedRegionList(ruleLine):
-  coordinates = gatherCoordinateList(ruleLine)
-  regionList = [regionalize(coord) for coord in coordinates]
-
-  dedupedRegions = []
-  # TODO: change this once we have bounds checking for data
-  i = 0
-  while i < len(regionList) and regionList[i] < 0:
-    i += 1
-  dedupedRegions.append(regionList[i])
-  foundItem = regionList[i]
-
-  while i < len(regionList):
-    if regionList[i] != foundItem and regionList[i] >= 0:
-      foundItem = regionList[i]
-      dedupedRegions.append(regionList[i])
-    i += 1
-
-  return dedupedRegions
-
-def safe_increment(table, key, value=1.0):
-  try:
-    table[key] += value
-  except KeyError:
-    table[key] = value

### 2017-12-09 project_final/scripts/generate_assoc_rules.py

@@ -1,112 +1,0 @@
-# Generates association rules given a list of frequent itemsets
-# It is assumed that each line in the file is of the form (IS) : Supp
-# Where IS is a ", " delimited series of numbers, which correspond to regions
-# Ignore singletons, they cannot produce rules
-
-import sys
-import time
-import re
-
-def getRawRules(itemset):
-  divider = len(itemset) - 1
-  rules = [] # Antecedent, Consequent pairs
-  while divider > 0:
-    antecedent = itemset[0:divider]
-    consequent = itemset[divider:]
-    rules.append( (antecedent, consequent) )
-    divider -= 1
-
-  return rules
-
-def convertToSearchLine(antecedentArray):
-  if len(antecedentArray) == 1:
-    line = antecedentArray[0]
-  else:
-    line = "(" + ", ".join(map(str, antecedentArray)) + ")"
-
-  # Will have to change this to parse strings correctly
-  return str(line).replace("'", "")
-
-def getAntecedentSupport(itemsetsFile, searchLine):
-  with open(itemsetsFile, "r") as itemsets:
-    for line in itemsets:
-      contents = line.split(":")
-      if contents[0].strip() == searchLine:
-        result = float(contents[1].strip())
-
-  return result
-
-def generateRules(itemsetsFile, currItemsetLine, minconf=0.25): # String line of form (t, t, t) : s
-  generatedRules = []
-
-  currItemsetLine = currItemsetLine.replace(" ", "")
-  currItemsetParsed = currItemsetLine.split(":")
-
-  currItemsetSupport = float(currItemsetParsed[1])
-
-  # Change this to list of tuples
-  currItemset = convertToTupleList(currItemsetParsed[0])
-  # Now we have the full itemset, and its support
-  # These are ordered rules, so we grow the antecedent and shrink the consequent together
-  # ABC, AB>C, A>BC
-  rulePartitions = getRawRules(currItemset)
-
-  # To find confidence, we need support of antecedent
-  for rule in rulePartitions:
-    searchLine = convertToSearchLine(rule[0])
-    # Do a search in the file for the corresponding frequent itemset, and get its support
-    # That gives the confidence of the rule to be inserted
-    antecedentSupport = getAntecedentSupport(itemsetsFile, searchLine)
-
-    conf = float(currItemsetSupport) / float(antecedentSupport)
[…]

### 2017-12-09 project_final/scripts/generate_frequent_patterns.py

@@ -1,172 +1,0 @@
-
-from apriori_utils import *
-import sys
-import itertools
-import time
-
-WEIGHT_INTERVAL_NOMINAL = 756
-WEIGHT_INTERVAL_YEAR = 50
-HURRICANE_ID_PLACEHOLDER = "AL015555"
-WEIGHT_SCALE = 2.5
-# For weighted data:
-# Data can be weighted nominally or year-over-year
-# Careful! Right now these values are dependent on the specific training data input. This must be fixed
-# A good minsupport value for weight training data appears to be 10. Minconf can stay the same
-
-def gather_new_frequent_items(resultTable, frequent_k_itemsets, minsup):
-  resultTable.append({})
-  for key in frequent_k_itemsets:
-    if frequent_k_itemsets[key] >= minsup:
-      resultTable[-1][key] = frequent_k_itemsets[key] 
-
-def timeseries_candidate_generation(frequent_k_itemsets, chunk_size):
-  # this is the candidate generation and pruning
-  # We have our L_k, use suffix-prefix to get C_k+1
-
-  frequent_k_itemsets_dict = { tuple(item):None for item in frequent_k_itemsets }
-  candidates = []
-  for curr_index, curr_item in enumerate(frequent_k_itemsets):
-
-    compare_index = 0
-    found = False
-    leftChunk = curr_item[(chunk_size*-1):]
-
-    while compare_index < len(frequent_k_itemsets) and not found:
-      found = leftChunk == frequent_k_itemsets[compare_index][:chunk_size]
-      if not found:
-        compare_index += 1
-
-    while compare_index < len(frequent_k_itemsets) and leftChunk == frequent_k_itemsets[compare_index][:chunk_size]:
-      new_candidate = join_entries(curr_item, frequent_k_itemsets[compare_index], chunk_size)
-      required_itemset_checks = itertools.combinations(new_candidate, chunk_size+1)
-
-      if all(item == curr_item or item == frequent_k_itemsets[compare_index] or tuple(item) in frequent_k_itemsets_dict for item in required_itemset_checks):
-        candidates.append(new_candidate)
-      compare_index += 1
-
-  return candidates
-
-def join_entries(left, right, chunk_size):
-  result = left + right[chunk_size:]
-  return result
-
-def gather_sublists(region_list):
-  result = []
-  for size in range(1, len(region_list)):
-    for index in range(len(region_list)):
-      result.append(region_list[index : index + size])
[…]

### 2017-12-09 project_final/scripts/test_best_fit.py

@@ -1,399 +1,0 @@
-# Given an incoming trajectory:
-# Cut it into two pieces: check, confirm
-# Check "check" against rule antecedents, get best fit (or show "missing")
-# if not missing, see if "confirm" was right
-
-# To do this, we will need to cut the patterns into two pieces. Training and Testing.
-# First try: quarters. 1335 training 445 testing
-
-# First try: divide test line in half. First half looks for antecedent
-
-# Parse a line into the antecedent, list of strings
-
-import math
-import sys
-import time
-import re
-from apriori_utils import *
-
-EXTRAPOLATION_LIMIT = 10
-EARTH_RADIUS = 6371
-MAX_MINIMUM_MATCH = 5 # Max comparison length from paper
-
-def getAttributes(ruleLine):
-  ruleLine = ruleLine.replace(" ", "")
-  confParsed = ruleLine.split(":")
-
-  confidence = float(confParsed[1])
-
-  ruleParsed = confParsed[0].split("=>")
-
-  antecedent = strToTupleArray(ruleParsed[0])
-  consequent = strToTupleArray(ruleParsed[1])
-
-  return antecedent, consequent, confidence
-
-def getMatchingLen(antecedent, checkRegions):
-  # We have antecedent A1 A2 ... An
-  # and regions R1 R1 ... Rm
-  # What is the longest sequence of A's present in R's?
-  # Note that the sequence has to be ordered too
-  maxLen = 0
-  i = 0
-  j = 0
-
-  while i < len(checkRegions):
-    currLen = 0
-    currItem = checkRegions[i]
-    if currItem in antecedent:
-      index = antecedent.index(currItem)
-      j = i
-
-      while j < len(checkRegions) and index < len(antecedent) and checkRegions[j] == antecedent[index]:
-        j += 1
-        index += 1
-        currLen += 1
-
-    if currLen > maxLen:
-      maxLen = currLen
-    i += 1
-
-  return maxLen
-
-def strToTupleArray(arrayString):
-  # May have to fix when we see tuples of size > 1
-  # some code repitition here but that's alright for now
-  tupleArray = []
-  result = arrayString.replace("'", "").replace("[", "").replace("]", "")
-  rawTuples = re.findall("\([0-9-,]+\)" , result)
-  for rawTuple in rawTuples:
-    leftT, rightT = rawTuple[1:len(rawTuple)-1].split(",")
[…]

### 2017-12-09 project_final/scripts/tools/basic_trajectory_draw.py

@@ -1,83 +1,0 @@
-"""
-A Hurricane Trajectory-Rendering script based on examples from matplotlib
-"""
-import numpy as np
-import matplotlib.pyplot as plt
-import sys
-from mpl_toolkits.basemap import Basemap
-from apriori_utils import *
-
-def drawTrajectory():
-  fraction = int(sys.argv[2])
-  showRegions = bool(int(sys.argv[3]))
-  # Lambert Conformal Conic map.
-  m = Basemap(llcrnrlon=-100.,llcrnrlat=0.,urcrnrlon=-20.,urcrnrlat=50.,
-              projection='lcc',lat_1=20.,lat_2=40.,lon_0=-60.,
-              resolution ='l',area_thresh=1000.)
-  # read shapefile.
-  # Need to write a longtitudional line as opposed to a euclidean horizontal line
-
-  with open(sys.argv[1], 'r') as inputFile:
-    i = 0
-    for line in inputFile.readlines():
-      color = 'r'
-      if i % fraction == 0: # Render 1 in fraction trajectories (nonzero)
-        lineVector = line.split(' ')[1].split(",")[:-1]
-
-        parsedVector = [item.replace("(", "").replace(")", "") for item in lineVector]
-        trajectoryVector = [ (float(item.split(":")[0]), float(item.split(":")[1])) for item in parsedVector]
-        gisXCoords = []
-        gisYCoords = []
-        for coord in trajectoryVector:
-          newCoord = m(coord[1], coord[0]) # Flip the input coordinates only on conversion
-          gisXCoords.append(int(newCoord[0]))
-          gisYCoords.append(int(newCoord[1]))
-
-        m.plot(gisXCoords, gisYCoords, linewidth=1.0, color=color)
-      i += 1
-
-
-  # with open(sys.argv[4], 'r') as otherInput:
-  #   i = 0
-  #   for line in otherInput.readlines():
-  #     if True or i % fraction == 0: # Render 1 in fraction trajectories (nonzero)
-  #       lineVector = line.split(' ')[1].split(",")[:-1]
-
-  #       parsedVector = [item.replace("(", "").replace(")", "") for item in lineVector]
-  #       trajectoryVector = [ (float(item.split(":")[0]), float(item.split(":")[1])) for item in parsedVector]
-  #       gisXCoords = []
-  #       gisYCoords = []
[…]

### 2017-12-09 project_final/scripts/tools/find_data_extrema.py

@@ -1,53 +1,0 @@
-def gatherCoordinateList(inputStr):
-  parseStr = inputStr.split(' ')[1]
-  if parseStr[-1] == ',':
-    parseStr = parseStr[:-1]
-
-  result = []
-  parseStr = parseStr.strip()
-
-  for vecStr in parseStr.split(','):
-    if len(vecStr):
-      currVec = vecStr.replace('(', '').replace(')', '').split(':')
-      result.append( (float(currVec[0]), float(currVec[1])) )
-
-  return result
-
-import sys
-
-def get_extrema():
-  minX = 1000000000
-  minY = 1000000000
-
-  maxX = 0
-  maxY = 0
-
-  with open(sys.argv[1], 'r') as vectors:
-    for line in vectors:
-      coords = gatherCoordinateList(line)
-      for co in coords:
-        if co[0] < minX:
-          minX = co[0]
-
-        if co[0] > maxX:
-          maxX = co[0]
-
-        if co[1] < minY:
-          minY = co[1]
-
-        if co[1] > maxY:
-          maxY = co[1]
-
-  print minX, maxX, minY, maxY
-  print maxX - minX, maxY - minY
-
-def main():
-  if len(sys.argv) != 2:
-    print "Usage: {0} <VectorsInput>".format(sys.argv[0])
-  else:
-    get_extrema()
-
-if __name__ == '__main__':
-  main()
-else:
-  print __name__

### 2017-12-09 project_final/scripts/tools/generate_coord_vectors.py

@@ -1,36 +1,0 @@
-# A simple script to parse atlantic_storms.csv into a plaintext list of HurricaneID -> Trajectory
-# Hurricanes with fewer than 3 trajectories are not included in the list, but are recorded as pruned
-import sys
-
-def parseHurricanes():
-  with open(sys.argv[1], 'r') as inputFile, open(sys.argv[2], 'w') as outputFile, open(sys.argv[3], 'w') as prunedFile:
-    currentId = ""
-    currentVector = []
-
-    for line in inputFile.readlines()[1:]:
-      lineId = line.split(",")[1]
-      if lineId != currentId: # New ID!
-        if len(currentVector) >= 3: # Did we build enough data for this hurricane to be interesting?
-          outputFile.write(currentId+" "),
-          for item in currentVector:
-            outputFile.write(item),
-          outputFile.write("\n")
-        elif len(currentId) > 0:
-          prunedFile.write(currentId+"\n")
-
-        currentVector = [] # Print or not, move onto next ID/coordinates
-        currentId = lineId
-
-      outStr = "(" + line.split(",")[6] + ":" + line.split(",")[7] + "),"
-      currentVector.append(outStr)
-
-def main():
-  if len(sys.argv) != 4:
-    print "Usage: {0} <StormsCsvInput> <VectorsOutput> <PrunedStormsOutput>".format(sys.argv[0])
-  else:
-    parseHurricanes()
-
-if __name__ == '__main__':
-  main()
-else:
-  print __name__

### 2017-12-09 project_final/scripts/tools/parse_paper_vectors.py

@@ -1,39 +1,0 @@
-import sys
-
-YEAR_INDEX = 0
-NUM_INDEX = 4
-
-LAT_INDEX = 6
-LON_INDEX = 7
-
-def parseHurricanes():
-  with open(sys.argv[1], 'r') as inputFile, open(sys.argv[2], 'w') as outputFile:
-    currentId = ""
-    currentVector = []
-
-    for line in inputFile.readlines()[1:]:
-      lineSplit = line.split(" ")
-      lineId = lineSplit[YEAR_INDEX] + lineSplit[NUM_INDEX]
-
-      if lineId != currentId: # New ID!
-        outputFile.write(currentId+" ")
-        for item in currentVector:
-          outputFile.write(item)
-        outputFile.write("\n")
-
-        currentVector = [] # Print or not, move onto next ID/coordinates
-        currentId = lineId
-
-      outStr = "(" + lineSplit[LAT_INDEX] + ":" + lineSplit[LON_INDEX] + "),"
-      currentVector.append(outStr)
-
-def main():
-  if len(sys.argv) != 3:
-    print "Usage: {0} <StormsCsvInput> <VectorsOutput>".format(sys.argv[0])
-  else:
-    parseHurricanes()
-
-if __name__ == '__main__':
-  main()
-else:
-  print __name__

### 2018-01-26 paper/manuscript_revised_jan25.tex
Eliminate repition in AprioriAll description

--- 2017-12-09 COMP4710-Bissonnette-Marofke-Cox-Project-Fall2017/paper/comp4710-bissonnette-marofke-cox-manuscript.tex
@@
 
-\author{\IEEEauthorblockN{Jasmin Bissonnette, Caden Marofke, Taylor Cox}
+\author{\IEEEauthorblockN{Taylor Cox, Caden Marofke}
 \IEEEauthorblockA{Department of Computer Science\\
@@
 Winnipeg, Manitoba\\
-Email: \{bissonnj, marofkec, coxt3\}@myumanitoba.ca}}
+Email: \{coxt3, marofkec\}@myumanitoba.ca}}
 
 \maketitle
+
+% Reduce repition: highlight modificatiosn of aprioriall less frequently
+% Clearly state that weighted support is our idea, inspired by some work but we have a diff approach
 
@@
 AprioriAll transforms a transaction database into a \textit{customer} database, where each entry contains a customer and an ordered sequence of purchased items.
-The ordered sequence database is then mined for frequent ordered patterns, or frequent sequences.
-The modification proposed in \cite{major-hurricane-model} directly mines frequent ordered patterns and thus skips the transformation step of AprioriAll.
+The ordered sequence database is then mined for frequent ordered sequences.
 
@@
 The hurricane trajectory model in this work does not implement step (1), as the training data from HURDAT2 was instead translated into a $\langle HurricaneID, CoordinateSequence\rangle$ database as a preprocessing step before any aspects of the model are executed.
+Figure (1) shows an example of how the trajectory database is structured after the initial mapping stage.
+Only two dimensions are needed in the database, being the hurricane ID and the sequence of latitude-longitude coordinates.
+All hurricanes in HURDAT2 have an ID corresponding to the location (i.e.
+the Atlantic), the index ordered by appearance in the hurricane year, and the year itself.
+For example, hurricane Katrina of year 2005 corresponds to AL122005.
 By directly translating the data instead of processing it with AprioriAll, all trajectories in the original database are preserved.
-This includes coordinates which may ultimately be infrequent.
+This includes coordinate sequences which may ultimately be infrequent.
 The sequential mapping of AprioriAll is performed as a preprocessing step so that the database used in the prediction model will always consist of hurricanes and their coordinate sequences.
-This allows the prepared (mapped) database to be mined directly without the need to complete the mapping step every time AprioriAll is executed.
+The prepared (mapped) database may therefore be mined directly without the need for repeated mapping in subsequent executions of AprioriAll.
 
@@
 \end{figure}
-
-Figure (1) shows an example of how the trajectory database is structured after the mapping preprocessing step.
-Only two dimensions are needed in the database, being the hurricane ID and the sequence of latitude-longitude coordinates.
-All hurricanes in HURDAT2 have an ID corresponding to the location (i.e.
-the Atlantic), the index ordered by appearance in the hurricane year, and the year itself.
-For example, hurricane Katrina of year 2005 corresponds to AL122005.
-The preprocessing step proposed is executed exactly once, allowing the hurricane database to be mined without the initial grouping and mapping stage.
 
@@
 Experiments in this work show that all datasets contain a subset of trajectories for which no association rule could be matched.
-While \cite{major-hurricane-model} proposed an extrapolation method to recover from rule-miss-incidents, more work must be completed to reduce the number of non-matchable trajectories.
+While \cite{major-hurricane-model} proposed an extrapolation method to recover from rule-miss incidents, more work must be completed to reduce the number of non-matchable trajectories.
 Many of the mismatched trajectories are outliers with very little conformance to existing hurricane trajectories as seen in figure (3).

### 2018-01-26 winter_term/manuscript_revised_jan25.tex
Eliminate repition in AprioriAll description

--- 2018-01-26 paper/manuscript_revised_jan25.tex
@@
 
-\author{\IEEEauthorblockN{Taylor Cox, Caden Marofke}
+\author{\IEEEauthorblockN{Jasmin Bissonnette, Caden Marofke, Taylor Cox}
 \IEEEauthorblockA{Department of Computer Science\\
@@
 Winnipeg, Manitoba\\
-Email: \{coxt3, marofkec\}@myumanitoba.ca}}
+Email: \{bissonnj, marofkec, coxt3\}@myumanitoba.ca}}
 

### 2018-01-29 paper/manuscript_revised_jan25.tex
Fix repo structure. Clarify uniqueness of weighted data contribution

@@
 \maketitle
-
-% Reduce repition: highlight modificatiosn of aprioriall less frequently
 % Clearly state that weighted support is our idea, inspired by some work but we have a diff approach
-
 \begin{abstract}
@@
 While meteorological methods of hurricane trajectory prediction have been effective in the past, little attention has been paid to hurricane trajectory prediction with respect to historical frequent pattern mining.
-In this work, we propose multiple dimensions for improving the accuracy of data mining-based hurricane trajectory prediction.
-We propose the use of the Haversine distance formula for realistic hurricane trajectory distance comparison.
-We also propose the use of weighted training data to favour recent hurricane trends when determining a trajectory prediction.
+In this work, we present multiple dimensions for improving the accuracy of data mining-based hurricane trajectory prediction.
+We first propose the use of the Haversine distance formula for realistic hurricane trajectory distance comparison.
+We also introduce the use of weighted training data to favour recent hurricane trends when determining a trajectory prediction.
 Finally, we use training data exclusively from the years 1950 to 2000 to reduce the impact of historical climate trends and inaccuracies in hurricane recording methodologies.
@@
 
-To the knowledge of the authors, the existing standard in hurricane trajectory prediction via data mining is established by Dong et al in \cite{major-hurricane-model}.
-The work cited proposes a hurricane trajectory prediction model based on data mining (HTPDM) that uses a modified variant of AprioriAll, a common sequential mining algorithm \cite{AprioriAll-original}.
+The existing standard in hurricane trajectory prediction via association rule mining is established by Dong et al in \cite{major-hurricane-model}.
+The work cited proposes a hurricane trajectory prediction model based on data mining (HTPDM) that uses a modified variant of AprioriAll, a standard algorithm for mining sequential data \cite{AprioriAll-original}.
 HTPDM executes AprioriAll against historical hurricane trajectories from 1900 to 2000 to generate interesting trajectory association rules.
@@
 
-The second approach to increasing trajectory prediction correctness deployed in WARD-HTP is the usage of weighted training data in AprioriAll.
-Applying weights to recorded trajectories in chronological order allows AprioriAll to favour recently produced trajectory patterns by increasing the support and confidence of such recently produced trajectories.
+The second approach to increasing trajectory prediction correctness deployed in WARD-HTP is the use of weighted training data in AprioriAll.
+Applying chronologically-ordered weights to recorded trajectories allows AprioriAll to favour recently produced trajectory patterns by increasing their support.
 While the weighting of training data results in some trajectories being counted with a support greater than 1, this effect is balanced by counting earlier recorded trajectories with a support less than 1.
-In WARD-HTP, we weigh hurricane trajectories on a scale increasing linearly.
+However, the most promising results of this work were produced when the support values for recent trajectories were increased \textit{more} than the support values for early trajectories were decreased.
+
+In WARD-HTP, we weigh hurricane trajectories on a linearly increasing scale.
 This results in a tipping point \textit{T} where records before \textit{T} are underweighted and records after \textit{T} are overweighted.
-Depending on the weighing approach used, the combined weight of all records does not necessarily need to be balanced.
-
+Depending on the weighing approach used, the combined weight of all records does not necessarily need to equal 1.
 The following algorithm applies further modifications to AprioriAll in order to weigh the stored hurricane records.
@@
 Therefore the final record in the HDB is guaranteed to have a weighted support exactly equal to W.
-Alternatively, the support weight modifier could also change every year.
+Alternatively, the support weight modifier may also be changed year by year.
 This means that W would be divided by the number of years which the database records span, and the value of w would increase every time the records reach a new year.
-Though index-based weighting is shown above in algorithm (3), both weighting strategies will be evaluated further in this work.
+Though index-based weighting is shown above in algorithm (3), both year and index-based weighting strategies will be evaluated further in this work.
 
-Under the assumption of chronologically ordered data, introducing weighted support to AprioriAll will result in higher support for patterns which occur in recent years.
-Fewer repeated records will be needed to generate frequent sequences from recent data, and more repeated records will be needed to generate frequent sequence from older data.
+Assuming the input data is ordered chronogically, introducing weighted support to AprioriAll will result in higher support for patterns which occur in recent years.
+Recent trajectories will require less repitition to produce associaton rules, and higher repitition will be needed to generate frequent sequences from older data.
+The approach of chronogically weighted data will produce association rules that favor recent trends in hurricane trajectories.
 Therefore, the introduction of weighted support to AprioriAll is expected to further increase the hurricane trajectory prediction correctness ratio for current and future trajectories being tested upon.
+Publications such as \cite{war-mining} and \cite{cai-weighted-thesis} first introduce the use of weighted support for association rule mining.
+However, to the knowledge of the authors this work is the first to (1) apply weighted support to sequential data, (2) weigh individual records based on a proportional scale and (3) allow records to have a weight less than 1.
 
@@
 \bibitem{earth-radius-kilometres} Chambat, F., and Valette, B. (2001). Mean radius, mass, and inertia for reference Earth models. Physics of the Earth and Planetary Interiors, 124(3), 237-253.
+\bibitem{war-mining} W. Wang, J. Yang and P. Yu Efficient  mining of  weighted association rules (WAR), Proc. of the ACM SIGKDD Conf. on Knowledge Discovery and Data Mining, 270-274, 2000.
+\bibitem{cai-weighted-thesis} Cai, Chun Hing, Ada Wai-Chee Fu, C.
+H.
+Cheng, and W.
+W.
+Kwong.
+Mining association rules with weighted items.
+In Database Engineering and Applications Symposium, 1998.
+Proceedings.
+IDEAS'98.
+International, pp.
+68-77.
+IEEE, 1998.
 \bibitem{hurricane-recording-history} Landsea, C.

### 2018-01-29 paper/manuscript_revised_jan25.tex
minor spelling fixes

@@
 There are two categories of motivation behind this work.
-The first motivation of this work is to add to the total investigation completed in the data mining-based hurricane trajectory prediction subdiscipline.
+The first motivation of this work is to add to the total investigation completed in the data mining-based hurricane trajectory prediction sub-discipline.
 Little research effort has been allocated to this particular problem space of high economic and social impact.
@@
 A block size of 5 corresponds approximately to a real-world block dimension of 500km by 500km.
-The block dimension size is approximate because the mapping between coordinate degrees and kilometres is imprecise, and is dependent on location in the case of longitudal points \cite{lat-long-distance}.
+The block dimension size is approximate because the mapping between coordinate degrees and kilometres is imprecise, and is dependent on location in the case of longitudinal points \cite{lat-long-distance}.
 The approximate size of 500km is similar to the average hurricane total diameter \cite{hurricane-distances}, permitting the coarse-grain expression of hurricane trajectories without undue loss in precision.
@@
 
-Assuming the input data is ordered chronogically, introducing weighted support to AprioriAll will result in higher support for patterns which occur in recent years.
-Recent trajectories will require less repitition to produce associaton rules, and higher repitition will be needed to generate frequent sequences from older data.
-The approach of chronogically weighted data will produce association rules that favor recent trends in hurricane trajectories.
+Assuming the input data is ordered chronologically, introducing weighted support to AprioriAll will result in higher support for patterns which occur in recent years.
+Recent trajectories will require less repetition to produce association rules, and higher repetition will be needed to generate frequent sequences from older data.
+The approach of chronologically weighted data will produce association rules that favour recent trends in hurricane trajectories.
 Therefore, the introduction of weighted support to AprioriAll is expected to further increase the hurricane trajectory prediction correctness ratio for current and future trajectories being tested upon.

### 2018-01-29 paper/manuscript_revised_jan25.tex
Slightly shorten AprioriAll section. Reword weighted data contribution

@@
 
-The hurricane trajectory model in this work does not implement step (1), as the training data from HURDAT2 was instead translated into a $\langle HurricaneID, CoordinateSequence\rangle$ database as a preprocessing step before any aspects of the model are executed.
+The hurricane trajectory model in this work does not implement step (1), as the training data from HURDAT2 was instead translated into a $\langle HurricaneID, CoordinateSequence\rangle$ database as a preprocessing step before launching the model.
 Figure (1) shows an example of how the trajectory database is structured after the initial mapping stage.
@@
 This includes coordinate sequences which may ultimately be infrequent.
-The sequential mapping of AprioriAll is performed as a preprocessing step so that the database used in the prediction model will always consist of hurricanes and their coordinate sequences.
-The prepared (mapped) database may therefore be mined directly without the need for repeated mapping in subsequent executions of AprioriAll.
+The sequential mapping of AprioriAll is performed as a preprocessing step so that the database used in the prediction model will always consist of hurricanes and their coordinate sequences, removing the need for repeated mapping in subsequent executions of AprioriAll.
 
@@
 Publications such as \cite{war-mining} and \cite{cai-weighted-thesis} first introduce the use of weighted support for association rule mining.
-However, to the knowledge of the authors this work is the first to (1) apply weighted support to sequential data, (2) weigh individual records based on a proportional scale and (3) allow records to have a weight less than 1.
+However, to the knowledge of the authors this work is the first to (1) apply weighted support to sequential data, (2) apply weight to records as opposed to individual items and (3) allow records to have a weight less than 1.
 

### 2018-01-29 paper/manuscript_revised_jan25.tex
Keep paper to 14 pages

@@
 In HTPDM and WARD-HTP, the algorithm that generates trajectory prediction possibilities is based on AprioriAll.
-AprioriAll was first proposed at IBM research as a method for discovering frequent sequential patterns in transactional databases \cite{AprioriAll-original}.
-In its original implementation, AprioriAll is able to generate association rules describing sentences of the form ``Customers who buy P may later buy Q''.
+In its original implementation, AprioriAll is able to generate association rules describing sentences of the form ``Customers who buy P may later buy Q'' \cite{AprioriAll-original}.
 In the hurricane trajectory prediction problem, AprioriAll is used to generate association rules describing sentences of the form ``Hurricanes located at P may later be located at Q''.
@@
 In the case of hurricane trajectory prediction, these sequences correspond to frequently recorded hurricane coordinate sequences.
-Frequent hurricane coordinate sequences are those which occur at least a minimum number of times, as defined by the \textit{minimum support}, or minsup.
+Frequent hurricane coordinate sequences are those which have a support (occurance count) greater than or equal to the \textit{minimum support}.
 In general terms, the sequence generation step of AprioriAll develops frequent ordered patterns of increasing length, beginning with frequent single-item sequences (corresponding to frequent single-point sequences).
-This step begins by scanning the $\langle HurricaneID, CoordinateSequence\rangle$ database for frequent single-point sequences.
-The frequent single-point trajectories are then converted to candidate two-point trajectories as the result of an inner join.
+This step begins by scanning the $\langle HurricaneID, CoordinateSequence\rangle$ database for frequent single-point trajectories.
+The frequent single-point trajectories are then converted to candidate two-point trajectories by an inner join.
 The candidate two-point trajectories are scanned against the database to determine which candidates satisfy the minimum support.
-Another inner join generates the candidate 3-length trajectories.
-Once the sequence generation level reaches three, each candidate sequence must be pruned if any of their candidate subsequences are infrequent.
-Continuous execution of AprioriAll until termination results in the collection of all frequent hurricane trajectory patterns defined by the minimum support.
+Running an inner join against the frequent k-point trajectories generates the candidate (k+1)-point trajectories.
+Once the sequence generation level reaches 3, each candidate sequence must be pruned if any of their subsequences are infrequent.
+The complete exectuion of AprioriAll results in a collection of all frequent hurricane trajectory patterns defined by the minimum support.
 
@@
 
-Figure (7) shows the correctness ratios generated by all possible training data and testing data pairings.
-Note that ratios in brackets denote the worse-case ratio while ratios outside of brackets refer to the best-case.
+Figure (7) shows the correctness ratios generated by all pairings of training and testing data.
+Note that ratios in brackets denote the worse-case ratio while unbracketed ratios refer to the best-case.
 Between all possible options, we find that the training data interval of 1950-2000 results in the best correctness ratios for testing data from 2001-2015 when applying all other parameters we deemed optimal throughout the previous results of this work.
 Regardless of correctness ratio, 2001-2015 should always be selected as the test data interval to encourage analysis against more recent hurricane patterns.
-The results show that no training-data/testing-data pairing was able to generate correctness ratios higher than the ones generated by training the model with data from 1950 to 2000, and testing it on data as recently recorded as 2015.
+The results show that no training-data/testing-data pairing was able to generate correctness ratios higher than those generated by training the model with data from 1950 to 2000, and testing it on data from 2001 to 2015.
 
@@
 There are three such contributions brought forth in the development and implementation of the model proposed.
-These contributions are the use of training data from 1950 to 2000, the use of weighted training data, or weighted assets (WA), and the use of the Haversine Distance Formula for realistic distance (RD).
+These contributions are the use of training data from 1950 to 2000, the use of weighted training data, and the use of the Haversine Distance Formula..
 The three contributions combined form WARD-HTP and produce a hurricane trajectory prediction correctness ratio of exactly 75.0\%, which is 10\% higher than the current state-of-the-art.
@@
 
-There are multiple areas to explore in the hurricane trajectory prediction problem space with respect to future work.
+There are multiple aspects of future work to explore in the hurricane trajectory prediction problem space.
 Experiments in this work show that all datasets contain a subset of trajectories for which no association rule could be matched.
@@
 Additionally, more experiments will be run on different inputs including training data from other regions such as the Pacific region.
-Further approaches to increasing the correctness ratio of WARD-HTP will also be investigated.
-This includes the investigation of alternative methods for defining training data weight.
+Further approaches to increasing the correctness ratio of WARD-HTP will also be investigated, including the pursuit of alternative methods for implementing AprioriAll with weighted transactions.
 

### 2018-02-05 paper/manuscript_revised.tex
Rename manuscript to remove specific date

--- 2018-01-29 paper/manuscript_revised_jan25.tex
@@
 \maketitle
-% Clearly state that weighted support is our idea, inspired by some work but we have a diff approach
+
 \begin{abstract}
@@
 A block size of 5 corresponds approximately to a real-world block dimension of 500km by 500km.
-The block dimension size is approximate because the mapping between coordinate degrees and kilometres is imprecise, and is dependent on location in the case of longitudinal points \cite{lat-long-distance}.
+The block dimension size is approximate because the mapping between coordinate degrees and kilometers is imprecise, and is dependent on location in the case of longitudinal points \cite{lat-long-distance}.
 The approximate size of 500km is similar to the average hurricane total diameter \cite{hurricane-distances}, permitting the coarse-grain expression of hurricane trajectories without undue loss in precision.
@@
 
-In contrast to the Euclidean formula, the Haversine Distance Formula is designed to determine the distance in kilometres between two latitude-longitude points over a large sphere, such as the earth's surface.
+In contrast to the Euclidean formula, the Haversine Distance Formula is designed to determine the distance in kilometers between two latitude-longitude points over a large sphere, such as the earth's surface.
 Despite the fact that the earth's surface is not a perfect sphere \cite{lat-long-distance}, the Haversine Distance Formula is sufficiently accurate when used within a specific region such as the Atlantic Basin.
@@
 
-% H is in RADS, Eucl in degrees
 Where $\Delta Lat$ and $\Delta Lon$ respectively denote the differences in latitude and longitude between points q and p.
@@
 
-\section{Conclusions and Future Work} % what we found:
+\section{Conclusions and Future Work}
 
@@
 \bibitem{hurricane-cost-non-financial} F. F. Townsend, The Federal Response to Hurricane Katrina: Lessons Learned, United States Federal Government, Washington, DC, USA, 2006
-\bibitem{data mining-def} S. Sumathi and S. N. Sivanandam, Introduction to Dataming and its Applications: Springer Publishing 2006, ISBN: 978-3-540-34350-9
+\bibitem{data mining-def} S. Sumathi and S. N. Sivanandam, Introduction to Data mining and its Applications: Springer Publishing 2006, ISBN: 978-3-540-34350-9
 \bibitem{major-hurricane-model} Dong, X and C.

### 2018-02-05 paper/manuscript_revised.tex
minor fixes

--- 2018-01-29 paper/manuscript_revised_jan25.tex
@@
 In the case of hurricane trajectory prediction, these sequences correspond to frequently recorded hurricane coordinate sequences.
-Frequent hurricane coordinate sequences are those which have a support (occurance count) greater than or equal to the \textit{minimum support}.
+Frequent hurricane coordinate sequences are those which have a support (occurrence count) greater than or equal to the \textit{minimum support}.
 In general terms, the sequence generation step of AprioriAll develops frequent ordered patterns of increasing length, beginning with frequent single-item sequences (corresponding to frequent single-point sequences).
@@
 Once the sequence generation level reaches 3, each candidate sequence must be pruned if any of their subsequences are infrequent.
-The complete exectuion of AprioriAll results in a collection of all frequent hurricane trajectory patterns defined by the minimum support.
+The complete execution of AprioriAll results in a collection of all frequent hurricane trajectory patterns defined by the minimum support.
 

### 2018-02-15 paper/manuscript_revised.tex
Modify drawing script to render figures with proper dimensions. Save those figures.

@@
 % Figures 2 and 3: Trajectories without and with region overlay
+% \begin{figure}[htp]
+% \caption{Hurricane Trajectory Paths Without and With Regionalization}
+% \centering
+% \includegraphics[scale=0.55]{Merged-Trajectory-Comparison}
+% \end{figure}
+
 \begin{figure}[htp]
@@
 \centering
-\includegraphics[scale=0.55]{Merged-Trajectory-Comparison}
+\includegraphics[scale=0.55]{Trajectories-Proper-Range-Regions-NoRegions-Combined}
 \end{figure}
@@
 Consider figure (2), generated using a sample of the recorded hurricane trajectories in HURDAT2 with the visualization tool Matplotlib \cite{matplotlib}.
+The region of interest is captured within the emboldened sections of the figures.
 As shown in the pair of images, region discretization organizes continuous hurricane coordinate sequences into discrete region sequences of approximately uniform size and square shape.
@@
 \centering
-\includegraphics[scale=0.6]{Rule-Miss-Diagram}
+\includegraphics[scale=0.6]{Trajectories-Proper-Range-Show-Miss}
 \end{figure}
@@
 Figure (3) illustrates the characteristics of trajectories which were not able to match with any interesting rules.
+As seen in figure (2), the area of interest to this work is captured within the bold area.
 The particular diagram was generated from training data ranging from 1950 to 2000, and testing data from 2001 to 2015.
@@
 It is visually clear to see that non-matching trajectories were highly dissimilar from training data trajectories.
-As a result, no amount of point extrapolation would be able to result in a reliable terminal trajectory prediction.
+As a result, no amount of point extrapolation would be able to provide a reliable terminal trajectory prediction.
 Some of the non-matching trajectories appear to align with training trajectories, but are moving in an opposite direction.

### 2018-02-15 COMP4710-Bissonnette-Marofke-Cox-Project-Fall2017/scripts/tools/basic_trajectory_draw.py

@@ -6,13 +6,14 @@ import matplotlib.pyplot as plt
 import sys
 from mpl_toolkits.basemap import Basemap
 from apriori_utils import *
+from matplotlib.patches import Rectangle 
 
 def drawTrajectory():
   fraction = int(sys.argv[2])
   showRegions = bool(int(sys.argv[3]))
   # Lambert Conformal Conic map.
-  m = Basemap(llcrnrlon=-100.,llcrnrlat=0.,urcrnrlon=-20.,urcrnrlat=50.,
-              projection='lcc',lat_1=20.,lat_2=40.,lon_0=-60.,
+  m = Basemap(llcrnrlon=-101.,llcrnrlat=-5.5,urcrnrlon=-0.,urcrnrlat=50.9,
+              projection='lcc',lat_1=0.,lat_2=39.,lon_0=-60.,
               resolution ='l',area_thresh=1000.)
   # read shapefile.
   # Need to write a longtitudional line as opposed to a euclidean horizontal line
@@ -37,38 +38,80 @@ def drawTrajectory():
       i += 1
 
 
-  # with open(sys.argv[4], 'r') as otherInput:
-  #   i = 0
-  #   for line in otherInput.readlines():
-  #     if True or i % fraction == 0: # Render 1 in fraction trajectories (nonzero)
-  #       lineVector = line.split(' ')[1].split(",")[:-1]
+  with open(sys.argv[4], 'r') as otherInput:
+    i = 0
+    for line in otherInput.readlines():
+      if True or i % fraction == 0: # Render 1 in fraction trajectories (nonzero)
+        lineVector = line.split(' ')[1].split(",")[:-1]
 
-  #       parsedVector = [item.replace("(", "").replace(")", "") for item in lineVector]
-  #       trajectoryVector = [ (float(item.split(":")[0]), float(item.split(":")[1])) for item in parsedVector]
-  #       gisXCoords = []
-  #       gisYCoords = []
-  #       for coord in trajectoryVector:
-  #         newCoord = m(coord[1], coord[0]) # Flip the input coordinates only on conversion
-  #         gisXCoords.append(int(newCoord[0]))
-  #         gisYCoords.append(int(newCoord[1]))
[…]
